\clearpage

# Methods

## Revised Chinook catch estimates from 2013-present

### Background on survey methods

We first provide a brief background on the regionally-specific creel and logbook, programs and the coast-wide iREC surveys in and how each contribute to Chinook catch estimates in a given month. More details on each method can be found in the relevant technical reports (*Cite: Houtman et al. 2015, English et al. 2002, Korman et al. 2005, Sturhahn and Nagtegaal 2001*). We note any improvements to the data quality assurance and control that may have occurred in order to create the datasets used in this analysis. While all methods report on multiple species, here we are concerned only with data involving Chinook salmon. The coastal area of interest is all Pacific Fishery Management Areas (PFMAs, Figure \@ref(fig:fig-PFMA)) , more information and detailed maps can be found [here](https://www.pac.dfo-mpo.gc.ca/fm-gp/maps-cartes/areas-secteurs/index-eng.html).

```{r myfigure, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-PFMA
#| fig-cap: "Pacific Fishery Management Areas"

fig1_path <- paste0("data/PFMA.png")

knitr::include_graphics(c(fig1_path))
```

*Note: references need to be added to bib and linked*

#### South Coast Catch Estimation *Kris/Rob/Aswea to review this section*

South Coast catch estimation is a combination of creel surveys and logbook and avid angler programs. The South Coast Creel Survey (SCCS), run by the department, began in the early 1980s with a limited scope of Strait of Georgia recreational fisheries. It has since expanded its scope to surveys in PFMAs 10 to 27, including offshore areas (start with a "1"). The surveys are generally concentrated in the summer months, however some spring and fall surveys occur in specific areas. The survey involves dock-side fisher interviews and aerial surveys for effort counts. Raw estimates and uncertainty are expanded to cover a given month and PFMA. Currently, this program produces separate catch estimates for marked and unmarked fish, for kept and released fish, and for legal and sublegal sized fish. In the past, unchecked kept fish were expanded to marked and unmarked using a monthly mark rate and to kept and released using a monthly release rate. For more information and methods about the SCCS, see the most recent program overview and bulletins, found [here](https://cat.fsl-bsf.scitech.gc.ca/record=b4125212~S1) (*Cite: South coast assessment bulletin, program overview 2024*). While the SCCS is the primary source of data for the South Coast, data from logbook and avid angler programs are provided on a voluntary basis and are detracted from area and month-specific estimates of the creel survey to avoid double-counting.

##### South Coast Creel Survey quality control

Since the South Coast creel surveys vary in sampling intensity and spatial and temporal coverage, we have introduced a systematic quality control of these estimates. Following the method used for iREC calibration to creel (*cite DFO iREC report - is this already published or is this the first time creel quality control published?*), we filtered the creel data based on quality criteria in a given area and month. These quality control criteria are designed to avoid uncertainty where a monthly estimate was generated by expanding from only a small number of days or bias when generated only from a certain type of day (e.g. weekends). Each area and monthly creel estimate included in our data was required to meet each of the following criteria:

*Note: Kris and Rob to flesh out these bullet points .... Why was at least 3 flights for each type of day chosen? Is that because in a month you typically get x amount of flights? Need to flesh out each bullet point with context.*

-   At least 3 flights for each type of day (weekday or weekend)
-   At least 25 interviews mid week OR at least 10% of interviews from mid-week
-   At least 25 interviews on weekends OR at least 10% of interviews from weekend
-   At least 15 day spread in flights
-   At least 15 day spread in interviews

The criteria failed 51% of all estimates since 2005 (Figure \@ref(fig:fig-creel-criteria)). Areas 15 and 16 had the highest proportion of failed criteria, followed by Area 29, 19 (GS) and 21, which all had over 80% of estimates fail the criteria. The core summer months July and August have the highest rate of criteria passing (more than 70% of estimates passing), whereas the winter months (October through February) have more than 75% failed criteria, likely due to weather making consistent overflights difficult.

```{r methods Loading packages useful functions, message=FALSE, warning=FALSE, results='hide', echo=FALSE}

library(ROracle)
library(pacRecCatch)

library(tidyverse)
library(odbc)
library(lubridate)
library(ggnewscale)
library(patchwork)
library(DHARMa)
library(lme4)
library(bbmle)
library(SuppDists)
library(MuMIn)

col1 = "#d8e1cf"
col2 = "#438484"
col3 = "lightblue"
col4 = "lightblue4"
col5 = "pink"
col6 = "darkred"
col7 = "purple"
col8 = "purple4"
colclear= "#1C00ff00"

#utils
"%notin%" <- Negate("%in%")


```

```{r methods plotting PFMA creel unfiltered, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12, fig.width=15, out.width = "7in", fig.align = "left"}
#| label: fig-creel-criteria
#| fig-cap: "Creel estimates that passed (green) or failed (red) quality control criteria by month and PFMA."

creel_month_PFMA<-filter_creel_catch("pacRecCatch_db.yaml")
Creel_month_PFMA_failed<-creel_month_PFMA %>% filter(criteria == "failed")
Creel_month_PFMA_passed<-creel_month_PFMA %>% filter(criteria == "passed")

g<-ggplot()+
  geom_tile(data=Creel_month_PFMA_passed %>% mutate(across(where(is.numeric), ~na_if(., 0)))  %>%  filter(AREA %notin% c("Area 3", "Area 4")), aes(x=YEAR, y=as.factor(MONTH), fill = VAL),colour = "white") +
  scale_fill_gradient(low = col1, high = col2,  na.value = colclear, limits=c(0,20000)) +
  scale_y_discrete(limits=rev)+
  guides(fill=guide_legend(title="Creel estimates that passed criteria")) +
  new_scale_fill() +
  geom_tile(data=Creel_month_PFMA_failed %>% mutate(across(where(is.numeric), ~na_if(., 0))) %>%  filter(AREA %notin% c("Area 3", "Area 4")), aes(x=YEAR, y=as.factor(MONTH), fill = VAL),colour = "white") +
  scale_fill_gradient(low = col5, high = col6,  na.value = colclear, limits=c(0,20000))+
  facet_wrap(~AREA)+
  guides(fill=guide_legend(title="Creel estimates that failed criteria")) +
  labs(title = "Coverage by PFMA",
       x = "Year", y = "Month") +
  theme_bw() + theme_minimal() + theme(legend.position="bottom")

g
```

#### Central Coast Catch Estimation - *Chelsea to review/expand this section*

Central Coast catch estimation is exclusively logbook data from lodges. The PFMAs included in central coast are 6, 7, 8, 9 and their outer counterparts 106, 107, 108, 109. This data is typically not by marked and kept status so those distinctions need to be reconstructed using mark rate and release rates.

-   *Do they adjust the catch for like 80% of lodges reported for eg. I.e. is this NOT an expanded estimate?*

-   *What about CCIRA? Note the development of CCIRA but that data not yet available to be incorporated.*

#### North Coast Catch Estimation - *Chelsea and Jessica to review/expand*

The North Coast catch estimation is a combination of independently-run creel programs and logbook data. Northern BC Abundance Based Fishery is comprised of PFMAs 1, 2 (East and West), 101, 102, 142. It is a combination of the Haida creel program and lodge-submitted logbooks. NBC ISBM is comprised of the Skeena creel program and logbooks run in PFMAs 3 and 4 (and 103, 104). Technically PFMA 5 and 105 is included in this category but there is no data from Area 5. Data for before 2008 is not available. As in Central coast, typically this data is typically not collected by marked and kept status so those distinctions need to be reconstructed using mark rates and release rates.

-   *North coast description of standalone programs - expand on the Haida creel and the 3/4 creel details, who are they run by, when the programs were started, when are they typically run (time of year).*
-   *description of how creel and logbook are combined*
-   *Do they adjust for % lodges reporting?*
-   *Explain if/where marked and kept status data is collected and since when?*

#### Internet Recreational Catch (iRec) Survey - *Rob/Nick/Catarina to review this section*

*Note: this section could be shorter? We don't need to go into full details?*

The mandatory internet recreational effort and catch (iREC) survey is a recent method for estimating recreational fishing effort and catch in Pacific Region tidal waters. The iREC survey annually provides catch estimates and associated precision across all tidal water Pacific Fishery Management Areas (PFMA's), licenced fishing methods, species and fates (retained or released) at a monthly resolution (Houtman et al, 2015). For all fishing methods besides boat-based angling -- specifically angling from shore, shellfish trapping from boat or shore, beach collecting, hand picking, diving, and other methods -- these are the first estimates available.

##### Survey License Selection

When the iRec survey initially began in 2012, survey participants were selected from a provided subset of licence holders that purchased through a web based license purchasing system. This subset excluded any paper-based licences that were issued outside the web-based system and typically consisted of approximately XX% of all recreational licence holders. This subset of licence holders were contacted by email with instructions on how to ....

*Note: this sentence sort of just cut off - Nick?*

In more recent years (since 2018??) the iREC survey takes advantage of the web based purchasing of recreational licences to communicate survey selection and creating a sampling frame. Random selections of fishers with licences active in a given month are invited, via email, to provide details about their fishing activity and catch through a web-based survey tool. This fishing information from survey respondents is expanded to estimate total effort and catch following the methods described by Houtman et al. (2015).

The iRec survey is designed to provide month recreational catch estimates for every area within DFO's Pacific Region. When the survey was initially designed, the survey used the legal definition of Pacific Fishery Management Areas (PFMA) within the @pfma_2007.

##### License Stratification

When assessing survey responses, annual licences are unique compared to term licences as they can potentially fish any date after the licence purchase date and may be selected for one of several survey months within the licence year. To improve the representativeness of annual licence holders and provide a more consistent precision across months within the year, the licence selection process has used different stratification methods over the years the iRec has operated. Over the years, the approach to stratification has be adjusted to better align to improvements the in licence selection process and the increased number of selected licenses.

At the beginning of the iRec survey in 2012, licences were selected monthly, after the licence was purchased. Based on whether the annual licence was selected to respond to the survey for the month it was purchased in or for a survey month and were stratified into pre- and post-selection strata based on the licence purchase date. While nearly all term licences fished near the licence purchase date, the were all stratified as post-selection. However, annual licences may be selected for either a month they purchased the licence in or a future survey month. Selected annual licences purchased date was during the survey month, the licence was estimated in the post-selection strata for the survey month. For example if a licence was purchased on August 15th and they were selected for the August survey, the licence was stratified as a post-selection licence they in relation to their selected survey period. If the licence was purchased during the ...

*Note: this sentence sort of just cut off - Nick?*

Despite this potential bias, the iREC estimates are currently used, in a relative way, to prioritize the allocation of resources of creel surveys. Creel surveys are funded through contributions from Science (Salmon Stock Assessment) and Fisheries Management branches of Fisheries and Oceans Canada, ostensibly for the estimation of recreational salmon and Halibut catches although estimates for other finfish species are made. The iREC estimates of catch of these two species, along with boat-based recreational angling effort, are used to rank the most important month and area combinations for creel survey allocation purposes. Currently, creel surveys are focused on only those month -- area combinations contributing the highest combined retained catches of Chinook Salmon and Halibut. In times and locations where fishing effort and catch of these important species are low, creel surveys are not currently conducted (Houtman et al. 2015). In fact, for many month -- PFMA combinations, creel surveys or other recreational monitoring approaches have never been used to estimate recreational fishing effort or catch. The apparent bias in iREC estimates could result from numerous sources. With an invitational fisher-dependent survey of this type, there is potential that the respondents to the survey are not representative of all licenced fishers. Such 'response bias' could result from differential response rates by successful or unsuccessful fishers, active or inactive fishers, people who tend to inaccurately report their catch for prestige or strategic reasons as well as problems with recalling the details of their fishing activity. The survey is designed to minimize bias where possible, but not all sources of bias can be completely addressed through design or education. Without a source of alternative estimates of high accuracy, the overall magnitude and direction of bias is difficult to quantify.

Total annual catch estimates of boat-based angling would thus be equal to the sum of estimates from creel surveys for all months and areas where they occurred plus the iREC\* estimates for all months and areas where they did not. Despite the apparent simplicity of this calibration approach, there are outstanding issues regarding the use of creel survey estimates to calibrate iREC estimates; three of which we address here. First, creel surveys vary in sampling intensity and spatial and temporal coverage even with a particular month -- area estimate strata. This variability in the 'quality' of creel estimates (both in terms of accuracy and precision) may adversely affect calibration of iREC estimates. Second, variability in the relative strength and direction of "response bias" may influence the spatial or temporal patterns of bias in iREC estimates, with important implications for calibration. Finally, we combine information about the previous two issues to make recommendations about where, when and how to conduct relatively costly creel surveys to maximize their calibration value.

##### iRec to Creel Calibration

*moved this section back to under the "irec" section*

Where absolute estimates of effort and catch are demanded, such as for allocation decisions, stock assessments and catch reporting requirements, calibration of iREC estimates would be required. Calibrated iREC estimates for boat-based angling would be achieved by scaling iREC to creel estimates using the regression parameters (slope [m]) of iREC -- creel regressions for equivalent month -- area pairs. The calibrated iREC point estimates of boat-based angling effort or species specific catch (iREC\*) would be simply:

$$iRec^*=\frac{(iRec)}{m}$$

Use of JAGS

```{r methods jags, echo=TRUE}
model.jags <- function() {   for(i in 1:Ndata) {      
  #generate what true value may be based on se (notice that dnorm in jags      
  # uses 1/variance for second argument thus 1/SE^2)      
  creel.true[i] ~ dnorm(creel.obs[i], 1/creel.se[i]*1/creel.se[i])            
  #now for the actual regression equation     
  mu[i] <- beta1 * creel.true[i]      
  iRec.obs[i] ~ dnorm(mu[i], tau)   }      
  beta1 ~ dnorm(0, 0.0001)   
  tau ~ dgram(0.01, 0.01) 
  #a common prior for a variance term in vertical direction     
  #derived variables   
  sigma <- 1/sqrt(tau) 
  #std deviation about the regression line 
  }
```

*Notes to add:*

-   *irec calibrated using south coast creel only*

-   *Forced through the origin (need background?)*

-   *Discuss separate calibration from kept and released?*

### Assembling and combining survey methods for best catch estimates (2013 - present)

In the past, iREC has been incorporated into domestic catch estimation products in various ways, though an infilling method (not described here) for South Coast products, and at a high stratification level for the CTC's catch and escapement report. Neither of these methods were robust or systematic approaches to incorporating iREC at the finest scale possible. Here, we outline a systematic method for combining survey methods to obtain the best catch estimate in a given month and PFMA.

We assembled and combined recreational Chinook catch data by developing the [pacRecCatch](https://github.com/nick-komick/pacRecCatch) R package. This package has several functions to pull data from the Catch and Released Estimates (CREST) database and load in other sources manually. The package has functions that will then clean, filter, align, and combine the various survey methods to the final dataset as described below.

We pulled raw recreational catch data from the Catch and Release Estimates (CREST) database. This database stores creel, logbook, and iREC data for both marine and freshwater regions. We filtered the data to contain only marine estimates for legal-sized marked and unmarked fish that were kept and released. We loaded in data from Northern BC (creel and logbook estimates 2005-2023) and Central BC (logbook only estimates 2022-2023) separately since the information in CREST is incomplete for these regions. We filtered the south coast creel data as described above to only include high quality creel estimates. These estimates (which don't include submissions from logbooks) were then augmented with logbook data where available (logbook and creel estimates were summed).

#### Stratification

In Table \@ref(tab:strattab), we show the stratification of fisheries to PFMAs and months. In order to complete this work and to fufill Treaty obligations to incorporate Mark Selective Fisheries, we recently reogranzied the fishery stratification levels used by the CTC. PFMAs are used domestically to collect the data, while finescale and exploitation rate analysis fisheries are groupings of fisheries used for CTC analyses. New Finescale Fisheries are a grouping of PFMAs under a given season, they were introduced recently and are the level at which marked release rates and unmarked release rates are generated. The Old Finescale Fisheries were the previously used groupings by the CTC and are shown for consistency. The ERA fisheries are the exploitation rate analysis fisheries and are the level at which CTC reporting happens and exploitation rate calculations happen.

*Notes:*

-   *may want to have detailed description of fisheries - like ERA level BC Juan de Fuca ....*

-   *May want to put this in an appendix instead, it takes up 3 pages.*

-   *May want to drop left-hand levels ... should be able to do this with Kable extra group but couldn't quite get it to work.*

```{r methods strattab, echo=FALSE, message=FALSE, warning=FALSE}
#| label: strattab

Sport_mark_rate_finescale<-wrangle_rec_catch("pacRecCatch_db.yaml", by="area")

Strat_table<-Sport_mark_rate_finescale %>% dplyr::select(finescale_fishery_old, finescale_fishery, AREA, MONTH) %>% unique() %>% arrange(finescale_fishery_old, finescale_fishery,  AREA, MONTH)

fishery_map<-Sport_mark_rate_finescale %>% dplyr::select(finescale_fishery_old, finescale_fishery) %>%
  mutate(erafishery = case_when(
    finescale_fishery_old %in% c("NWCVI S AABM", "SWCVI S AABM") ~ "WCVI AABM S",
    finescale_fishery_old %in% c("NGS S", "SGS S") ~  "GEO ST S",
    finescale_fishery_old == "CA JDF S" ~ "BC JF S",
    finescale_fishery_old %in%  c("NWCVI S ISBM", "SWCVI S ISBM") ~ "WCVI ISBM S",
    TRUE ~ finescale_fishery_old
  )) %>% unique()

Strat_table<-Strat_table %>% left_join(fishery_map) %>% arrange(erafishery) %>% relocate(erafishery) %>%                group_by(finescale_fishery) %>% 
             mutate(month_category = case_when(
              finescale_fishery == "CBC S" ~ "1:12",
              finescale_fishery == "NWCVI S SUMMER AABM" & AREA %in% c("Area 25", "Area 26", "Area 27") ~ "5:6",
              finescale_fishery == "NWCVI S SUMMER ISBM"  ~ "7:9",
               finescale_fishery == "SWCVI S SUMMER AABM" & AREA %in% c("Area 21", "Area 23", "Area 23 (Alberni Canal)", "Area 23 (Barkley)", "Area 24") ~ "5:7",
              finescale_fishery == "SWCVI S SUMMER ISBM"  ~ "8:9",
              MONTH %in% c(10,11,12) ~ "10:12", 
              MONTH %in% c(1:4) ~ "1:4", 
              MONTH %in% c(5:9) ~ "5:9",
              MONTH %in% c(7:8) ~ "7:8",
              TRUE ~ "NA"
             )) %>% 
             dplyr::select(-MONTH) %>% unique() %>% drop_na() %>% 
             rename("ERA Fishery" = "erafishery", "Finescale Fishery Old" = "finescale_fishery_old", 
                    "Finescale Fishery New" ="finescale_fishery", "PFMA" ="AREA","Months" ="month_category")
              


csasdown::csas_table(Strat_table, caption = "Stratification mapping") %>% 
  kableExtra::collapse_rows(columns = 1:3)

```

#### Alignment issues

iREC and creel both operate on the Pacific Fisheries Management Area system. However, there are a few areas that were initially not split by the iREC system into their component parts, as done in the creel database. These include areas 2, 19, 20, and 23. We therefore combined data from 2013 until the split was successfully incorporated into iREC (licence year 2014 for Areas 2,19, and 23; licence year 2020 for Area 20).

*Note:*

-   *Add something about ISBM corridor alignment issue and how we handle it -i.e. iREC is just the whole PFMA whereas creel is specific sub-areas*

-   *Kris Hein has a good explanation for the alignment problem (creel surveys aerial need to be not on PFMA boundaries where a lot of people fish) ---- creel sub areas vs. pfma sub-areas*

#### Extrapolating unchecked catch by mark rate

*Note: this section could be moved above to the data sources section ... this essentially is appliued to south coast creel data but we did a similar thing to the north and central data just by hand using mark rates from MRP and release rates estimated by the programs.*

All data is typically collected in number of marked and unmarked fish kept and released. However, when mark status is unknown, the catch estimates can be provided as unchecked kept and unchecked released catch. We extrapolated these unchecked estimates to marked and unmarked catch using mark rate.

*Note: add in Nick's MRP mark rate data to this list. Is it preferred? Is it from biodata?*

We calculated mark rate in a number of different ways and prioritize the assignment of mark rate in the following order:

1.  Monthly area calculation - Mark rate calculated for a given source (iREC, creel, or logbook) within a given year, month, and area
2.  Monthly average by source - Mark rate calculated within a given year, month, and area, averaged across sources
3.  Monthly regional average - Mark rate calculated for a given source within a given year, month, and region (averaged across areas within a region)
4.  Seasonal area average - Mark rate calculated for a given source (iREC, creel, or logbook) within a given year, season, and area
5.  Seasonal regional average -Mark rate calculated for a given source within a given year, season, and region
6.  Monthly area average across years - Mark rate calculated for a given source within a given month, area, and across years
7.  Monthly regional average across years- Mark rate calculated for a given source within a given month, region, and across years

#### Choosing best monthly catch estimate per PFMA

We used the following rules to choose the best catch estimate between creel (augmented with logbook data) and calibrated iREC:

For Southern BC:

-   In months May-September (5-9) use creel+ logbook, otherwise use calibrated iREC

-   In months outside of 5-9, use calibrated iREC

For Northern and Central BC:

-   Use calibrated iREC from 2015 onwards

These rules prioritize quality-controlled creel estimates during months where the most creel effort is concentrated, and then prioritizes calibrated iREC data during shoulder seasons and anywhere there is a gap in creel coverage. Northern and Central BC do not have standardized creel, and therefore the iREC data is preferred over logbook-only data. iREC took longer to become adopted in Northern BC and therefore data from 2015 onwards is used.

In Figure \@ref(fig:fig-creel-irec-JDF) we demonstrate how iREC fills in the gaps in each PFMA for a given finescale fishery (Juan De Fuca shown as an example).

*Note: fix the scale of this plot - consider using an areas that's not JDF and doesn't have the weird 20 split*

```{r methods pfma_month again, message=FALSE, warning=FALSE, results='hide', echo=FALSE}

Sport_mark_rate_finescale<-wrangle_rec_catch("pacRecCatch_db.yaml", by="area")

pfma_yearMonth <- plyr::ddply(Sport_mark_rate_finescale, c( "YEAR", "MONTH", "AREA", "finescale_fishery_old"), summarise,
                         sum_creel = sum(creel, na.rm = TRUE), sum= sum(creel_plus, na.rm = TRUE), sum_historic= sum(historic_plus, na.rm = TRUE), sum_catch_estimate = sum(catch_estimate, na.rm = TRUE), sum_irec = sum(irec_calibrated, na.rm = TRUE)) %>% mutate(across(where(is.numeric), ~na_if(., 0))) %>% filter(finescale_fishery_old!="NA")

pfma_yearMonth_catch_estimate_1<-pfma_yearMonth  %>% filter(sum_catch_estimate!=sum,sum_catch_estimate!=sum_historic, sum_catch_estimate!=sum_irec)

pfma_yearMonth_irec_1<-pfma_yearMonth  %>% filter(sum_catch_estimate==sum_irec)
```

```{r methods plotting PFMA - Juan de Fuca, echo=FALSE, message=FALSE, warning=FALSE, fig.height=2, fig.width=8}
#| label: fig-creel-irec-JDF
#| fig-cap: "Creel estimates in Juan de Fuca by area and month and chosen best catch estimate between creel and iREC"

g<-ggplot()+
  geom_tile(data=pfma_yearMonth %>% filter(finescale_fishery_old=="CA JDF S"), aes(x=YEAR, y=as.factor(MONTH), fill = sum_catch_estimate),colour = "white") +
  scale_fill_gradient(low = col1, high = col2,  na.value = colclear, limits=c(0,90000)) +
  scale_y_discrete(limits=rev)+
  guides(fill=guide_legend(title="creel or logbook estimates")) +
  new_scale_fill() +
  geom_tile(data=pfma_yearMonth_irec_1 %>% filter(finescale_fishery_old=="CA JDF S"), aes(x=YEAR, y=as.factor(MONTH), fill = sum_catch_estimate),colour = "white") +
  scale_fill_gradient(low = col5, high = col6,  na.value = colclear, limits=c(0,90000))+
  facet_wrap(~finescale_fishery_old+AREA, nrow=1)+
  guides(fill=guide_legend(title="iREC estimates")) +
  labs(title = "Coverage by PFMA",
       x = "Year", y = "Month") +
  theme_bw() + theme_minimal() + geom_vline(xintercept = 2013)+ theme(legend.position="bottom")

g
```

\clearpage

## Revised Chinook catch estimates from 2005-2012

Where iREC is unavailable (before 2013), we predict iREC-incorporated seasonal catch estimates by finescale fishery using a modelled relationship between summertime creel estimates without iREC and seasonal iREC-incorporated catch estimates from 2013-2023. These seasonal estimates are then rolled up into yearly estimates. This process is automated and accomplished in the model function of the PacRecCatch R package.

### Splitting fisheries by season

Following the rules for incorporating iREC, we split previous fisheries in spring (months 1:4), summer (months 5:9) and fall (10:12) seasons. A few exceptions were made, for example WCVI ISBM doesn't have a spring fishery since no ISBM fisheries occur in the spring. We split fisheries by season in order to better pool estimates since there are not always iREC or creel estimates for each month. Terminal fisheries were not split since they typically are defined by limited time period for a given stock.

### Identifying a summer creel period

We identified a fishery specific consistently creeled or logbooked coverage period for the summer period where there is consistent data throughout the time series (from 2005 to 2023) from the creel + logbook data. The summer coverage period included all months within 7:8 that are consistent through the timeseries. This consistent chunk of data will be used to predict the estimated catch in the section of the time series where we don't have iREC. For South Coast regions this data is unfiltered South Coast creel + logbook catch estimates; for North and Central coast this data is referred to as "historical" but is logbook only for Central Coast and a combination of unfiltered creel and logbook catch estimates in the North Coast.

*Note: include explanation figure here?*

### Modeling protocol

We investigated various models using a systematic approach:

-   One model for southern BC fisheries and separate models for Northern and Central BC since the data inputs are different

-   One model for spring and fall pre-terminal sport fisheries since these are iREC-only data

-   One model for summer pre-terminal sport fisheries since these are iREC and creel combined data

-   Same dataset tested for each model - therefore done without NAs

-   We didn't consider random effects models since we don't have enough levels of each factor to warrant random effects inclusion (need \>10 typically)

-   For all candidate models we used the DHARMa package to create a qq-plot to detect deviations in observations from expected distribution and a plot of the residuals against predicted values to detect patterns in residuals. We visually assessed whether candidate models improved plots.

-   We also compared candidate models using AIC, a lower AIC indicating a better model fit.

The order of comparisons was as follows:

-   First we compared three distributions: normal (gaussian), poisson, and Gamma distribution to see which fit our data the best on the full model. These distributions are all consistent with our positive, continuous data

    $$
    catch = creel.summer * status * finescale.fishery * season 
    $$

-   Next, using the chosen distribution, we compared the full model with all terms to models with each of the terms dropped sequentially except for creel.summer which was our main interest. We did this first by dropping out full terms from the full interaction model, then used the dredge function in MuMIn to sequentially drop out the various sub-interactions

-   In each case, we selected the model with the lowest AIC and best visual diagnostics.

### Model validation with Sport Head Recovery Program

We validated the modelling output with the sport head recovery program by regressing both the observed catch (2013-2023) and the modelled catch (2005-2012) with heads recovered by the the sport head recovery program which receives voluntary head submissions for recovering CWT tags.

*Note: More detail here. I.e. comparing the slopes of heads submitted with marked kept catch and the modelled catch, similar slopes = similar submission rates.*

\clearpage

## Incorporation of both sets of revised estimates in bilateral PST processes

The Chinook Technical Committee (CTC) of the Pacific Salmon Commission (PSC) has several annual analyses and reports. Notably, the Catch and Escapement Report which details catch and escapement across all fisheries. The Exploitation Rate Analysis and accompanying report which uses a cohort analysis on hatchery stocks with Coded Wire Tags (CWTs) to calculate Calendar Year Exploitation Rates on a stock-specific basis. Finally, the Calibration Report, which reportd on the annual Chinook model calibration which takes output from the ERA and other sources.

The revised Chinook catch estimates from modeled data from 2005 to 2012 and from improved data from 2013-2023 will be incorporated into a number of CTC process starting in 2025. Notably, the total number of Kept and Released Chinook will be updated in the Annual Catch and Escapement Report. Next, catch estimates are used as inputs to the annual exploitation rate analysis (ERA) in two places. First, kept and marked catch estimates are used to expand Coded Wire Tag (CWT) observations from the Sport Head Recovery Program to CWT estimates uploaded to the Regional Mark Information System ([RMIS](https://www.rmpc.org/#:~:text=This%20data%20system%2C%20named%20the%20Regional%20Mark%20Information,Commission%E2%80%99s%20information%20needs%20under%20the%20Pacific%20Salmon%20Treaty.)) and ultimately used in the ERA analysis. Second, kept and released catch estimates are incorporated for most sport fisheries into the Chinook non-retention process in the ERA.

### Updating CWT estimates

*Nick fill in here details about how marked kept catch estimates are used to create cwt estimates*

-   *caveats about if there is no data you use a regional average etc...*

### Updating Chinook non-retention (release) estimates

In 2025, in an attempt to account for mark-selective fisheries, the ERA is switching from a model that only accounts Chinook non-retention without selectively to non-retention by mark status. Therefore, the applicable rates in the ERA that are affected by the revised catch estimates are the marked released rate (MRR) and unmarked kept rate (UKR) (See more detail in the CYER Technical Report - 53). In a given PFMA and month, MRR and UKR are calculated as follows:

$Marked.release.rate = marked.released / (marked.released + marked.kept)$

$Unmarked.kept.rate = unmarked.kept/(unmarked.kept + unmarked.released)$

We calculated the MR and UK rates at the PFMA and month level and then used a weighted average based on total catch estimate per area and month to generate an MR and UK rate for each finescale fishery.

With the exception of Juan de Fuca, which did have mark-selective fisheries before 2012, all other sport fisheries were non-selective before 2013. Since the data is modelled, we did not want to bias the previous non-selective fisheries by allowing for selectivity, therefore we have fixed the MRR and URR rates to be equal in the modelled data, by calculating a single release rate for marked and unmarked fish combined. After 2013, the data is observed and therefore allowed to vary selectively even if no specific mark selective fishery was imposed. Terminal fisheries typically happen in the summer period, therefore they were assigned the same MRR and UKR rates as the equivalent summer seasonal fishery.
