---
output:
  pdf_document: default
  html_document: default
  word_document: default
---

\clearpage

# Results {#sec:results}

In the following section we present the results of revising Chinook catch using both the combined creel and iREC catch estimates for more recent years and modelled estimates for pre-iREC years back to 2005. In addition to revising catch, the relationship between marked kept Chinook catch and CWT sampling through the SHRP is presented, including sample rate time series plots for each CWT region.

## Revised catch estimates since iREC
```{r results catch1, echo=FALSE}
#### Summarizing all: 
models_combineda<-
  sport_catch_models$model_catch |>
  mutate(region_mgmt = if_else(region %in% c("SWCVI", "NWCVI", "NC"),
                               paste(region, management), 
                               region)) |>
  group_by(year, kept_status, pred_cat, region_mgmt) |> 
  summarise(across(where(is.double), \(x) sum(x, na.rm = TRUE)), .groups = "drop") |>
  mutate(fishery_full_name = recode_values(region_mgmt,
                                           "JDF" ~ "Juan de Fuca Strait",
                                           "JNST" ~ "Johnstone Strait",
                                           "NGS" ~ "Northern Strait of Georgia",
                                           "SGS" ~ "Southern Strait of Georgia",
                                           "NWCVI AABM" ~ "Northern West Vancouver Island - AABM",
                                           "SWCVI AABM" ~ "Southern West Vancouver Island - AABM",
                                           "NWCVI ISBM" ~ "Northern West Vancouver Island - ISBM",
                                           "SWCVI ISBM" ~ "Southern West Vancouver Island - ISBM", 
                                            "NC AABM" ~ "Northern BC - AABM",
                                           "NC ISBM" ~ "Northern BC - ISBM",
                                           "CBC" ~ "Central BC")) 

models_combined2a <- 
  models_combineda%>% 
  mutate(erafishery = case_when(
    grepl("JDF", region_mgmt) ~ "BC JF S",
    grepl("NGS|SGS", region_mgmt) ~ "GEO ST S",
    grepl("WCVI", region_mgmt) & grepl("AABM", region_mgmt) ~  "WCVI AABM S",
    grepl("WCVI", region_mgmt) & grepl("ISBM", region_mgmt) ~  "WCVI ISBM S",
    grepl("JNST", region_mgmt) ~  "JNST S",
     grepl("NC AABM", region_mgmt) ~  "NBC AABM S",
    grepl("NC ISBM", region_mgmt) ~  "NBC ISBM S",
    grepl("CBC", region_mgmt) ~  "CBC S",
    .default = NA_character_
  ))|>
  group_by(year, region_mgmt, erafishery, kept_status)|>
     summarize_if(is.numeric, sum, na.rm = TRUE)|>
  mutate(creel_unfiltered_plus = case_when(erafishery %in% c("NBC AABM S", "NBC ISBM S", "CBC S") ~ historic_plus, 
                                           .default= creel_unfiltered_plus))|> dplyr::select(year, region_mgmt, erafishery, kept_status, creel_unfiltered_plus, catch_estimate_predicted)


  models_combined2a_diffs<-  models_combined2a |> mutate(pdiff_irec_creel = (catch_estimate_predicted - creel_unfiltered_plus)/(creel_unfiltered_plus)) |> mutate(diff_irec_creel = (catch_estimate_predicted - creel_unfiltered_plus))

    models_combined2a_diffs<-  models_combined2a_diffs %>%
 mutate(across(where(is.numeric), ~replace(., is.infinite(.), NA)))
  
    models_combined2a_avs<-  models_combined2a_diffs |> group_by(region_mgmt, kept_status)|>
      summarize_if(is.numeric, c(mean=mean, sd=sd), na.rm=TRUE)
    
    models_combined2a_avs_pre2019<-  models_combined2a_diffs |> filter(between(year, 2013, 2018)) |> group_by(region_mgmt, kept_status)|>
      summarize_if(is.numeric, c(mean=mean, sd=sd), na.rm=TRUE)

    models_combined2a_avs_post2019<-  models_combined2a_diffs |> filter(between(year, 2019, 2023)) |> group_by(region_mgmt, kept_status)|>
      summarize_if(is.numeric, c(mean=mean, sd=sd), na.rm=TRUE)


```

In finescale fisheries, we found that the combined creel and iREC catch estimates were broadly higher than previous creel catch estimates. As an example, Figure \@ref(fig:fig-ngs-catch-revised-1) shows the effect of the summed and revised 2013 to 2023 kept catch estimates for the Northern Straight of Georgia finescale fishery (solid line) compared to the previous creel catch estimates without iREC inclusion (dashed line). Though the relative difference between the two time series vary, the revised combined creel and iREC catch estimates are consistently higher than the previous creel catch estimates. For Northern Georgia Strait, the difference between creel and combined creel and iREC is larger in the early years of the iREC program (mean `r fmt_stat(models_combined2a_avs_pre2019$pdiff_irec_creel_mean[models_combined2a_avs_pre2019$region_mgmt == "NGS" & models_combined2a_avs_pre2019$kept_status =="kept"], scale=100)`% 2013-2018, an average of `r fmt_stat(models_combined2a_avs_pre2019$diff_irec_creel_mean[models_combined2a_avs_pre2019$region_mgmt == "NGS" & models_combined2a_avs_pre2019$kept_status =="kept"])` pieces per year), and less pronounced in more recent years (mean `r fmt_stat(models_combined2a_avs_post2019$pdiff_irec_creel_mean[models_combined2a_avs_post2019$region_mgmt == "NGS" & models_combined2a_avs_post2019$kept_status =="kept"], scale=100)`% and `r fmt_stat(models_combined2a_avs_post2019$diff_irec_creel_mean[models_combined2a_avs_post2019$region_mgmt == "NGS" & models_combined2a_avs_post2019$kept_status =="kept"])` pieces on average per year 2019-2023). This pattern could be a combination of improvements in creel surveying in recent years (see methods: creel quality control) and improvements to the iREC survey which increased the number of survey responses and therefore reduced uncertainty in iREC estimates (see Appendix \@ref(app:apdx-irec-mthd) for more details). In some fisheries, the difference between combined creel and iREC and creel is consistent across all years of iREC (e.g., NWCVI AABM). 

As demonstrated by Figure \@ref(fig:fig-creel-iREC-NGS), iREC fills in spatial and temporal gaps in creel sampling, so it is expected that the summed yearly estimates are typically higher when iREC is included, especially in regions with many gaps to fill. The highest percent differences between the revised estimates (combined creel and iREC) and previous estimates occurred in Strait of Georgia Sport and Central BC Sport. The fishery with the least amount of change across the time series was Haida Gwaii Sport. As the comparison plots for the other fisheries show in Appendix \@ref(app:revised-catch-timeseries), in most cases kept catch was increased with a few exceptions where there was a small reduction in specific years. For example, in Johnstone Strait in 2019, the combined creel and iREC catch estimate was lower than the creel catch estimate (see Appendix \@ref(app:revised-catch-timeseries)). This could be due to overestimation by the unfiltered creel since 7 out of 9 estimates in this region and year failed the criteria and are therefore likely estimating catch based only on a small number of days in a month that is then expanded to the entire month. This could result in a monthly overestimation if for example a long holiday weekend is extrapolated to represent the entire monthly catch. 

Released catch estimates in finescale fisheries were also higher using combined creel and iREC compared to creel (see Appendix \@ref(app:revised-catch-timeseries)). Revised kept and released catch estimates summed to ERA fishery level are discussed in detail below.  

```{r results example summer m plots, results='asis', echo=FALSE, fig.height=3, fig.width=8, out.width="100%", fig.asp=NA}
#| label: fig-ngs-catch-revised-1
#| fig-cap: "North Strait of Georgia yearly kept catch estimates based on both previous unfiltered creel/logbook estimates and the revised estimates that include iREC"


### make this by finescale fishery old... sum up. 
ngs_graph_data <-   
  sport_catch_models$model_catch |> 
  filter(region == "NGS",
         between(year, 2013, 2023),
         kept_status == "kept") |>
  select(year, creel_unfiltered_plus, catch_estimate_predicted) |>
  pivot_longer(c(creel_unfiltered_plus, catch_estimate_predicted),
               names_to = "estimate_type",
               values_to = "catch_estimate") |>
  group_by(year, estimate_type) |>
  summarize(catch_estimate = sum(catch_estimate, na.rm = TRUE), 
            .groups = "drop")

ggplot(ngs_graph_data,
       aes(y = catch_estimate,
           x = year,
           shape = estimate_type,
           linetype = estimate_type))+
  geom_point(size = 2.5,
             colour = "#440154")+
  geom_line(colour = "#440154") +
  scale_shape_manual(values = c(16, 1),
                     labels = c("revised catch (creel + iREC)", "previous catch (creel)")
  ) +
  scale_linetype_manual(values = c(1, 2),
                        labels = c("revised catch (creel + iREC)", "previous catch (creel)")
  ) +
  scale_x_continuous(breaks = c(2013, 2015, 2017, 2019, 2021, 2023)) +
  ylim(0,60000) +
  ggtitle("Northern Strait of Georgia")  + 
  theme_bw() + 
  ylab("Kept catch estimate") + 
  xlab("Year") +
  labs(linetype = "Catch Estimate", shape = "Catch Estimate")


```

\clearpage

## Revised catch estimates before iREC

### Model fit

When revising Chinook catch estimates prior to the introduction of iREC, the following regional-temporal stratification and time series were used to produce separate models to better represent fishery and catch monitoring dynamics in different regions and seasons.

- Summer models for each of Northern West Coast Vancouver Island (WCVI), Southern WCVI, Northern Strait of Georgia, Southern Strait of Georgia, Juan de Fuca Strait, and Johnstone Strait as six separate models using 2013-2023 time period.
- Spring/Fall models for each of North and South WCVI, Northern and Southern Strait of Georgia, Juan de Fuca Strait, and Johnstone Strait as five separate models using 2013-2023 time period.
- Annual models for Central BC, Haida Gwaii (Northern BC AABM), and Northern BC ISBM, as three separate models using 2015 to 2023 time period.

The model selection process determined that separating South Coast into smaller regional models made for a better fit overall. See Appendix \@ref(app:fifth-appendix) for model selection tables comparing groupings of fisheries. These groupings make intuitive sense as creel programs in each of the regions are operated in different ways. For example, WCVI may include more extensive log book programs than other regions. Furthermore, as described above, the quality of creel surveys may vary regionally based on geography, logistics, and weather. In Central BC, Haida Gwaii, and Northern BC, there is very limited off-season fisheries and therefore a single annual model was fit for each of these regions. Furthermore, while this process allowed us to find the best fitting model among a number of candidates, residual plots show the summer models overall tend to fit better than the seasonal or annual models, and generally the summer models had excellent residual fit and show minimal patterns in the residuals (see Appendix \@ref(app:second-appendix) for residual plots).

These models develop a relationship between combined creel and iREC catch in the a given season (spring (months 1:4), summer (months 5:9), or fall (months 10:12)) and the creel catch (or logbook catch for CBC) in the summer period. This relationship is developed with each data point as a single year in the 2013-2023 time series (or 2015-2023 for NBC and CBC). The relationship is then used to predict the kept and released catch by mark status from 2005 to year before the model fit. For example the Haida Gwaii Annual model is used to predict annual estimates of kept and released catch by mark status from 2005 to 2014 based on the existing creel and logbook catch estimates.

In the following subsections we describe the chosen model and its parameters and show an example of the modelled relationships from the best fitting models for each spatio-temporal combination. The full suite of model outputs can be found in Appendix \@ref(app:third-appendix). 

```{r plotting-function, include = FALSE}
#plotting function that will be used for the next few sections
# Helper: runs predict_with_ci + plot_region_model (not by-year) for one region,
# returns a named list of ggplot objects (one per fishery). For main text use.

ModelYearEnd <- 2023
SouthModelYearStart <- 2013
NorthModelYearStart <- 2015
most_recent_data_year <- 2023
PredYearStart <- 2005


main_regions <- list(

  # --- Summer fisheries ---
  "SGS Summer"      = list(model_name = "SGS_summer_model",   x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = "summer",
                            fishery_codes = c("SGS S SUMMER")),
  "NGS Summer"      = list(model_name = "NGS_summer_model",   x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = "summer",
                            fishery_codes = c("NGS S SUMMER")),
  "JNST Summer"     = list(model_name = "jnst_summer_model",  x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = "summer",
                            fishery_codes = c("JOHN ST S SUMMER")),
  "SWCVI Summer"    = list(model_name = "SWCVI_summer_model", x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = "summer",
                            fishery_codes = c("SWCVI S SUMMER AABM", "SWCVI S SUMMER ISBM")),
  "NWCVI Summer"    = list(model_name = "NWCVI_summer_model", x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = "summer",
                            fishery_codes = c("NWCVI S SUMMER AABM", "NWCVI S SUMMER ISBM")),
  "JDF Summer"      = list(model_name = "jdf_summer_model",   x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = "summer",
                            fishery_codes = c("CA JDF S SUMMER")),

  # --- Seasonal (spring/fall) fisheries ---
  "SGS Seasonal"    = list(model_name = "seasonal_SGS_only", x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = c("spring", "fall"),
                            fishery_codes = c("SGS S SPRING", "SGS S FALL")),
  "NGS Seasonal"    = list(model_name = "seasonal_NGS_only", x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = c("spring", "fall"),
                            fishery_codes = c("NGS S SPRING", "NGS S FALL")),
  "JDF Seasonal"    = list(model_name = "seasonal_JDF",      x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = c("spring", "fall"),
                            fishery_codes = c("CA JDF S SPRING", "CA JDF S FALL")),
  "JNST Seasonal"   = list(model_name = "seasonal_JNST",     x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = c("spring", "fall"),
                            fishery_codes = c("JOHN ST S SPRING", "JOHN ST S FALL")),
  "WCVI Seasonal"   = list(model_name = "seasonal_WCVI",     x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = c("spring", "fall"),
                            fishery_codes = c("NWCVI S SPRING AABM", "SWCVI S SPRING AABM",
                                               "NWCVI S FALL AABM",   "SWCVI S FALL AABM")),

  # --- Northern BC (ISBM/AABM) fisheries ---
  "North AABM"      = list(model_name = "north_aabm",   x_var = "historic_plus",
                            x_var_mr = "mr_historic_plus", x_var_rr = "rr_historic_plus",
                            year_start = NorthModelYearStart, season_codes = "annual",
                            fishery_codes = c("NORTH AABM S")),
  "North ISBM"      = list(model_name = "north_isbm",   x_var = "historic_plus",
                            x_var_mr = "mr_historic_plus", x_var_rr = "rr_historic_plus",
                            year_start = NorthModelYearStart, season_codes = "annual",
                            fishery_codes = c("NORTH ISBM S")),
  "Central BC ISBM kept" = list(model_name = "central_isbm_kept", x_var = "historic_plus",
                            x_var_mr = "mr_historic_plus", x_var_rr = "rr_historic_plus",
                            year_start = NorthModelYearStart, season_codes = "annual",
                            fishery_codes = c("CENTRAL S")), 
   "Central BC ISBM released" = list(model_name = "central_isbm_released", x_var = "historic_plus",
                            x_var_mr = "mr_historic_plus", x_var_rr = "rr_historic_plus",
                            year_start = NorthModelYearStart, season_codes = "annual",
                            fishery_codes = c("CENTRAL S"))
)


run_main_region <- function(model_name, x_var, x_var_mr, x_var_rr, year_start,
                             season_codes, fishery_codes,
                             return_data = FALSE) {

  model <- sport_catch_models$model_list[[model_name]]

  if (is.null(model)) {
    stop("No model found for model_name = '", model_name,
         "'. Available names: ", paste(names(sport_catch_models$model_list), collapse = ", "))
  }

  df <- prepare_region_data(
    data            = sport_catch_finescale,
    combine_to      = NULL,
    fishery_codes   = fishery_codes,
    season_codes    = season_codes,
    status_codes    = c("marked_kept_total", "unmarked_kept_total",
                         "marked_released_total", "unmarked_released_total"),
    pred_year_start = PredYearStart,
    year_start      = year_start,
    year_end        = ModelYearEnd,
    x_var           = x_var,
    x_var_mr        = x_var_mr,
    x_var_rr        = x_var_rr,
    exclude_years   = NULL
  )

  pred_out <- predict_with_ci(
    model, df,
    pred_year_start = PredYearStart,
    year_start      = year_start,
    year_end        = ModelYearEnd,
    x_var           = x_var,
    x_var_mr        = x_var_mr,
    x_var_rr        = x_var_rr
  )

  pred_df        <- pred_out$predictions
  pred_df_actual <- pred_out$predictions_actual   # NEW — this is what plot_catch_heads needs
  fisheries      <- unique(df$finescale_fishery)

  if (return_data) {
    return(list(data = df, predictions = pred_df, predictions_actual = pred_df_actual))
  }

  plots <- purrr::map(
    fisheries,
    ~ plot_region_model(
      fishery    = .x,
      x_var      = x_var,
      x_var_mr   = x_var_mr,
      x_var_rr   = x_var_rr,
      plot_data  = df,
      model_data = pred_df,
      year_start = year_start,
      year_end   = ModelYearEnd
    )
  )
  names(plots) <- fisheries
  plots
}

```

### Summer models

#### Northern WCVI

The best model for Northern WCVI summer combined creel and iREC catch ($C_{creel+iREC, summer}$), was a model with a gaussian distribution that included summer creel catch ($C_{creel,summer}$), marked status ($m$) and management ($M$), an interaction between mark status and management, and an intercept. Kept status was dropped from the model during model selection, as were more complex variations of interactions.

$$C_{creel+iREC, summer} = C_{creel,summer} + M + m + M:m + 1
\label{eq:nwcvi_summer}$$

This model parameterization allows for filtered combined creel and iREC catch to be predicted independently by the creel catch, management type (e.g., AABM vs. ISBM), and mark status (e.g., marked vs. unmarked). In addition, the interaction between management and mark status allows the effect of management type on combined creel and iREC catch to be modified by mark status.

Figure \@ref(fig:fig-nwcvi-isbm-summer-model) shows ISBM modelled relationship for NWCVI Summer. The AABM counterpart plot can be found in Appendix \@ref(app:third-appendix). Overall, we found a strong linear relationship between summer combined creel and iREC catch and summer creel catch (p\<0.001, pseudo $R^{2}$ = 0.98), the intercept (but not slope) of which changed with mark status (overall there were more unmarked fish than marked fish), and the interaction between mark status and management type, such that more unmarked fish were caught during ISBM than AABM.

```{r results example seasonal summer, results='asis', echo=FALSE, fig.height=6, fig.width=8, out.width="70%", fig.asp=NA}
#| label: fig-nwcvi-isbm-summer-model
#| fig-cap: "Northern WCVI ISBM Summer fishery modelled relationship for between summertime unfiltered creel estimate (months 5 and 9) and total catch estimate that incorporates iREC. Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."

cfg <- main_regions[["NWCVI Summer"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)


  (plots[["NWCVI S SUMMER ISBM"]] + ylab("summer creel + iREC catch estimate") + xlab("summer creel catch estimate") ) 


```
\newpage
#### Southern WCVI

The best model for Southern WCVI Summer combined creel and iREC catch ($C_{creel+iREC, summer}$), was a model with a gaussian distribution that included summer creel catch ($C_{creel,summer}$), kept status ($k$) , marked status ($m$) and management ($M$), an interaction between kept status and summer creel, an interaction between mark status and summer creel, and an intercept. More complex variations and interactions were dropped during model selection.

$$C_{creel+iREC, summer} = C_{creel,summer} + M + m + k + C_{creel,summer}:k + C_{creel,summer}:m + 1
\label{eq:swcvi_summer}$$

This model parameterization allows for combined creel and iREC catch to be predicted by the creel catch, management type (e.g., AABM vs. ISBM), mark status (e.g., marked vs. unmarked), and kept status (kept vs. released). In addition, the interactions allow the effect of creel summer on catch (i.e., the slope) to be modified by kept status and mark status separately.

Figure \@ref(fig:fig-swcvi-aabm-summer-model) shows the AABM modelled relationship for SWCVI Summer. The ISBM counterpart plot can be found in Appendix \@ref(app:third-appendix). Overall, we found a strong positive linear relationship between summer combined creel and iREC catch and summer creel catch (p\<0.001, pseudo $R^{2} = 0.98$), the slope of which was modestly modified by kept status (released steeper than kept) and marked status (unmarked steeper than marked). No significant independent effects were found for kept or mark status. There was significant effect of management (AABM caught more fish overall than ISBM), and AABM appeared to have a tighter relationship than ISBM.

```{r results example seasonal summer, results='asis', echo=FALSE, fig.height=6, fig.width=8, out.width="70%", fig.asp=NA}
#| label: fig-swcvi-aabm-summer-model
#| fig-cap: "Southern WCVI AABM Summer fishery modelled relationship for between summertime unfiltered creel estimate (months 5 and 9) and total catch estimate that incorporates iREC. Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."

cfg <- main_regions[["SWCVI Summer"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)


(plots[["SWCVI S SUMMER AABM"]] + ylab("summer creel + iREC catch estimate") + xlab("summer creel catch estimate") )
  #+ facet_wrap(~management, scales="fixed")) + (plots[["SWCVI S SUMMER ISBM"]] + ylab("combined creel and iREC catch estimate") ) + facet_wrap(~management, scales="fixed")

```
\newpage
#### Northern Strait of Georgia

The best model for Northern Strait of Georgia summer combined creel and iREC catch ($C_{creel+iREC, summer}$), was a model with a gaussian distribution that included summer creel ($C_{creel,summer}$) and mark status ($m$) and an intercept. Kept status was dropped from the model during model selection, as were more complex variations and interactions.

$$C_{creel+iREC, summer} = C_{creel,summer} + m + 1
\label{eq:ngs_summer}$$

This model parameterization allows for summer combined creel and iREC catch to be predicted independently by the unfiltered creel and logbook catch and mark status (e.g., marked vs. unmarked).

Figure \@ref(fig:fig-ngs-summer-model) shows the modelled relationship for NGS Summer. Overall, we found a strong positive linear relationship between summer combined creel and iREC catch and summer creel catch (p\<0.001, pseudo $R^{2} = 0.94$). There were significantly more unmarked fish than marked fish encountered.

```{r results example seasonal summer, results='asis', echo=FALSE, fig.height=6, fig.width=8, out.width="70%", fig.asp=NA}
#| label: fig-ngs-summer-model
#| fig-cap: "Northern Strait of Georgia fishery modelled relationship between summertime unfiltered creel estimate and total catch estimate including iREC in summer months (5-9). Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."
cfg <- main_regions[["NGS Summer"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plot <- plots[[1]] + ylab("summer creel + iREC catch estimate") + xlab("summer creel catch estimate") 
plot
```

\newpage
#### Southern Strait of Georgia

The best model for Southern Strait of Georgia summer combined creel and iREC catch ($C_{creel+iREC,summer}$), was also a model with a gaussian distribution that included summer creel catch ($C_{creel,summer}$) and mark status ($m$) and an intercept. Kept status was dropped from the model during model selection, as were more complex variations and interactions.

$$C_{creel+iREC, summer} = C_{creel,summer} + m + 1
\label{eq:sgs_summer}$$

This model parameterization allows for combined creel and iREC catch to be predicted independently by the creel catch and mark status (e.g., marked vs. unmarked).

Figure \@ref(fig:fig-sgs-summer-model) shows the modelled relationship for SGS Summer. Overall, we found a moderate positive linear relationship between summer combined creel and iREC catch and summer creel catch (p\<0.001, pseudo $R^{2} = 0.71$). Overall, there were more unmarked fish than marked fish encountered.

```{r results example seasonal summer, results='asis', echo=FALSE, fig.height=6, fig.width=8, out.width="70%", fig.asp=NA}
#| label: fig-sgs-summer-model
#| fig-cap: "Southern Strait of Georgia fishery modelled relationship between summertime unfiltered creel estimate and total catch estimate including iREC in summer months (5-9). Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."
cfg <- main_regions[["SGS Summer"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plot <- plots[[1]] + ylab("summer creel + iREC catch estimate") + xlab("summer creel catch estimate") 
plot
```

\newpage
#### Johnstone Strait

The best model for Johnstone Strait (JNST) summer combined creel and iREC catch ($C_{creel+iREC,summer}$), was a model with a Gamma distribution that included summer creel ($C_{creel,summer}$), kept status ($k$), mark status ($m$), an interaction between kept status and summer creel, an interaction between mark status and summer creel, and an intercept. More complex variations and interactions were dropped.

$$C_{creel+iREC, summer} = C_{creel,summer} + m + k + C_{creel,summer}:k + C_{creel,summer}:m + 1
\label{eq:jnst_summer}$$

This model parameterization allows for combined creel and iREC catch in the whole summer period (months 5:9) to be predicted independently by the unfiltered creel and logbook catch, mark status (e.g., marked vs. unmarked), and kept status (kept vs. released). In addition, the interactions allow the effect of creel summer on combined catch (i.e., the slope) to be modified by kept status and mark status separately.

Figure \@ref(fig:fig-jnst-summer-model) shows the modelled relationship for JNST Summer. Overall, we found a strong non-linear positive relationship between summer combined creel and iREC catch and summer creel catch (p\<0.001, pseudo $R^{2} = 0.96$), the slope of which was modestly modified by kept status (released steeper than kept) and marked status (marked steeper than unmarked). There was an overall effect of mark status, as above, more unmarked fish were encountered than marked fish, and an overall effect of kept status, more fish kept than released.

```{r results example seasonal summer, results='asis', echo=FALSE, fig.height=6, fig.width=8, out.width="70%", fig.asp=NA}
#| label: fig-jnst-summer-model
#| fig-cap: "Johnstone Strait fishery modelled relationship between summertime unfiltered creel estimate and total catch estimate including iREC in summer months (5-9). Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."
cfg <- main_regions[["JNST Summer"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

 plots[[1]] + ylab("combined creel and iREC catch estimate") + xlab("summer creel catch estimate")  

```

\newpage
#### Juan de Fuca

The best model for Juan de Fuca (JDF) Strait summer combined creel and iREC catch ($C_{creel+iREC,summer}$), was a model with a Gamma distribution that included summer creel ($C_{creel,summer}$), kept_status ($k$), mark status ($m$), an interaction between kept status and summer creel, an interaction between mark status and summer creel, and an intercept. More complex variations and interactions were dropped.

$$C_{creel+iREC, summer} = C_{creel,summer} + m + k + C_{creel,summer}:k + C_{creel,summer}:m + 1
\label{eq:jdf_summer}$$

This model parameterization allows for combined creel and iREC catch in the whole summer period (months 5:9) to be predicted independently by the creel catch, mark status (e.g., marked vs. unmarked), and kept status (kept vs. released). In addition, the interactions allow the effect of summer creel on summer combined creel and iREC catch (i.e., the slope) to be modified by kept status and mark status separately.

Figure \@ref(fig:fig-jdf-summer-model) shows the modelled relationship for JDF Summer. Overall, we found a strong non-linear positive relationship between summer creel catch and summer combined creel and iREC catch (p\<0.001, pseudo $R^{2} = 0.91$), the slope of which was modestly modified by kept status (released steeper than kept) and marked status (marked steeper than unmarked). There was an overall effect of mark status, as above, more unmarked fish were encountered than marked fish, and an overall effect of kept status, more fish kept than released.

```{r results example seasonal summer, results='asis', echo=FALSE, fig.height=6, fig.width=8, out.width="70%", fig.asp=NA}
#| label: fig-jdf-summer-model
#| fig-cap: "Juan de Fuca Summer fishery modelled relationship between unfiltered creel and total catch estimates including iREC in summer months (5-9). Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."

cfg <- main_regions[["JDF Summer"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plots[[1]] + ylab("combined creel and iREC catch estimate")  + xlab("summer creel catch estimate") 


```

\newpage
### Spring and Fall Models
#### WCVI

The best model for WCVI Spring and Fall AABM iREC catch ($C_{iREC, seasonal}$), was a model with a gaussian distribution that included region ($r$), season ($s$), their interaction, and an intercept. Notably, summer creel, marked status, and kept status were dropped from the model during model selection, as were more complex variations of interactions.

$$C_{iREC, seasonal} = r + s + r:s + 1
\label{eq:wcvi_seasonal}$$

This model parameterization allows for iREC catch in the off-season to be predicted independently by region (NWCVI vs. SWCVI) and season (spring vs. fall), and for the effect of season to be modified by region.

Figure \@ref(fig:fig-swcvi-aabm-spring-model) shows the SWCVI AABM modelled relationship for Spring. The Fall and NWCVI counterpart plots can be found in Appendix \@ref(app:third-appendix). Overall, we found no relationship between summer creel and off-season iREC catch. There was a significant interaction such that SWCVI Spring had more catch than the other regions and season combinations, but no overall main effect of region or season. This flat relationship explains minimal variation in the data (pseudo $R^{2} = 0.18$), but gives an offset (intercept) to use as a predictor for off-season catch. Unlike the summer models for NWCVI and SWCVI, which both demonstrated overall more unmarked catch than marked catch, mark status was not included in the model, indicating similar catch for marked and unmarked fish in the off-season.

```{r results example seasonal summer, results='asis', echo=FALSE, fig.height=6, fig.width=8, out.width="70%", fig.asp=NA}
#| label: fig-swcvi-aabm-spring-model
#| fig-cap: "Southern WCVI AABM Summer fishery modelled relationship for between summertime unfiltered creel estimate (months 5 and 9) and total catch estimate that incorporates iREC. Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."

cfg <- main_regions[["WCVI Seasonal"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)


  (plots[["SWCVI S SPRING AABM"]] + ylab("seasonal iREC catch estimate") + xlab("summer creel catch estimate") ) 

#+ facet_wrap(~management, scales="fixed")

```
\newpage
#### Northern Strait of Georgia

The best model for Northern Strait of Georgia off-season iREC catch ($C_{iREC, seasonal}$), was a model with a gaussian distribution that included summer creel ($C_{creel,summer}$), kept status ($k$), season ($s$), an interaction between summer creel and season, an interaction between summer creel and kept status, and an intercept. Mark status and more complex variations and interactions were dropped during model selection.

$$C_{iREC, seasonal} = C_{creel,summer} + k + s + C_{creel,summer}:s + C_{creel,summer}:k + 1
\label{eq:ngs_seasonal}$$

This model parameterization allows for off-season iREC catch to be predicted independently by the summer creel catch, kept status (kept vs. released), and season (spring vs. fall). Furthermore, this model formula allows for the effect of summer creel on off-season iREC catch to be modified by season and kept status.

Figure \@ref(fig:fig-ngs-spring-model) shows the modelled relationship for NGS Spring. The Fall counterpart plot can be found in Appendix \@ref(app:third-appendix). Overall, we found a moderate positive linear relationship between summer creel and off-season iREC catch (p\<0.001, pseudo $R^{2} = 0.57$). There was an overall effect of season (more iREC catch in spring than fall), and season altered the slope of the summer creel to seasonal iREC such that the slope is steeper for spring than for fall catch. There was no overall effect of kept status, but kept status modified the slope of the creel summer catch to iREC seasonal catch: the slope is steeper for released than kept fish. Unlike the summer model for NGS, which demonstrated overall more unmarked catch than marked catch, mark status was dropped from this model, indicating similar catch for marked and unmarked fish in the off-season.

```{r results example seasonal summer, results='asis', echo=FALSE, fig.height=6, fig.width=8, out.width="70%", fig.asp=NA}
#| label: fig-ngs-spring-model
#| fig-cap: "Northern Strait of Georgia fishery modelled relationship between summertime unfiltered creel estimate and total catch estimate including iREC in spring (months 1-4). Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."
cfg <- main_regions[["NGS Seasonal"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plots[["NGS S SPRING"]] + ylab("seasonal iREC catch estimate") + xlab("summer creel catch estimate")
```
\newpage
#### Southern Strait of Georgia

The best model for Southern Strait of Georgia off-season iREC catch ($C_{iREC, seasonal}$), was a model with a gaussian distribution that included summer creel ($C_{creel,summer}$), mark status ($m$), season ($s$), an interaction between summer creel and mark status, and an intercept. Kept status and more complex variations and interactions were dropped during model selection.

$$C_{iREC, seasonal} = C_{creel,summer} + m + s + C_{creel,summer}:m + 1
\label{eq:sgs_seasonal}$$

This model parameterization allows for off-season iREC catch to be predicted independently by the unfiltered creel summer catch, mark status (e.g., marked vs. unmarked), and season (spring vs. fall). Furthermore, this model formula allows for the effect of summer creel on off-season iREC catch to be modified by mark status.

Figure \@ref(fig:fig-sgs-spring-model) shows the modelled relationship for SGS Spring. The Fall counterpart plot can be found in Appendix \@ref(app:third-appendix). Overall, we found a weak positive linear relationship between summer creel and off-season iREC catch (p\<0.011, pseudo $R^{2} = 0.39$). There was an overall effect of season (more iREC catch in spring than fall). Similar to the summer SGS model, there was an overall effect of mark status (more unmarked than marked catch). However, in this model mark status modified the slope of the creel summer catch to iREC seasonal catch: the slope is steeper for marked than unmarked fish.

```{r results example seasonal summer, results='asis', echo=FALSE, fig.height=6, fig.width=8, out.width="70%", fig.asp=NA}
#| label: fig-sgs-spring-model
#| fig-cap: "Southern Strait of Georgia fishery modelled relationship between summertime unfiltered creel estimate and total catch estimate including iREC in spring (months 1-4). Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."
cfg <- main_regions[["SGS Seasonal"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plots[["SGS S SPRING"]] + ylab("seasonal iREC catch estimate") + xlab("summer creel catch estimate")
```
\newpage
#### Johnstone Strait

The best model for Johnstone Strait seasonal iREC catch ($C_{iREC, seasonal}$), was a model with a gaussian distribution that included mark status ($m$) and an intercept. Notably, summer creel, kept status, season, and any more complex variations and interactions were dropped.

$$C_{iREC, seasonal} = m + 1
\label{eq:jnst_seasonal}$$

This model parameterization allows for iREC catch in the off-season to be predicted independently by mark status (e.g., marked vs. unmarked).

Figure \@ref(fig:fig-jnst-spring-model) shows the modelled relationship for JNST Spring. The Fall counterpart plot can be found in Appendix \@ref(app:third-appendix). Overall, we found no relationship between summer creel and off-season iREC catch. As in the summer model for JNST, there was a significant effect of mark status, more unmarked fish caught than marked fish. The flat relationship explains minimal variation in the data (pseudo $R^{2} = 0.08$), but gives an offset (intercept) to use as a predictor for off-season iREC catch.

```{r results example seasonal summer, results='asis', echo=FALSE, fig.height=6, fig.width=8, out.width="70%", fig.asp=NA}
#| label: fig-jnst-spring-model
#| fig-cap: "Johnstone Strait fishery modelled relationship between summertime unfiltered creel estimate and total catch estimate including iREC in spring months (1:4). Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."
cfg <- main_regions[["JNST Seasonal"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

 plots[["JOHN ST S SPRING"]] + ylab("seasonal iREC catch estimate") + xlab("summer creel catch estimate")

```
\newpage
#### Juan de Fuca

The best model for Juan de Fuca Strait off-season iREC catch ($C_{iREC, seasonal}$), was a model with a Gamma distribution that included summer creel ($C_{creel,summer}$), kept status ($k$), mark status ($m$), season($s$), an interaction between mark status and summer creel, and an intercept. More complex variations and interactions were dropped.

$$C_{iREC, seasonal} = C_{creel,summer} + m + k + s + C_{creel,summer}:m + 1
\label{eq:jdf_seasonal}$$

This model parameterization allows for iREC catch in the off-season to be predicted independently by the summertime creel catch, mark status (e.g., marked vs. unmarked), kept status (kept vs. released), and season. In addition, the interaction allows the effect of creel summer on off-season iREC catch (i.e., the slope) to be modified by mark status.

Figure \@ref(fig:fig-jdf-spring-model) shows the modelled relationship for JDF Spring. The Fall counterpart plot can be found in Appendix \@ref(app:third-appendix). Overall, we found a modest non-linear positive relationship between summer creel and seasonal iREC catch (p\<0.003, pseudo $R^{2} = 0.43$), the slope of which was modified by marked status (marked steeper than unmarked), similar to the summer JDF model. Although there was no overall effect of mark status, the trend was in the opposite direction to the summer JDF model, slightly more marked fish than unmarked fish caught in the off-season (p=0.08). There was an overall effect of kept status, slightly more fish released than kept (p\<0.003), the opposite direction to the summer JDF model. Finally, there was an overall effect of season, where more fish were caught in the spring than the fall.

```{r results example seasonal summer, results='asis', echo=FALSE, fig.height=6, fig.width=8, out.width="70%", fig.asp=NA}
#| label: fig-jdf-spring-model
#| fig-cap: "Juan de Fuca Spring fishery modelled relationship between unfiltered creel and total catch estimates including iREC in spring months (1-4). Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."

cfg <- main_regions[["JDF Seasonal"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plots[["CA JDF S SPRING"]] + ylab("seasonal iREC catch estimate") + xlab("summer creel catch estimate") 


```
\newpage
### Annual Fisheries models
#### Haida Gwaii
The best model for the Haida Gwaii (Northern BC Sport AABM) annual iREC catch incorporated the annual creel catch $C_{creel, annual}$), kept status ($k$), mark status ($m$), and the interaction between kept and mark status, and an intercept, as identified in Equation \eqref{eq:haida_gwaii_annual}. We used a gaussian link instead of a Gamma log link because earlier values in the time series were outside the range of the model fit (see details in Appendix \@ref(apdx-haida-model)).

$$C_{iREC, annual} = C_{creel, annual} + k + m + k:m + 1 
\label{eq:haida_gwaii_annual}$$

This model parameterization allows for iREC catch to be predicted by annual creel catch, kept status and mark status. The interaction between kept and mark status allows the effect of kept status to be moderated by mark status. 

Figure \@ref(fig:fig-nbcaabm-model) shows the model results with Haida Gwaii. Overall, we found a strong positive linear relationship between annual creel and annual iREC catch (p\<0.001, pseudo $R^{2} = 0.91$). There were no overall significant effects of either mark or kept status, but a trending effect of mark status (p=0.056), which suggested that there were more unmarked fish encountered than marked fish. Furthermore, the interaction suggests that for unmarked fish, more fish were released than kept, while there were similar numbers of kept and released fish in the marked category.

```{r results NBC AABM Annual, results='asis', echo=FALSE, fig.height=6, fig.width=8, out.width="70%", fig.asp=NA}
#| label: fig-nbcaabm-model
#| fig-cap: "Haida Gwaii (Northern BC Sport AABM) annual fishery modelled relationship between historic unfiltered creel estimate and total iREC catch estimates in the whole year. Data points are yearly estimates from 2015 to 2023. Note varying scales on the y-axes to better visualize the model fit."

cfg <- main_regions[["North AABM"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plots[[1]] + ylab("annual iREC catch estimate") + xlab("annual creel catch estimate")

```
\newpage
#### Northern BC ISBM Fisheries

The best model for annual Northern BC (NBC) ISBM iREC catch was one with a Gamma distribution using a log link that included annual creel catch, $C_{creel,annual}$), mark status ($m$), and an interaction between historic data and mark status. Kept status and more complex variations and interactions were dropped.

$$C_{iREC, annual} = C_{creel,annual} + m + C_{creel,annual}:m \label{eq:nbc_isbm_annual}$$

This model parameterization allows for annual iREC catch to be predicted both independently by the annual creel catch, by mark status, and by their interaction.

Figure \@ref(fig:fig-nbcisbm-model) shows the annual model results in NBC ISBM. There is a positive non-linear relationship between creel catch and iREC annual catch estimates (p\<0.0001, pseudo $R^{2} = 0.91$). The slope of this relationship was slightly steeper for marked fish than unmarked fish. There was a significant effect of mark status, such that overall the marked catch was substantially less (\< 1,500 pieces) than the catch of unmarked fish (\< 15,000 pieces).

```{r results NBC ISBM annual, results='asis', echo=FALSE, fig.height=6, fig.width=8, out.width="70%", fig.asp=NA}
#| label: fig-nbcisbm-model
#| fig-cap: "Northern BC Sport ISBM fishery modelled relationship between unfiltered creel estimate and iREC catch estimates over the whole year. Data points are yearly regional estimates from 2015 to 2023. Note varying scales on the y-axes to better visualize the model fit."


cfg <- main_regions[["North ISBM"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plots[[1]] +  ylab("annual iREC catch estimate") + xlab("annual creel catch estimate")
#+ facet_wrap(~mark_status, scales="fixed")



```
\newpage
#### Central BC annual fishery

Central BC was split into two models, one for kept fish and one for released fish where there were almost exclusively zeros provided in the historical data. The best model for the kept Central BC annual catch was one with a Gamma distribution with a log link which included logbook catch and mark status. The best model for the released Central BC annual catch was one with a Gamma distribution with a log link which included mark status and an intercept. As historic logbook estimates were mostly zero, this factor was excluded during model selection, as was kept status and various interactions.

$$ C_{iREC,annual,kept} = C_{logbook,annual} + m \label{eq:cbc_annual_kept}$$
$$ C_{iREC,annual,released} = m + 1 \label{eq:cbc_annual_released}$$

This model parameterization allows for kept iREC catch to be predicted independently by the annual logbook catch and by mark status and released iREC catch to be predicted independently by mark status.

Figure \@ref(fig:fig-cbc-model) shows the combined CBC model results. Generally for low catch estimates a flat line is predicted, but for unmarked Kept fish, the iREC estimate is an increasing function of summertime logbook catch. In both models there was an effect of mark status such that there were more unmarked fish encountered than marked fish. There was an overall positive non-linear relationship between kept logbook catch and iREC (p\<0.01, pseudo $R^{2} = 0.90$). As released catch is not captured by the log book program, the predicted release data before iREC is a mean of the released catch in 2013-2023. 

```{r results CBC model, results='asis', echo=FALSE, fig.height=6, fig.width=8, out.width="70%", fig.asp=NA}
#| label: fig-cbc-model
#| fig-cap: "CBC annual fishery modelled relationship between annual logbook estimates and annual iREC catch estimates. Data points are yearly estimates from 2015 to 2023. Note varying scales on the y-axes to better visualize the model fit."

cfg <- main_regions[["Central BC ISBM kept"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plot <- plots[[1]] + ylab("annual iREC catch estimate") + xlab("annual logbook catch estimate")
plot

```

\newpage
## Validation using sport head recoveries

```{r heads-to-creel plotting function, echo=FALSE}
#| label: fig-heads-to-creel-and-updated-catch-2
#| fig-cap: "Relationship between the marked kept catch estimates from the the creel program (left panel) and the combined creel and iREC catch estimates (right panel) compared to the Chinook head recoveries provided to the SHRP."
run_catch_heads_group <- function(region_names, x_var_original, x_var_mr,
                                   region_list = main_regions, start_y=2005, end_y=2023) {

  raw <- purrr::map(region_names, function(rn) {
    cfg <- region_list[[rn]]

    run_main_region(
      model_name    = cfg$model_name,
      x_var         = cfg$x_var,
      x_var_mr      = cfg$x_var_mr,
      x_var_rr      = cfg$x_var_rr,
      year_start    = cfg$year_start,
      season_codes  = cfg$season_codes,
      fishery_codes = cfg$fishery_codes,
      return_data   = TRUE
    )
  })

  df_group        <- purrr::map_dfr(raw, "data") 
  pred_group_actual <- purrr::map_dfr(raw, "predictions_actual")

  plot_catch_heads(
    x_var_original   = x_var_original,
    x_var_mr         = x_var_mr,
    plot_data        = df_group,
    pred_data_actual = pred_group_actual,
    pred_year_start  = PredYearStart,
    year_start       = min(purrr::map_dbl(region_names, ~ region_list[[.x]]$year_start)),
    year_start_plot = start_y,
    year_end_plot   = end_y
  )
}
```

```{r heads-to-creel-results, echo=FALSE}

summer_regions <- c("SGS Summer","NGS Summer","JNST Summer","SWCVI Summer","NWCVI Summer","JDF Summer")
out_summer <- run_catch_heads_group(summer_regions,
                              x_var_original = "creel_unfiltered_plus",
                              x_var_mr       = "mr_creel_plus_summer", start_y = 2013, end_y=2023)

seasonal_regions <- c("SGS Seasonal","NGS Seasonal","JDF Seasonal","JNST Seasonal","WCVI Seasonal")
out_seasonal <- run_catch_heads_group(seasonal_regions,
                              x_var_original = "creel_unfiltered_plus",
                              x_var_mr       = "mr_creel_plus_summer", start_y = 2013, end_y=2023)


northern_aabm <- c("North AABM")
out_north_aabm <- run_catch_heads_group(northern_aabm,
                              x_var_original = "historic_plus",
                              x_var_mr       = "mr_historic_plus", start_y = 2015, end_y=2023)

northern_isbm <- c("North ISBM")
out_nbc_isbm <- run_catch_heads_group(northern_isbm,
                              x_var_original = "historic_plus",
                              x_var_mr       = "mr_historic_plus", start_y = 2015, end_y=2023)

cbc_region <- c("Central BC ISBM kept")
out_cbc <- run_catch_heads_group(cbc_region,
                              x_var_original = "historic_plus",
                              x_var_mr       = "mr_historic_plus", start_y = 2015, end_y=2023)
```

As the total number of heads collected by the SHRP is expected to be primarily driven by the the total kept marked catch in recreational fisheries, the total kept marked catch and total heads collected by the SHRP should be correlated. The strength of the correlation (i.e., $R^{2}$ value) along with the submission rate (i.e., slope) can be used as an indicator of the accuracy of the catch estimate. In contrast to the creel program, the SHRP relies on several sampling approaches (e.g., lodge sampling, public collection points, avid anglers) and should not necessarily be driven by creel survey coverage nor iREC survey completion. To better align with the CWT estimation methods, the catch and heads were spatially grouped by the region defined in Figure \@ref(fig:fig-CwtMap) and the seasonal period. Here we use the sport head recoveries to compare the marked kept catch in the previously used estimates (creel only) and revised estimates (combined creel and iREC observed from 2013-2023 and modelled from 2005-2012). To align with the modelling stratification, we examine these relationships in both the summer season and the off-season in Southern fishery regions and annually in Northern fishery regions.

### Revised catch estimates since iREC

First, in Figure \@ref(fig:fig-heads-to-creel-and-updated-catch-summer) we compare the observed summertime relationship between marked creel catch to recovered heads (left panel) to that of marked combined creel and iREC catch to recovered heads (right panel) using data from 2013 to 2023 and combining regions from the South Coast of BC. As described above, in the summer months we prioritized the creel estimate and the iREC estimate was used to fill in months and areas where a high-quality creel estimate was unavailable. Therefore, while the number of points remains the same (one for each region and year), the underlying number of areas and months contributing catch data to each point is higher for the combined creel and iREC (see Figure \@ref(fig:fig-creel-iREC-NGS) for a visualization of the increased spatio-temporal coverage of combined creel and iREC). Across the South Coast region, creel and combined creel and iREC perform relatively equally, and are able to explain `r fmt_stat(out_summer$r2_creel, scale=100)` and `r fmt_stat(out_summer$r2_irec, scale=100)`% respectively of the variation, a small improvement using combined creel and iREC. These regressions also have similar slopes, indicating an average sample rate across South Coast regions of `r fmt_stat(out_summer$slope_creel, scale=100)`% for creel and `r fmt_stat(out_summer$slope_irec, scale=100)`% for combined creel and iREC. While there is some variation among finescale fishery (e.g., NWCVI and NGS have steeper slopes, higher sample rates due to high lodge participation), the slope doesn't change substantially with the addition of iREC information and the $R^{2}$ improves slightly.

Next, Figure \@ref(fig:fig-heads-to-creel-and-updated-catch-seasonal) shows the off-season relationship between marked creel catch to recovered heads and marked iREC catch to recovered heads. In the off-season (spring and fall), the fit is substantially improved: from $R^{2}=$ `r fmt_stat(out_seasonal$r2_creel)` using creel to $R^{2}=$ `r fmt_stat(out_seasonal$r2_irec)` using iREC. This is likely because as iREC covers more months and areas, it allows more head recoveries to be used with the appropriate regional and seasonal catch (n=`r fmt_stat(out_seasonal$n_irec)` vs. n=`r fmt_stat(out_seasonal$n_creel)`). Previously these heads were not assigned to a seasonal catch and were given an estimated rather than direct submission rate. In addition, using creel would indicate a sample rate of `r fmt_stat(out_seasonal$slope_creel, scale=100)`%, which, given that the fit is so poor, is likely biased high. iREC has a more realistic sample rate of `r fmt_stat(out_seasonal$slope_irec, scale=100)`% on average. The low rate is likely due to low lodge use in the off-season and a new strategy may need to be employed to increase the head submission rate in the off-season. Sample rates by region and season over time can be found in Figure \@ref(fig:fig-fishery-rates-2) and are discussed below. 

For the South Coast, both the summer and seasonal sport head recovery analysis supports the concern that creel catch alone provides an under estimate of catch that could impact assessment of CWT indicators. These plots validate the use of iREC as the best catch estimate in the off-season and support the combination of creel with iREC during the summertime.

```{r heads-to-creel-summer, echo=FALSE, fig.height=4, fig.width=10, out.width="100%", fig.asp=NA}
#| label: fig-heads-to-creel-and-updated-catch-summer
#| fig.cap: "Relationship between the summer marked kept catch estimates from the the creel program (left panel) and the combined creel and iREC catch estimates (right panel) compared to the Chinook head recoveries provided to the SHRP from 2013-2023."

out_summer$plot[[1]] <- out_summer$plot[[1]] + labs(title = "previous catch (creel)")
out_summer$plot[[2]] <- out_summer$plot[[2]] + labs(title = "revised catch (creel + iREC)")
out_summer$plot
```

```{r heads-to-creel-seasonal, echo=FALSE, fig.height=4, fig.width=10, out.width="100%", fig.asp=NA}
#| label: fig-heads-to-creel-and-updated-catch-seasonal
#| fig-cap: "Relationship between the seasonal (spring/fall) marked kept catch estimates from the creel program (left panel) and the iREC catch estimates (right panel) compared to the Chinook head recoveries provided to the SHRP"

out_seasonal$plot[[1]] <- out_seasonal$plot[[1]] + labs(title = "previous catch (creel)")
out_seasonal$plot[[2]] <- out_seasonal$plot[[2]] + labs(title = "revised catch (iREC)")
out_seasonal$plot
```
\newpage
For Haida Gwaii (Northern AABM Sport), between 2015 and 2023, there was an excellent submission rate of `r fmt_stat(out_north_aabm$slope_creel, scale=100)`% using creel methods and this slightly increased with iREC to `r fmt_stat(out_north_aabm$slope_irec, scale=100)`%. Both methods have excellent fit, $R^{2}=$ `r fmt_stat(out_north_aabm$r2_creel)` for creel and $R^{2}=$ `r fmt_stat(out_north_aabm$r2_irec)` for iREC. This indicates that creel and iREC in this region are both independently creating reliable estimates of marked kept catch. It further may suggest that Haida creel can be used in place of iREC-only methods for marked kept catch, although released catch tends to be more fully captured by iREC.

```{r heads-to-creel-northern-aabm, echo=FALSE, fig.height=4, fig.width=10, out.width="100%", fig.asp=NA}
#| label: fig-heads-to-creel-and-updated-catch-northern-aabm
#| fig-cap: "Relationship between the Northern BC AABM marked kept catch estimates from the creel program (left panel) and the combined creel and iREC catch estimates (right panel) compared to the Chinook head recoveries provided to the SHRP"

out_north_aabm$plot+ labs(title="iREC")
```
In Northern ISBM, iREC improves the fit of the heads to catch relationship ($R^{2}=$ `r fmt_stat(out_nbc_isbm$r2_creel)` for creel vs. $R^{2}=$ `r fmt_stat(out_nbc_isbm$r2_irec)` for iREC), and the sample rate drops slightly using iREC to `r fmt_stat(out_nbc_isbm$slope_irec, scale=100)`% from `r fmt_stat(out_nbc_isbm$slope_creel, scale=100)`% using creel.
```{r heads-to-creel-northern-isbm, echo=FALSE, fig.height=4, fig.width=10, out.width="100%", fig.asp=NA}
#| label: fig-heads-to-creel-and-updated-catch-northern-isbm
#| fig-cap: "Relationship between the Northern BC ISBM marked kept catch estimates from the creel program (left panel) and the combined creel and iREC catch estimates (right panel) compared to the Chinook head recoveries provided to the SHRP"

out_nbc_isbm$plot[[1]] <- out_nbc_isbm$plot[[1]] + labs(title = "previous catch (creel)")
out_nbc_isbm$plot[[2]] <- out_nbc_isbm$plot[[2]] + labs(title = "revised catch (iREC)")
out_nbc_isbm$plot
```

In Central BC, the logbook program was underestimating catch from 2013-2023, as the sample rate is `r fmt_stat(out_cbc$slope_creel, scale=100)`% and when we use iREC we get a more reasonable sample rate of `r fmt_stat(out_cbc$slope_irec, scale=100)`%. The fit is very similar at $R^{2}=$ `r fmt_stat(out_cbc$r2_creel)` for creel and $R^{2}=$ `r fmt_stat(out_cbc$r2_irec)` for iREC.

```{r heads-to-creel-cbc, echo=FALSE, fig.height=4, fig.width=10, out.width="100%", fig.asp=NA}
#| label: fig-heads-to-creel-and-updated-catch-central
#| fig-cap: "Relationship between Central BC marked kept catch estimates from the logbook program (left panel) and the  iREC catch estimates (right panel) compared to the Chinook head recoveries provided to the SHRP"

out_cbc$plot[[1]] <- out_cbc$plot[[1]] + labs(title = "previous catch (logbook)")
out_cbc$plot[[2]] <- out_cbc$plot[[2]] + labs(title = "revised catch (iREC)")
out_cbc$plot
```



### Revised catch estimates before iREC

```{r heads-to-creel-results-2005, echo=FALSE}

out_summer05 <- run_catch_heads_group(summer_regions,
                              x_var_original = "creel_unfiltered_plus",
                              x_var_mr       = "mr_creel_plus_summer", start_y = 2005, end_y=2012)

out_seasonal05 <- run_catch_heads_group(seasonal_regions,
                              x_var_original = "creel_unfiltered_plus",
                              x_var_mr       = "mr_creel_plus_summer", start_y = 2005, end_y=2012)

out_north_aabm05 <- run_catch_heads_group(northern_aabm,
                              x_var_original = "historic_plus",
                              x_var_mr       = "mr_historic_plus", start_y = 2005, end_y=2015)

out_nbc_isbm05 <- run_catch_heads_group(northern_isbm,
                              x_var_original = "historic_plus",
                              x_var_mr       = "mr_historic_plus", start_y = 2005, end_y=2015)

out_cbc05 <- run_catch_heads_group(cbc_region,
                              x_var_original = "historic_plus",
                              x_var_mr       = "mr_historic_plus", start_y = 2005, end_y=2015)
```
We first compare the observed summertime relationship between marked creel catch to recovered heads and that of modelled marked combined creel and iREC catch to recovered heads using data from 2005 to 2012 and regions from the South Coast of BC (Figure \@ref(fig:fig-heads-to-creel-and-updated-catch-summer-2005)). The modelled estimate (right panel) explains `r fmt_stat(out_summer05$r2_irec, scale=100)`% of variation, a good fit and similar to the original creel estimate performance (`r fmt_stat(out_summer05$r2_creel, scale=100)`% left panel). These regressions also have similar slopes, indicating an average sample rate across South Coast regions of `r fmt_stat(out_summer05$slope_creel, scale=100)`% for creel and `r fmt_stat(out_summer05$slope_irec, scale=100)`% for modelled combined creel and iREC. In both cases, the sample rates were higher in the 2005-2012 period than the 2013-2023 period, this is discussed further below.

Next, Figure \@ref(fig:fig-heads-to-creel-and-updated-catch-seasonal-2005) shows the off-season relationship between marked creel catch to recovered heads and modelled iREC only marked catch to recovered heads. In the off-season, the fit is improved: from $R^{2}=$ `r fmt_stat(out_seasonal05$r2_creel)` using creel to $R^{2}=$ `r fmt_stat(out_seasonal05$r2_irec)` using modelled iREC. This is likely because iREC covers more months and areas, it allows more head recoveries to be used in more times are areas (n=`r fmt_stat(out_seasonal05$n_irec)`, vs. n=`r fmt_stat(out_seasonal05$n_creel)`). This validates our use of modelled iREC catch in the off-season. In addition, using creel would indicate a sample rate of `r fmt_stat(out_seasonal05$slope_creel, scale=100)`%, which is much more likely than the 2013-2023 estimate of 2019 but still perhaps a bit high, while the modelled iREC catch indicates a sample rate of `r fmt_stat(out_seasonal05$slope_irec, scale=100)`% ($R^{2}=$ `r fmt_stat(out_seasonal$r2_irec)`), similar to the 2013-2023 period.

```{r heads-to-creel-summer-2005-2012, echo=FALSE, fig.height=4, fig.width=10, out.width="100%", fig.asp=NA}
#| label: fig-heads-to-creel-and-updated-catch-summer-2005
#| fig-cap: "Relationship between the summer marked kept catch estimates from the the creel program (left panel) and the modelled combined creel and iREC catch estimates (right panel) compared to the Chinook head recoveries provided to the SHRP in 2005-2012"

out_summer05$plot[[1]] <- out_summer05$plot[[1]] + labs(title = "previous catch (creel)")
out_summer05$plot[[2]] <- out_summer05$plot[[2]] + labs(title = "revised catch (modelled combined creel and iREC)")
out_summer05$plot


```

```{r heads-to-creel-summer-2005-2012, echo=FALSE, fig.height=4, fig.width=10, out.width="100%", fig.asp=NA}
#| label: fig-heads-to-creel-and-updated-catch-seasonal-2005
#| fig-cap: "Relationship between the seasonal marked kept catch estimates from the the creel program (left panel) and the modelled iREC catch estimates (right panel) compared to the Chinook head recoveries provided to the SHRP in 2005-2012"

out_seasonal05$plot[[1]] <- out_seasonal05$plot[[1]] + labs(title = "previous catch (creel)")
out_seasonal05$plot[[2]] <- out_seasonal05$plot[[2]] + labs(title = "revised catch (modelled iREC)")
out_seasonal05$plot
```

Figure \@ref(fig:fig-heads-to-creel-and-updated-catch-nbc-aabm-2005) shows the relationships for the three Northern regions. Again, NBC AABM has a high sample rate (`r fmt_stat(out_north_aabm05$slope_creel, scale=100)`%) and very good fit ($R^{2}=$ `r fmt_stat(out_north_aabm05$r2_creel)`) for the creel program and only slight derivation from this when using modelled iREC data from 2005-2012 (sample rate = `r fmt_stat(out_north_aabm05$slope_irec, scale=100)`%, $R^{2}=$ `r fmt_stat(out_north_aabm05$r2_irec)`). NBC ISBM creel catch estimates in 2005-2012 have a poor fit to the head recoveries ($R^{2}=$ `r fmt_stat(out_nbc_isbm05$r2_creel)`%) which is slightly improved by using modelled iREC data ($R^{2}=$ `r fmt_stat(out_nbc_isbm05$r2_irec)`). The sample rates are also lower but similar to the 2013-2023 period, for creel `r fmt_stat(out_nbc_isbm05$slope_creel, scale=100)`% and modelled iREC `r fmt_stat(out_nbc_isbm05$slope_irec, scale=100)`%. CBC ISBM again shows a biased sample rate of `r fmt_stat(out_cbc05$slope_creel, scale=100)`% which is improved to a more realistic `r fmt_stat(out_cbc05$slope_irec, scale=100)`% when modelled iREC is considered. And although the modelled iREC has many static values in this period, the fit is also improved from $R^{2}=$ `r fmt_stat(out_cbc05$r2_creel)` to $R^{2}=$ `r fmt_stat(out_cbc05$r2_irec)`.

```{r heads-to-creel-nbc-aabm-2005, echo=FALSE, fig.height=10, fig.width=10, out.width="100%", fig.asp=NA}
#| label: fig-heads-to-creel-and-updated-catch-nbc-aabm-2005
#| fig-cap: "Relationship between marked kept catch from the creel program (left panel) and the combined creel and iREC catch (right panel) compared to the Chinook head recoveries provided to the SHRP from 2005 to 2015 for Haida Gwaii (top row), Northern BC ISBM (middle row) and Central BC (bottom row)."

out_north_aabm05$plot[[1]] <- 
  out_north_aabm05$plot[[1]] + 
  labs(title = "Haida Gwaii Previous Catch (creel)") + 
  theme(legend.position = "none")

out_north_aabm05$plot[[2]] <- 
  out_north_aabm05$plot[[2]] + 
  labs(title = "Haida Gwaii Revised Catch (modelled iREC)") + 
  theme(legend.position = "none")

out_north_aabm05$plot <-
  out_north_aabm05$plot[[1]] + 
  out_north_aabm05$plot[[2]]


out_nbc_isbm05$plot[[1]] <- 
  out_nbc_isbm05$plot[[1]] + 
  labs(title = "NBC ISBM Previous Catch (creel)") + 
  theme(legend.position = "none")

out_nbc_isbm05$plot[[2]] <- 
  out_nbc_isbm05$plot[[2]] + 
  labs(title = "NBC ISBM Revised Catch (modelled iREC)") + 
  theme(legend.position = "none")
#out_nbc_isbm05$plot


out_cbc05$plot[[1]] <- 
  out_cbc05$plot[[1]] + 
  labs(title = "Central BC Previous Catch (logbook)") + 
  theme(legend.position = "none")

out_cbc05$plot[[2]] <- 
  out_cbc05$plot[[2]] + 
  labs(title = "Central BC Revised Catch (modelled iREC)") + 
  theme(legend.position = "none")
#out_cbc05$plot

out_all <- 
  out_north_aabm05$plot[[1]] + 
  out_north_aabm05$plot[[2]] + 
  out_nbc_isbm05$plot[[1]] + 
  out_nbc_isbm05$plot[[2]] + 
  out_cbc05$plot[[1]] + 
  out_cbc05$plot[[2]] + 
  plot_layout(guides="keep", ncol=2, axes="collect")

out_all

```

Overall, we demonstrate that incorporating iREC both in current years and projecting into the past is as good a predictor of recovered heads or better than creel in most places.

\newpage
### Submission rates over time
```{r shrp-head-catch-model-fits, echo=FALSE}

lm_eqn <- function(df){
  m <- lm(head_total ~ catch_estimate_predicted + 0, df)
  
  eq <- substitute(italic(y) == a %.% italic(x)*", Adj."~~italic(R)^2~"="~r2*"," ~~italic(n)*"="~df_n,
                   list(a = format(unname(coef(m)[1]), digits = 3),
                        b = format(unname(coef(m)[2]), digits = 3),
                        r2 = format(summary(m)$adj.r.squared, digits = 3),
                        df_n = nrow(df)))
  as.character(as.expression(eq))
}

lm_eqn_vec <- function(head_total, catch_estimate){
  m <- lm(head_total ~ catch_estimate + 0)
  
  eq <- substitute(italic(y) == a %.% italic(x)*", Adj."~~italic(R)^2~"="~r2*"," ~~italic(n)*"="~df_n,
                   list(a = format(unname(coef(m)[1]), digits = 3),
                        b = format(unname(coef(m)[2]), digits = 3),
                        r2 = format(summary(m)$adj.r.squared, digits = 3),
                        df_n = length(head_total)))
  as.character(as.expression(eq))
}

lm_log_eqn_vec <- function(head_total, catch_estimate){
  m <- lm(log(head_total) ~ log(catch_estimate) + 0)
  
  eq <- substitute(italic(ln(y)) == a %.% italic(ln(x))*", Adj."~~italic(R)^2~"="~r2*"," ~~italic(n)*"="~df_n,
                   list(a = format(unname(coef(m)[1]), digits = 3),
                        b = format(unname(coef(m)[2]), digits = 3),
                        r2 = format(summary(m)$adj.r.squared, digits = 3),
                        df_n = length(head_total)))
  as.character(as.expression(eq))
}

creel_to_heads <-
  sport_catch_models$model_catch |>
  filter(status == "marked_kept_total") |> 
  full_join(pacRecCatch::get_head_totals(), 
            by = c("year", finescale_fishery = "region")) |>
  mutate(finescale_fishery = if_else(grepl("NWCVI|SWCVI", finescale_fishery), 
                                     trimws(gsub(" ISBM| AABM", "", finescale_fishery)),
                                     finescale_fishery))  |>
  group_by(year, finescale_fishery) |>
  select(year, finescale_fishery, creel_unfiltered_plus, head_total, catch_estimate_predicted, catch_estimate) |>
  summarize(across(c(head_total, catch_estimate_predicted, creel_unfiltered_plus, catch_estimate),
                   \(x) sum(x, na.rm=TRUE)),
            .groups = "drop") |>
  filter(head_total > 0, head_total < 6000)


south_creel_to_heads <- 
  creel_to_heads |>
  filter(!grepl("^NORTH ", finescale_fishery), !grepl("^CENTRAL " ,finescale_fishery))

all_heads_lm <- lm(head_total ~ catch_estimate_predicted + 0, creel_to_heads)


pre_iREC_catch_model_df <- 
  creel_to_heads |>
  filter(year < 2013)

pre_iREC_lm <- lm(head_total ~ catch_estimate_predicted + 0, pre_iREC_catch_model_df)


post_iREC_catch_model_df <- 
  creel_to_heads |>
  filter(year >= 2013)

post_iREC_lm <- lm(head_total ~ catch_estimate_predicted + 0, post_iREC_catch_model_df)


south_creel_to_heads_df <-
  south_creel_to_heads |>
  filter(creel_unfiltered_plus > 0)

south_creel_to_heads_lm <- lm(head_total ~ creel_unfiltered_plus + 0, south_creel_to_heads_df)

south_combine_to_heads_df <-
  south_creel_to_heads |>
  filter(catch_estimate_predicted > 0)

combine_creel_to_heads_lm <- lm(head_total ~ catch_estimate_predicted + 0, south_combine_to_heads_df)
```

The SHRP has generally run consistently over the transition time between the modelled catch (pre-2013) and subsequent catch period during which iREC has been incorporated into estimates. So, the linear relationship between marked kept catch and head recoveries from both the 2005 - 2012 modelled catch time series and the 2013 - 2023 combined creel and iREC catch time series should be similar. We explore these trends further in this section. 

In Figure \@ref(fig:fig-fishery-rates-2) we visualize updated submission rates to the SHRP across time in each region. A submission rate of 20% of the catch is the desired minimum sampling rate, and some regions are meeting this target while others are not. Submission rates fluctuate year to year and between seasons. Above, for the South coast and North AABM (Haida Gwaii) we noted a decline in submission rates from the 2005-2012 period to the 2013-2023 period, reflected both in creel and combined creel and iREC data, but the opposite trend in NBC ISBM and CBC. Overall, while there are some trends over time in submission rates, there is no clear and consistent step change occurring in 2013 (or 2015 for Northern regions) that would indicate that pre- and post-iREC are separate catch regimes. Furthermore, when combining all regions and comparing the model fit before iREC (Figure \@ref(fig:fig-heads-kept-iREC) left panel) and with iREC (Figure \@ref(fig:fig-heads-kept-iREC) right panel), CWT sample rates are fairly consistent with a before iREC average sample rate (i.e., slope) of `r format(unname(coef(pre_iREC_lm)[1]), digits = 3)` and an iREC period sample rate of `r format(unname(coef(post_iREC_lm)[1]), digits = 3)` with similar adjusted $R^{2}$ values. This suggests that there is no consistent bias in the modelled vs. observed combined creel and iREC estimates. 

The summer submission rates in Northern WCVI, Johnstone Strait, and the annual Haida Gwaii rates have typically been over 20% for most of the time series. This is likely due to high lodge participation in the sport head recovery program. Haida Gwaii Sport in particular has consistently high submission rates, which likely contributes to the accuracy of the cwt estimates. Annual submission rates for Northern BC ISBM and Central BC Sport and summer submission rates for Northern Strait of Georgia initially were quite low (</= 10 %) early in the time series, but in the last ten years have had more than half of the submission rates above or very close to the target. This could be due to increasing lodge or guide participation in the SHRP over time. The summer submission rates for Southern WCVI Sport, Southern Strait of Georgia Sport, and Juan de Fuca Sport are all consistently under the 20% target, and have mostly been under 15% in the last ten years.

As noted above, in general, summer submission rates are typically higher than off-season submission rates. Overall, the off-season submission rates could use improvement and are typically under 20%. The off-season submission rates in Juan de Fuca Sport and Southern Georgia Strait tend to be under 10%, and improving these rates is of particular importance since off-season catch in each of those areas was over 7000 in 2023. Northern Strait of Georgia also had off-season catch of over 5000 in 2023, and could use improvement in off-season submission rates. In contrast, the sum total off-season catch in NWCVI, SWCVI, and Johnstone Strait Sport in 2023 was under 900. Overall, all regions outside of Haida Gwaii would benefit from increased participation in the sport head recovery program to strengthen cwt estimates.

\newpage

\begin{landscapepage}
```{r results Fishery Rates, echo=FALSE, fig.height=8, fig.width=12, fig.align="left", out.width="100%", fig.asp=NA}
#| label: fig-fishery-rates-2
#| fig-cap: "Chinook Head submission rate time series for each CWT region and associated time periods (i.e., season or annual)"

region_labels <- c(
  "JDF" = "Juan de Fuca Sport",
  "CBC" = "Central BC Sport",
  "JNST" = "Johnstone Strait Sport",
  "NC" = "Northern BC",
   "NBC AABM" = "Haida Gwaii Sport",
   "NBC ISBM" = "Northern BC ISBM Sport",
  "NGS" = "Strait of Georgia Sport (North)",
  "NWCVI" = "West Coast Vancouver Island Sport (North)",
  "SGS" = "Strait of Georgia Sport (South)",
  "SWCVI" = "West Coast Vancouver Island Sport (South)"
)

region_order<-c("NWCVI", "SWCVI", "NGS", "SGS", "JNST", "JDF", "NBC AABM","NBC ISBM", "CBC")

all_regions <- c("SGS Summer","NGS Summer","JNST Summer","SWCVI Summer","NWCVI Summer","JDF Summer", "SGS Seasonal","NGS Seasonal","JDF Seasonal","JNST Seasonal","WCVI Seasonal", "North AABM", "North ISBM", "Central BC ISBM kept")

out <- run_catch_heads_group(all_regions,
                              x_var_original = "creel_unfiltered_plus",
                              x_var_mr       = "mr_creel_plus_summer", start_y = 2005, end_y=2023)
creel_to_heads2 <- out$creel_to_heads


region_df <- 
  creel_to_heads2 |>
  mutate(season_combo = case_when(season %in% c("spring", "fall") ~ "off-season",  
                                  .default = season)) |>
  mutate(region_combo = case_when(region =="NC" & management == "AABM" ~ "NBC AABM", 
                                  region =="NC" & management == "ISBM" ~ "NBC ISBM",
                                  .default = region)) |>
  group_by(year, status, kept_status, mark_status, region_combo, season_combo, estimate_type, estimate_shape_type) |>
  summarize_if(is.numeric, sum, na.rm = TRUE) |>
  mutate(submit_rate = head_total / catch_estimate) |>
  filter(catch_estimate > 0, catch_estimate > head_total)|>
  mutate(region_combo = factor(region_combo, levels = region_order))|>
  filter(estimate_type=="catch_estimate")

ggplot(data = region_df,
       aes(x=year, y=submit_rate, color=season_combo)) +
  geom_line(linewidth = 1,
            alpha = 0.6) +
  geom_point(size = 1)+
  facet_wrap(~region_combo,
             ncol = 3,
             labeller = labeller(region_combo = region_labels)) +
  scale_color_manual(values =  c("#440154", "#d4be02", "#35b779", "#31688e")) +
  labs(x = "Year", 
       y = "Chinook Head Submission Rate", 
       color = "Season")+
  theme_bw() +
  theme(legend.position="bottom") + geom_hline(yintercept = 0.2, 
             color = "red", 
             linetype = "dashed", 
             linewidth = 0.5, show.legend=TRUE) 
```
\end{landscapepage}
\clearpage

section break  


## Revised catch data in CTC processes

Revised recreational catch estimates are inputs to the CTC's analyses and reporting processes in three primary ways.

- Reported total kept and released catch 
- Marked kept catch estimates are used to estimate CWT recoveries
- Marked and unmarked release rates are used by the ERA to estimate fishery-related incidental mortality

Recreational catch estimate revisions from this analysis were first incorporated into the February 2025 CTC ERA [@CTC2026]. Upon further review of the methods and implementation of the `pacRecCatch` R package for this technical report, some improvements and corrections were identified. While this report focuses on the results of latest version, the changes to methods and data are summarized in Appendix \@ref(apdx-analytic-improvements).

### Reporting Kept and Released Catch
```{r results CandEkept1, echo=FALSE}

erafishery_order<-c("WCVI AABM S", "WCVI ISBM S", "GEO ST S", "JNST S", "BC JF S", "NBC AABM S","NBC ISBM S", "CBC S")

era_model_catch <- 
  sport_release_rates %>% 
  filter(grepl("^T", finescale_fishery) == FALSE) |> #Exclude Terminal finescale fisheries, these just duplicate non-terminal catch in marine
  mutate(erafishery = case_when(
    grepl("CA JDF S", finescale_fishery) ~ "BC JF S",
    grepl("NGS S|SGS S", finescale_fishery) ~ "GEO ST S",
    grepl("WCVI", finescale_fishery) & grepl("AABM", finescale_fishery) ~  "WCVI AABM S",
    grepl("WCVI", finescale_fishery) & grepl("ISBM", finescale_fishery) ~  "WCVI ISBM S",
    grepl("NORTH AABM", finescale_fishery) ~  "NBC AABM S",
    grepl("NBC ISBM|NORTH ISBM", finescale_fishery) ~  "NBC ISBM S",
    grepl("CENTRAL", finescale_fishery) ~  "CBC S",
    grepl("JOHN ST", finescale_fishery) ~  "JNST S",
    .default = NA_character_
  )) %>%
  mutate(kept_total = marked_kept_total + unmarked_kept_total,
         released_total = marked_released_total + unmarked_released_total) %>%
  group_by(year, erafishery) %>%
  summarize(updated_kept_total = sum(kept_total, na.rm=TRUE), 
            updated_released_total = sum(released_total, na.rm=TRUE),
            .groups = "drop")

fishery_model_line <- data.frame(
  erafishery = c("BC JF S", "CBC S", "GEO ST S", 
                 "JNST S", "NBC AABM S", "NBC ISBM S",
                 "WCVI AABM S", "WCVI ISBM S"),
  xintercept = c(2013, 2015, 2013, 2013, 2015, 2015, 2013, 2013))|>
  mutate(erafishery = factor(erafishery, levels = erafishery_order))


rec_summary <- 
  read.csv("data/rec_summary_2005-2023.csv") %>% 
  as_tibble() |>
  rename_with(tolower)

total_catch_compare <- 
  era_model_catch  %>% 
  left_join(rec_summary, by = c("erafishery", "year")) |>
  filter(!is.na(erafishery)) |>
  filter(year<2024) |>
  mutate(erafishery = factor(erafishery, levels = erafishery_order))


total_catch_compare2<- total_catch_compare |> mutate(kept_diff = (updated_kept_total - total_kept)/(total_kept))|>  mutate(kept_diff_abs = abs(kept_diff))

total_catch_compare3<- total_catch_compare2 |>
  group_by(erafishery)|>
  summarize_if(is.numeric, mean, na.rm=TRUE)


#### Summarizing all: 
### make this by finescale fishery old... sum up. 
models_combined<-
  sport_catch_models$model_catch |>
  mutate(region_mgmt = if_else(region %in% c("SWCVI", "NWCVI", "NC"),
                               paste(region, management), 
                               region)) |>
  group_by(year, kept_status, pred_cat, region_mgmt) |> 
  summarise(across(where(is.double), \(x) sum(x, na.rm = TRUE)), .groups = "drop") |>
  mutate(fishery_full_name = recode_values(region_mgmt,
                                           "JDF" ~ "Juan de Fuca Strait",
                                           "JNST" ~ "Johnstone Strait",
                                           "NGS" ~ "Northern Strait of Georgia",
                                           "SGS" ~ "Southern Strait of Georgia",
                                           "NWCVI AABM" ~ "Northern West Vancouver Island - AABM",
                                           "SWCVI AABM" ~ "Southern West Vancouver Island - AABM",
                                           "NWCVI ISBM" ~ "Northern West Vancouver Island - ISBM",
                                           "SWCVI ISBM" ~ "Southern West Vancouver Island - ISBM", 
                                            "NC AABM" ~ "Northern BC - AABM",
                                           "NC ISBM" ~ "Northern BC - ISBM",
                                           "CBC" ~ "Central BC")) 

models_combined2 <- 
  models_combined%>% 
  mutate(erafishery = case_when(
    grepl("JDF", region_mgmt) ~ "BC JF S",
    grepl("NGS|SGS", region_mgmt) ~ "GEO ST S",
    grepl("WCVI", region_mgmt) & grepl("AABM", region_mgmt) ~  "WCVI AABM S",
    grepl("WCVI", region_mgmt) & grepl("ISBM", region_mgmt) ~  "WCVI ISBM S",
    grepl("JNST", region_mgmt) ~  "JNST S",
     grepl("NC AABM", region_mgmt) ~  "NBC AABM S",
    grepl("NC ISBM", region_mgmt) ~  "NBC ISBM S",
    grepl("CBC", region_mgmt) ~  "CBC S",
    .default = NA_character_
  ))|>
  group_by(year, erafishery, kept_status)|>
     summarize_if(is.numeric, sum, na.rm = TRUE)|>
  mutate(creel_unfiltered_plus = case_when(erafishery %in% c("NBC AABM S", "NBC ISBM S", "CBC S") ~ historic_plus, 
                                           .default= creel_unfiltered_plus))

 rec_summary2 <- 
  read.csv("data/rec_summary_2005-2023.csv") %>% 
  as_tibble() |>
  rename_with(tolower)|>
  rename(kept = total_kept, released = legal_releases)|> 
  dplyr::select(-sublegal_releases)|>
  pivot_longer(cols=c(kept, released), values_to = "ce_report", names_to="kept_status")
 
 models_combined2<- models_combined2 |> left_join(rec_summary2) |> dplyr::select(year, erafishery, kept_status, creel_unfiltered_plus, catch_estimate_predicted, ce_report)


  models_combined2_diffs<-  models_combined2 |> mutate(pdiff_irec_creel = (catch_estimate_predicted - creel_unfiltered_plus)/(creel_unfiltered_plus), pdiff_irec_rep = (catch_estimate_predicted - ce_report)/(ce_report), pdiff_creel_rep = (creel_unfiltered_plus- ce_report)/(ce_report)) |> mutate(diff_irec_creel = (catch_estimate_predicted - creel_unfiltered_plus), diff_irec_rep = (catch_estimate_predicted - ce_report), diff_creel_rep = (creel_unfiltered_plus- ce_report))

    models_combined2_diffs<-  models_combined2_diffs %>%
 mutate(across(where(is.numeric), ~replace(., is.infinite(.), NA)))
  
    models_combined2_avs<-  models_combined2_diffs |> group_by(erafishery, kept_status)|>
      summarize_if(is.numeric, c(mean=mean, sd=sd), na.rm=TRUE)

```
As noted previously, revised estimates derived using the methods described in this report were incorporated into CTC reports starting with the February 2025 ERA. Both kept and released catch estimates are typically reported in the Catch and Escapement report by ERA fishery. Here we plot the difference between the revised kept (Figure \@ref(fig:fig-cande-kept)) and released (Figure \@ref(fig:fig-cande-released)) catch estimates, and those recently reported in the 2024 Catch and Escapement report[@CTC2025b], before revised estimates were available. The revised estimates are the product of the finescale fishery results found in Appendix \@ref(app:revised-catch-timeseries), summed to the ERA fishery level. The previously reported catch estimates generally align with the unfiltered creel data, although iREC has been incorporated in some fisheries in recent years. There are also some discrepancies where the reported data does not match either the unfiltered creel data nor the revised estimates and we point out these cases and provide explanations where possible. 

At the ERA fishery level, as at the finescale fishery level, we found that the revised kept catch estimates were typically higher than previously reported estimates that did not incorporate iREC information. In some fisheries, iREC information had been included in recent years (e.g., since 2018 in Strait of Georgia Sport), but the revised estimates incorporate iREC in a standardized way across all fisheries. The largest percent changes in kept catch occurred in Central BC Sport where iREC had not been previously incorporated, but the inclusion of iREC more than doubles recent estimates (e.g. in 2020 the revised catch is `r fmt_stat(models_combined2_diffs$pdiff_irec_rep[models_combined2_diffs$erafishery == "CBC S" & models_combined2_diffs$year ==2020 & models_combined2_diffs$kept_status =="kept"], scale=100)`% larger than the previous estimate, an addition of `r fmt_stat(models_combined2_diffs$diff_irec_rep[models_combined2_diffs$erafishery == "CBC S" & models_combined2_diffs$year ==2020 & models_combined2_diffs$kept_status =="kept"])` pieces). On average, across the time series, the revised catch was `r fmt_stat(models_combined2_avs$pdiff_irec_rep_mean[models_combined2_avs$erafishery == "CBC S" & models_combined2_avs$kept_status =="kept"], scale=100)`% higher (an average difference of `r fmt_stat(models_combined2_avs$diff_irec_rep_mean[models_combined2_avs$erafishery == "CBC S" & models_combined2_avs$kept_status =="kept"])` pieces) than previously reported. This is consistent with the sport head recovery data that indicates previous CBC catch from logbooks was an underestimate. Other large changes occurred in Strait of Georgia sport before 2012 (iREC had already been incorporated from 2013 onward), where the revised catch is `r fmt_stat(models_combined2_diffs$pdiff_irec_rep[models_combined2_diffs$erafishery == "GEO ST S" & models_combined2_diffs$year ==2012 & models_combined2_diffs$kept_status =="kept"], scale=100)` to `r fmt_stat(models_combined2_diffs$pdiff_irec_rep[models_combined2_diffs$erafishery == "GEO ST S" & models_combined2_diffs$year ==2008 & models_combined2_diffs$kept_status =="kept"], scale=100)`% larger (`r fmt_stat(models_combined2_diffs$diff_irec_rep[models_combined2_diffs$erafishery == "GEO ST S" & models_combined2_diffs$year ==2012 & models_combined2_diffs$kept_status =="kept"])` to `r fmt_stat(models_combined2_diffs$diff_irec_rep[models_combined2_diffs$erafishery == "GEO ST S" & models_combined2_diffs$year ==2005 & models_combined2_diffs$kept_status =="kept"])` more pieces) than previous catch. Overall, on average Georgia Strait increased by `r fmt_stat(models_combined2_avs$pdiff_irec_rep_mean[models_combined2_avs$erafishery == "GEO ST S" & models_combined2_avs$kept_status =="kept"], scale=100)`% from previously reported values. These large changes are expected as this fishery has some of the highest off-season fishing which was not previously well captured and has some of the most challenges with creel surveying. Next, Northern BC ISBM Sport had `r fmt_stat(models_combined2_avs$pdiff_irec_rep_mean[models_combined2_avs$erafishery == "NBC ISBM S" & models_combined2_avs$kept_status =="kept"], scale=100)`% higher estimates than previously reported on average, an average of `r fmt_stat(models_combined2_avs$diff_irec_rep_mean[models_combined2_avs$erafishery == "NBC ISBM S" & models_combined2_avs$kept_status =="kept"])` additional pieces per year. The catch from 2005-2007 in Northern ISBM was `r fmt_stat(models_combined2_diffs$pdiff_irec_rep[models_combined2_diffs$erafishery == "NBC ISBM S" & models_combined2_diffs$year ==2005 & models_combined2_diffs$kept_status =="kept"], scale=100)`% higher than previous reported, however, the creel and logbook data has since been updated from ~8000 to ~10,000 pieces, making the percent change to the newly modelled catch at ~11,000 pieces much less dramatic (`r fmt_stat(models_combined2_diffs$pdiff_irec_creel[models_combined2_diffs$erafishery == "NBC ISBM S" & models_combined2_diffs$year ==2005 & models_combined2_diffs$kept_status =="kept"], scale=100)`%). The remainder of the upward revisions were less than 10% change on average across the time series (although individual years varied and were up to 30%). More details can be found in table format in Appendix \@ref(apdx-analytic-improvements).

```{r results CandEkept, results='asis', echo=FALSE, fig.height=5, fig.width=10, out.width="100%", fig.asp=NA}
#| label: fig-cande-kept
#| fig-cap: "Revised (blue) and previously reported (green) kept catch estimates for ERA fisheries from 2005-2023. The vertical line on each graph identifies the first year that revised estimates are based on only on observed data (e.g., iREC and creel).  Years before the vertical line are modelled estimates that utilize past catch estimates."

## Kept only
k<-ggplot() +
  geom_point(data=total_catch_compare, aes(y=updated_kept_total, x=year, col="lightblue4")) + 
  geom_line(data=total_catch_compare, aes(y=updated_kept_total, x=year, col="lightblue4"))+
  geom_point(data=total_catch_compare, aes(y=total_kept, x=year, col="lightgreen")) + 
  geom_line(data=total_catch_compare, aes(y=total_kept, x=year, col="lightgreen"))+
  scale_color_manual(values=c("lightblue4", "lightgreen"),  
                     labels = c("Revised estimates", "2024 C&E Report"))+
  facet_wrap(~erafishery, 
             scales = "free", 
             ncol=4)+
  geom_vline(data = fishery_model_line, aes(xintercept = xintercept)) +
  labs(col = "Data source",
       y = "Kept Catch",
       x = "Year") +
  theme_bw() +
  theme(legend.position="bottom") +
  coord_cartesian(ylim = c(0, NA))

print(k)

```

```{r results CandErelease, results='asis', echo=FALSE, fig.height=5, fig.width=10, out.width="100%", fig.asp=NA}
#| label: fig-cande-released
#| fig-cap: "Revised (blue) and previously reported (green) legal-sized released catch estimates from 2005-2023. The vertical line on each graph identifies the first year that revised estimates are based only on observed data (e.g., iREC and creel). Years before the vertical line are modelled based on past catch estimates."

rec_summary <- 
  read.csv("data/rec_summary_2005-2023.csv") |>
  as_tibble() |>
  rename_with(tolower)

total_catch_compare <- 
  era_model_catch  %>% 
  left_join(rec_summary, by = c("erafishery", "year")) |>
  filter(!is.na(erafishery))|>
  filter(year<2024)|>
  mutate(erafishery = factor(erafishery, levels = erafishery_order))


## Released only
ggplot() +
  geom_point(data=total_catch_compare, aes(y=updated_released_total, x=year, col="lightblue4")) + 
  geom_line(data=total_catch_compare, aes(y=updated_released_total, x=year, col="lightblue4"))+
  geom_point(data=total_catch_compare, aes(y=legal_releases, x=year, col="lightgreen")) + 
  geom_line(data=total_catch_compare, aes(y=legal_releases, x=year, col="lightgreen"))+
  scale_color_manual(values=c("lightblue4", "lightgreen"),  
                     labels = c("Revised estimates", "2024 C&E Report"))+
  facet_wrap(~erafishery, 
             scales = "free", 
             ncol=4)+
  geom_vline(data = fishery_model_line, aes(xintercept = xintercept)) +
  labs(col = "Data source",
       y = "Legal-Size Released Catch",
       x = "Year") +
  theme_bw() +
  theme(legend.position="bottom")+
  coord_cartesian(ylim = c(0, NA))

```

There were a few downward changes in kept catch, most due to previous errors in reporting that resulted in incongruous data between the reported estimate in the 2024 C&E report and the unfiltered creel and logbook estimate from the database (see Appendix \@ref(app:revised-catch-timeseries)). For example, in NBC ISBM S 2022 where the reported value in the 2024 ERA was `r fmt_stat(models_combined2_diffs$pdiff_irec_rep[models_combined2_diffs$erafishery == "NBC ISBM S" & models_combined2_diffs$year ==2022 & models_combined2_diffs$kept_status =="kept"], scale=-100)`% more than the revised estimate, despite creel and logbook estimate for that finescale fishery being less than the revised estimate). In this case it is that likely that the true creel and logbook estimate had been revised and updated in the database in recent years but not reflected in the reporting. Another example of downward revision is in 2005 and 2006 in Johnstone Strait, where methods from Argue et al. 1997 were used to enhance previously reported catch based on sport head recoveries. These methods were used in the C&E report from 2012 but are now discontinued. In Juan de Fuca Sport in 2018 and 2019, iREC was incorporated in the 2024 C&E report but was overestimated due to an incorrect bias correction factor from the iREC calibration, more details can be found in Appendix \@ref(apdx-analytic-improvements). A downward revision that is consistent with the creel estimate is a `r fmt_stat(models_combined2_diffs$pdiff_irec_rep[models_combined2_diffs$erafishery == "JNST S" & models_combined2_diffs$year ==2019 & models_combined2_diffs$kept_status =="kept"], scale=-100)`% reduction in the Johnstone Strait 2019 estimate using revised methods. This is likely due to poor quality creel causing overestimation as described above. A similar pattern was found in 2021 in the same fishery (`r fmt_stat(models_combined2_diffs$pdiff_irec_rep[models_combined2_diffs$erafishery == "JNST S" & models_combined2_diffs$year ==2021 & models_combined2_diffs$kept_status =="kept"], scale=100)`%). Another downward revision is in NBC ISBM 2019, a `r fmt_stat(models_combined2_diffs$pdiff_irec_rep[models_combined2_diffs$erafishery == "NBC ISBM S" & models_combined2_diffs$year ==2019 & models_combined2_diffs$kept_status =="kept"], scale=-100)`% reduction (or `r fmt_stat(models_combined2_diffs$diff_irec_rep[models_combined2_diffs$erafishery == "NBC ISBM S" & models_combined2_diffs$year ==2019 & models_combined2_diffs$kept_status =="kept"], scale=-1)` fewer pieces). While there is no immediate explanation, the creel and logbook data are likely an overestimation in this case as the revised data fits the Northern BC Sport Head Recovery data much better. The remainder of the negative revisions are under 10% in a given year and can be found in table format in Appendix \@ref(apdx-analytic-improvements).

Released catch was substantially increased from the reported releases, particularly in the pre-iREC period where very few releases were estimated (Figure \@ref(fig:fig-cande-released)). While our models assume that the relationship between the creel estimate and combined creel and iREC estimate does not change between the pre-iREC and the during-iREC periods, this revision necessarily added large amounts of releases for regions where lodges and creel surveys did not request that information. For example, the highest percent changes in released catch occurred in Central BC Sport where previously very almost no releases were reported (an addition of `r fmt_stat(models_combined2_avs$diff_irec_rep_mean[models_combined2_avs$erafishery == "CBC S" & models_combined2_avs$kept_status =="released"])` pieces on average per year across the time series, `r fmt_stat(models_combined2_avs$pdiff_irec_rep_mean[models_combined2_avs$erafishery == "CBC S" & models_combined2_avs$kept_status =="released"], scale=100)`% increase). This is because lodges in Central BC do not regularly report released catch, whereas iREC allows for released catch to be reported. In CBC, the static value of ~15000 releases between 2005-2014 is essentially an average of the 2015-2023 period, and a reflection of the models having zero input for releases. In other fisheries, where creel did estimate some small number of releases, the modelled releases can fluctuate and accommodate for changing fisher behaviour over time (e.g. Strait of Georgia, Strait of Juan de Fuca Sport). 

On average across the time series, where releases were partially estimated, the number of releases increased by `r fmt_stat(models_combined2_avs$pdiff_irec_rep_mean[models_combined2_avs$erafishery == "GEO ST S" & models_combined2_avs$kept_status =="released"], scale=100)`% (`r fmt_stat(models_combined2_avs$diff_irec_rep_mean[models_combined2_avs$erafishery == "GEO ST S" & models_combined2_avs$kept_status =="released"])` pieces) in the Strait of Georgia, `r fmt_stat(models_combined2_avs$pdiff_irec_rep_mean[models_combined2_avs$erafishery == "BC JF S" & models_combined2_avs$kept_status =="released"], scale=100)`% (`r fmt_stat(models_combined2_avs$diff_irec_rep_mean[models_combined2_avs$erafishery == "BC JF S" & models_combined2_avs$kept_status =="released"])` pieces) in Juan de Fuca, `r fmt_stat(models_combined2_avs$pdiff_irec_rep_mean[models_combined2_avs$erafishery == "JNST S" & models_combined2_avs$kept_status =="released"], scale=100)`% (`r fmt_stat(models_combined2_avs$diff_irec_rep_mean[models_combined2_avs$erafishery == "JNST S" & models_combined2_avs$kept_status =="released"])` pieces) in Johnstone Strait, `r fmt_stat(models_combined2_avs$pdiff_irec_rep_mean[models_combined2_avs$erafishery == "WCVI AABM S" & models_combined2_avs$kept_status =="released"], scale=100)`% (`r fmt_stat(models_combined2_avs$diff_irec_rep_mean[models_combined2_avs$erafishery == "WCVI AABM S" & models_combined2_avs$kept_status =="released"])` pieces) in WCVI AABM and `r fmt_stat(models_combined2_avs$pdiff_irec_rep_mean[models_combined2_avs$erafishery == "WCVI ISBM S" & models_combined2_avs$kept_status =="released"], scale=100)`% (`r fmt_stat(models_combined2_avs$diff_irec_rep_mean[models_combined2_avs$erafishery == "WCVI ISBM S" & models_combined2_avs$kept_status =="released"])` pieces) in WCVI ISBM. Northern BC ISBM Sport also had large percent changes, particularly in 2005-2012 period where no releases were previously reported (on average `r fmt_stat(models_combined2_avs$diff_irec_rep_mean[models_combined2_avs$erafishery == "NBC ISBM S" & models_combined2_avs$kept_status =="released"])` pieces and `r fmt_stat(models_combined2_avs$pdiff_irec_rep_mean[models_combined2_avs$erafishery == "NBC ISBM S" & models_combined2_avs$kept_status =="released"], scale=100)`% change across the whole time series. Haida Gwaii Sport also had the largest average numeric increases of `r fmt_stat(models_combined2_avs$diff_irec_rep_mean[models_combined2_avs$erafishery == "NBC AABM S" & models_combined2_avs$kept_status =="released"])` pieces, `r fmt_stat(models_combined2_avs$pdiff_irec_rep_mean[models_combined2_avs$erafishery == "NBC AABM S" & models_combined2_avs$kept_status =="released"], scale=100)`% change. 

Released catch was revised downwards in a handful of cases. Georgia Strait releases in 2014 were lower than reported in C&E 2024 (`r fmt_stat(models_combined2_diffs$pdiff_irec_rep[models_combined2_diffs$erafishery == "GEO ST S" & models_combined2_diffs$kept_status =="released" & models_combined2_diffs$year ==2014], scale=100)`%), however, the reported value is much higher than the creel value, so this is likely an earlier iteration of iREC that was previously incorporated. WCVI ISBM releases in 2012 follow the same pattern (`r fmt_stat(models_combined2_diffs$pdiff_irec_rep[models_combined2_diffs$erafishery == "WCVI AABM S" & models_combined2_diffs$kept_status =="released" & models_combined2_diffs$year ==2012], scale=100)`%), and as 2012 was the first year of the piloting of iREC, it was possibly included in the reporting and carried forward. NBC in 2019 had a negative revision of `r fmt_stat(models_combined2_diffs$pdiff_irec_rep[models_combined2_diffs$erafishery == "NBC AABM S" & models_combined2_diffs$kept_status =="released" & models_combined2_diffs$year ==2019], scale=100)`%  for released catch. This aligns with the unexplained kept catch downwards revision above for the same year. The remaining negative revisions are all small (</=10%). 


\newpage
### Updating CWT estimates
```{r results indcwtest2, results='asis', echo=FALSE}

cwt_est_filename <- "data/cwt_est_by_fishery.sql"
cwt_est_sql <- readChar(cwt_est_filename, file.info(cwt_est_filename)$size)

new_camp <- ctcUtil::open_camp_connection("camp_2025_local.yaml")
prev_camp <- ctcUtil::open_camp_connection("camp_2024.yaml")

new_cwt_est <- 
  dbGetQuery(new_camp, cwt_est_sql) |>
  mutate(fishery = paste0(fishery_id, " - ", fishery_name))|>
  mutate(fishery = factor(fishery, levels=c("50 - WCVI AABM S", "51 - WCVI ISBM S", "52 - TWCVI TERM S",  "63 - GEO ST S", "64 - TGEO ST TERM S", "61 - JNST S", 
                "62 - TJNST TERM S", 
                "65 - BC JF S", "66 - TBC JF TERM S", 
                "47 - NBC AABM S", "48 - NBC ISBM S", "49 - TNBC TERM S",  "45 - CBC S", "46 - TCBC TERM S")))


prev_cwt_est <- 
  dbGetQuery(prev_camp, cwt_est_sql) |>
  mutate(fishery = paste0(fishery_id, " - ", fishery_name))|>
  mutate(fishery = factor(fishery, levels=c("50 - WCVI AABM S", "51 - WCVI ISBM S", "52 - TWCVI TERM S",  "63 - GEO ST S", "64 - TGEO ST TERM S", "61 - JNST S", 
                "62 - TJNST TERM S", 
                "65 - BC JF S", "66 - TBC JF TERM S", 
                "47 - NBC AABM S", "48 - NBC ISBM S", "49 - TNBC TERM S",  "45 - CBC S", "46 - TCBC TERM S")))


dbDisconnect(prev_camp)
dbDisconnect(new_camp)

cwt_fishery_model_line <-
  data.frame(
    fishery = c("65 - BC JF S", "66 - TBC JF TERM S", 
                "63 - GEO ST S", "64 - TGEO ST TERM S", "61 - JNST S", 
                "62 - TJNST TERM S", 
                "50 - WCVI AABM S", "51 - WCVI ISBM S", "52 - TWCVI TERM S",
                "45 - CBC S", "46 - TCBC TERM S",
                "47 - NBC AABM S", "48 - NBC ISBM S", "49 - TNBC TERM S"),
    xintercept = c(2013, 2013, 2013, 2013, 2013, 2013, 2013, 2013, 2013,
                   2015, 2015, 2015, 2015, 2015)
  )|> mutate(fishery = factor(fishery, levels=c("50 - WCVI AABM S", "51 - WCVI ISBM S", "52 - TWCVI TERM S",  "63 - GEO ST S", "64 - TGEO ST TERM S", "61 - JNST S", 
                "62 - TJNST TERM S", 
                "65 - BC JF S", "66 - TBC JF TERM S", 
                "47 - NBC AABM S", "48 - NBC ISBM S", "49 - TNBC TERM S",  "45 - CBC S", "46 - TCBC TERM S")))

new_cwt_est2<- new_cwt_est |> rename(new_cwt_estimate = cwt_estimate)

cwt_all<- new_cwt_est2 |> left_join(prev_cwt_est) |> mutate(cwt_diff = new_cwt_estimate - cwt_estimate, 
                                                            cwt_pdiff = (new_cwt_estimate - cwt_estimate)/(cwt_estimate))


cwt_all_avg<- cwt_all |> group_by(fishery_name) |> summarize_if(is.numeric, c(mean=mean, sd=sd), na.rm=TRUE)

```
As described in the methods, kept and marked catch estimates are used to expand CWT observations from the SHRP to CWT estimates uploaded to RMIS. Canada updated CWT estimates based on revised catch and published them to RMIS in support of ERA process starting in February 2025. Figure \@ref(fig:fig-indcwtest) shows the time series of total CWT estimates for all ERA indicator stocks, including both the original 2024 estimates and the revised estimates based on revised catch data. Generally, the revised data increase the total CWT estimates across all time series, although there are some declines in a few cases. As expected, most of the differences between the 2024 and revised CWT estimates are in earlier years, before the incorporation of iREC (identified by the vertical line). The smallest impacts of the revisions occurred in Haida Gwaii Sport, (mean of `r fmt_stat(cwt_all_avg$cwt_pdiff_mean[cwt_all_avg$fishery_name == "NBC AABM S"], scale=100)`% change from previously reported to revised). As above, the largest impacts of the revisions occur in Strait of Georgia Sport (mean of `r fmt_stat(cwt_all_avg$cwt_pdiff_mean[cwt_all_avg$fishery_name == "GEO ST S"], scale=100)`%, an average of `r fmt_stat(cwt_all_avg$cwt_diff_mean[cwt_all_avg$fishery_name == "GEO ST S"])` more CWTs estimated per year) and Central BC Sport (mean of `r fmt_stat(cwt_all_avg$cwt_pdiff_mean[cwt_all_avg$fishery_name == "CBC S"], scale=100)`%, an average of `r fmt_stat(cwt_all_avg$cwt_diff_mean[cwt_all_avg$fishery_name == "CBC S"])` more CWTs estimated), and their terminal fisheries. Although it should be noted that Central BC in recent years has used iREC catch in CWT estimation to provide more accurate CWT estimates. WCVI AABM also had an average increase of `r fmt_stat(cwt_all_avg$cwt_diff_mean[cwt_all_avg$fishery_name == "WCVI AABM S"])` CWTs per year, a `r fmt_stat(cwt_all_avg$cwt_pdiff_mean[cwt_all_avg$fishery_name == "WCVI AABM S"], scale=100)`% increase, and similar pattern for the terminal component. The remainder of the fisheries had fewer than 100 CWT differences (indicator stocks summed) per year between the revised and reported estimates on average across the time series. This update shows more consistent ranges in values over the whole revised time series (e.g., ERA fishery 64 - T GEO ST TERM S) compared to those in the 2024 time series. This suggests that the general fishery sample rate has been relatively consistent and that changes in indicator stock CWT estimates were potentially driven more by biases in historic marked catch estimates. Therefore, revised CWT estimates should provide a more consistent time series to evaluate indicator stock impacts. 

```{r results indcwtest, results='asis', echo=FALSE,fig.height=10, fig.width=10, out.width="100%", fig.asp=NA}
#| label: fig-indcwtest
#| fig-cap: "Time series of CWT estimate totals for all PSC indicator stocks within the CTC 2025 ERA for each ERA fishery. The vertical line on each panel identifies the first year that revised CWT estimates used catch estimates based primarily on observed data (e.g., iREC and creel). Years before the vertical line use modelled catch based on past catch estimates."

## CWT Estimate Total time series
ggplot() +
  geom_point(data=new_cwt_est, aes(y=cwt_estimate, x=run_year, col="lightblue4")) + 
  geom_line(data=new_cwt_est, aes(y=cwt_estimate, x=run_year, col="lightblue4"))+
  geom_point(data=prev_cwt_est, aes(y=cwt_estimate, x=run_year, col="lightgreen")) + 
  geom_line(data=prev_cwt_est, aes(y=cwt_estimate, x=run_year, col="lightgreen"))+
  scale_color_manual(values=c( "lightblue4", "lightgreen"),
                     labels = c("Revised CWT Estimates", "2024 CWT Estimates"))+
  ylab("CWT Estimate (All Indicators)")+
  xlab("Run Year")+
  facet_wrap(~fishery, scales = "free", ncol=4)+
  theme_bw() + 
  theme(legend.position="bottom")+
  geom_vline(data = cwt_fishery_model_line, aes(xintercept = xintercept)) +
  labs(col = "Data source")+
  coord_cartesian(ylim = c(0, NA))

```

\newpage

### Updating Chinook non-retention (release) estimates
```{r results mrrukr1, results='asis', echo=FALSE}

fishery_map <-
  sport_catch_models$model_catch %>% 
  select(region, management) %>%
  mutate(erafishery = case_when(
    region %in% c("SWCVI", "NWCVI") & management == "AABM" ~ "WCVI AABM S",
    region %in% c("SWCVI", "NWCVI") & management == "ISBM" ~ "WCVI ISBM S",
    region %in% c("NGS", "SGS") ~  "GEO ST S",
    region == "JDF" ~ "BC JF S",
    region == "CBC" ~ "CBC S",
    region == "NC" & management == "AABM" ~ "NBC AABM S",
    region == "NC" & management == "ISBM" ~ "NBC ISBM S",
    region == "JNST" ~ "JNST S",
    .default = "Unknown"
  )) %>% 
  distinct()

cnr_canada_sport <- 
  read.csv("data/cnr_canada_sport.csv") %>% 
  as_tibble() %>% 
  rename_with(tolower) %>%
  filter(year > 2004) %>%
  full_join(fishery_map, by= c("erafishery"), relationship = "many-to-many")


finescale_fishery_map <- 
  sport_catch_models$model_catch %>% 
  select(finescale_fishery, region, management) %>% 
  distinct() |>
   mutate(erafishery = case_when(
    grepl("CA JDF S", finescale_fishery) ~ "BC JF S",
    grepl("NGS S|SGS S", finescale_fishery) ~ "GEO ST S",
    grepl("WCVI", finescale_fishery) & grepl("AABM", finescale_fishery) ~  "WCVI AABM S",
    grepl("WCVI", finescale_fishery) & grepl("ISBM", finescale_fishery) ~  "WCVI ISBM S",
    grepl("NORTH AABM", finescale_fishery) ~  "NBC AABM S",
    grepl("NBC ISBM|NORTH ISBM", finescale_fishery) ~  "NBC ISBM S",
    grepl("CENTRAL", finescale_fishery) ~  "CBC S",
    grepl("JOHN ST", finescale_fishery) ~  "JNST S",
    .default = NA_character_
  )) 


updated_relese_rates <-
  sport_release_rates |>
  inner_join(finescale_fishery_map) |>
  group_by(year, erafishery) |>
  summarize(marked_kept_total = sum(marked_kept_total, na.rm = TRUE),
            marked_released_total = sum(marked_released_total, na.rm = TRUE),
            unmarked_kept_total = sum(unmarked_kept_total, na.rm = TRUE),
            unmarked_released_total = sum(unmarked_released_total, na.rm = TRUE),
            .groups = "drop") |>
  mutate(mark_total = if_else(marked_kept_total > 0 | marked_released_total > 0,
                              marked_kept_total + marked_released_total,
                              0.0000001), #small value to prevent divide by zero
         unmark_total = if_else(unmarked_kept_total > 0 | unmarked_released_total > 0,
                                unmarked_kept_total + unmarked_released_total,
                                0.0000001), #small value to prevent divide by zero
         mrr = marked_released_total / mark_total,
         urr = unmarked_released_total / unmark_total, 
         rr = (marked_released_total + unmarked_released_total) / (mark_total + unmark_total)) |>
  select(year, erafishery, mrr, urr, rr)


sport_mark_rate_mrr <-
  cnr_canada_sport |>
  full_join(updated_relese_rates) |>
  mutate(across(c(mrr, urr, rr), \(x) coalesce(x, 0)),
         region = if_else(region == "NC", "NBC", region),
         region_mgmt = paste(region, management))|> filter(year<2024)|>
  mutate(region_mgmt = factor(region_mgmt, levels = c("NWCVI AABM","NWCVI ISBM","SWCVI AABM",  
                     "SWCVI ISBM","NGS ISBM", "SGS ISBM", "JNST ISBM", "JDF ISBM","NBC AABM", 
                    "NBC ISBM",  "CBC ISBM"
                     )))|>
  mutate(erafishery = factor(erafishery, levels = erafishery_order))


mrr_fishery_model_line <-
  data.frame(
    erafishery = c("CBC S", "NBC AABM S", 
                    "NBC ISBM S", "WCVI AABM S", 
                    "WCVI ISBM S", "JNST S",
                    "GEO ST S", "BC JF S"),
    xintercept = c(rep(2015, 3), rep(2013, 5))
  )|>
  mutate(erafishery = factor(erafishery, levels = erafishery_order))

rr_sum<- sport_mark_rate_mrr |> group_by(erafishery)|>
  summarize_if(is.numeric, mean, na.rm=TRUE)
```
Kept and released catch estimates are incorporated for most sport fisheries into the ERA non-retention method. In 2024, the Chinook ERA model was modified to account for mark selective retention fisheries with the addition of the marked release rate ($MRR$) and unmarked kept rate ($UKR$) [see @CTC2025 for methods]. We calculated the $MRR$ and $UKR$ at the PFMA- and month-level, and then used a weighted average based on total catch estimate per area and month to generate an $MRR$ and $UKR$ for each finescale fishery. We then calculate a stock-specific MRR and UKR at the ERA fishery level weighted by CWT recoveries. Therefore, in revised estimates, we allow release rates to vary by season, whereas in the past, finescale fisheries were not split into seasons and therefore release rates in each ERA fishery applied to all seasons.

Here, for comparison to previous years ERA-level release rates, we show the summed (non-seasonal and across stocks) marked and unmarked release rates at the ERA level (Figure \@ref(fig:fig-mrrukr2). In particular, for more direct comparison to $MRR$ in the figures we show the unmarked release rate $URR$ (i.e., $1 - UKR$), although $UKR$ is actually used in the ERA. We compare these marked and unmarked release rates based on revised catch estimates to release rates generated by the ratio between kept and released fish in the Chinook Non-Retention (CNR) file used in the 2024 ERA. With the exception of Juan de Fuca Strait, which did have some mark-selective fisheries before 2012, all other sport fisheries were regulated as non-selective before 2013. Since the underlying catch is modelled before 2013 we have fixed the $MRR$ and $URR$ based on modelled catch to be equal (i.e., no mark selectivity), by calculating a single release rate for marked and unmarked fish combined. This removes any potential bias from the modelling process that might have affected marked and unmarked catch estimates differently (e.g., larger uncertainty in modelled marked derived estimates because of lower magnitude of catch). After 2013, the data is observed and therefore allowed to vary selectively even if no specific mark selective fishery was specified in regulations. When $MRR$ and $URR$ are identical, this indicates non-selectivity. If $MRR$ is higher than $URR$, this indicates marked fish are released at a higher rate than unmarked fish, which would be indicative of preferential selection of unmarked fish for retention. If $URR$ is higher than $MRR$, then marked fish are released at a lower rate than unmarked fish, indicating preferential mark retention. 

The Exploitation rate analysis has been accounting for releases since before 2005 for some fisheries (e.g.West Coast Vancouver Island and Haida Gwaii Sport), while for others release information has only recently been incorporated (e.g. since 2019 in Strait of Georgia, Juan de Fuca Sport, and Johnstone Strait Sport). Where release rates were reported, revised release rates are generally higher than reported 2024 CNR release rates, although NBC ISBM S has a few years where the opposite is the case. While only recently incorporated into the ERA, releases have been occurring throughout the time series and release rates have fluctuated over time. The highest average release rates occur in Haida Gwaii (`r fmt_stat(rr_sum$rr[rr_sum$erafishery == "NBC AABM S"], scale=100)`%) and Central BC (`r fmt_stat(rr_sum$rr[rr_sum$erafishery == "CBC S"], scale=100)`%) which are relatively stable throughout the time series. WCVI ISBM has the lowest release rates on average (`r fmt_stat(rr_sum$rr[rr_sum$erafishery == "WCVI ISBM S"], scale=100)`%), followed by Johnstone Strait (`r fmt_stat(rr_sum$rr[rr_sum$erafishery == "JNST S"], scale=100)`%) and WCVI AABM (`r fmt_stat(rr_sum$rr[rr_sum$erafishery == "WCVI AABM S"], scale=100)`%), all relatively stable throughout time. There not appear to be strong selectivity (MRR>URR) in any of these fisheries. The Strait of Georgia, Juan de Fuca Strait, and Northern BC ISBM have all increased release rates since 2019. In Southern BC, these changes correspond with introductions of mark selective fishing regulations, and indeed mark selectivity is occurring in these fisheries in recent years (MRR > URR). In Northern BC, release rates are increasing but it does not appear to be linked with changes in regulation. Furthermore, there is a higher release rate for unmarked fish than marked fish (URR > MRR) in this fishery, which indicates a selectivity for unmarked fish. This could be a true preference for unmarked fish in this fishery or a product of uncertainty in mark rates. 

```{r results mrrukr2, results='asis', echo=FALSE, fig.height=10, fig.width=10}
#| label: fig-mrrukr2
#| fig-cap: "Release rate time series at the finescale fishery level, from 2005 to 2023. Light blue lines are release rates of unmarked fish and dark blue lines are release rates of marked fish from revised data. Green lines are release rates (regardless of mark status) used in the 2024 CNR files applied to the 2024 ERA."
mu <- 
  ggplot() +
  geom_point(data=sport_mark_rate_mrr %>% filter(between(year, 2015, 2023)), aes(y=urr, x=year, col="lightblue")) +
  geom_line(data=sport_mark_rate_mrr %>% filter(between(year, 2015, 2023)), aes(y=urr, x=year, col="lightblue"))+
  geom_point(data=sport_mark_rate_mrr%>% filter(between(year, 2015, 2023)), aes(y=mrr, x=year, col="lightblue4")) +
  geom_line(data=sport_mark_rate_mrr%>% filter(between(year, 2015, 2023)), aes(y=mrr, x=year, col="lightblue4"))+
  geom_point(data=sport_mark_rate_mrr %>% filter(between(year, 2013, 2015), erafishery %notin% c("NBC AABM S", "NBC ISBM S", "CBC S")), aes(y=urr, x=year, col="lightblue")) +
  geom_line(data=sport_mark_rate_mrr %>% filter(between(year, 2013, 2015), erafishery %notin% c("NBC AABM S", "NBC ISBM S", "CBC S")), aes(y=urr, x=year, col="lightblue"))+
  geom_point(data=sport_mark_rate_mrr%>% filter(between(year, 2013, 2015), erafishery %notin% c("NBC AABM S", "NBC ISBM S", "CBC S")), aes(y=mrr, x=year, col="lightblue4")) +
  geom_line(data=sport_mark_rate_mrr%>% filter(between(year, 2013, 2015), erafishery %notin% c("NBC AABM S", "NBC ISBM S", "CBC S")), aes(y=mrr, x=year, col="lightblue4"))+
  geom_point(data=sport_mark_rate_mrr, aes(y=release_rate, x=year, col="lightgreen")) +
  geom_line(data=sport_mark_rate_mrr, aes(y=release_rate, x=year, col="lightgreen"))+
   geom_point(data=sport_mark_rate_mrr %>% filter(between(year, 2005, 2013)), aes(y=rr, x=year, col="orange")) +
  geom_line(data=sport_mark_rate_mrr %>% filter(between(year, 2005, 2013)), aes(y=rr, x=year, col="orange"))+
    geom_point(data=sport_mark_rate_mrr %>% filter(between(year, 2013, 2015), erafishery %in% c("NBC AABM S", "NBC ISBM S", "CBC S")), aes(y=rr, x=year, col="orange")) +
  geom_line(data=sport_mark_rate_mrr %>% filter(between(year, 2013, 2015), erafishery %in% c("NBC AABM S", "NBC ISBM S", "CBC S")), aes(y=rr, x=year, col="orange"))+
  scale_color_manual(values=c("lightblue", "lightblue4", "lightgreen", "orange"),  
                     labels = c("Revised Unmarked Release Rate", 
                                "Revised Marked Release Rate", 
                                "2024 CNR Release Rate", "Revised Release Rate"))+
  facet_wrap(~erafishery) +
  theme_bw() + 
  geom_vline(data = mrr_fishery_model_line, aes(xintercept = xintercept)) +
  labs(col="Data source",
       x = "Year",
       y = "Release Rate") +
  theme(legend.position = "bottom")

print(mu)



```