---
output:
  pdf_document: default
  html_document: default
  word_document: default
---

\clearpage

# Results

In the following section we present the results of revising Chinook catch using both the combined iREC, logbook, and creel estimates for more recent years and modeled estimates for pre-iREC years back to 2005. In addition to revising catch, the relationship between marked kept Chinook catch and CWT sampling through the SHRP is presented, including sample rate time series plots for each CWT region.

## Revised Chinook catch estimates with iREC (2013-2023)

We found that the "best" catch estimates that included filtered creel and calibrated iREC data were broadly higher than previous estimates derived from only unfiltered creel and logbook information. As an example, Figure \@ref(fig:fig-ngs-catch-revised-1) shows the effect of the summed and revised 2013 to 2023 kept catch estimates for the Northern Straight of Georgia finescale fishery (solid line) compared to the previous estimates without iREC inclusion (dashed line). Though the relative difference between the two time series vary, the revised estimates are consistently higher than the estimates without iREC. As demonstrated by Figure \@ref(fig:fig-creel-irec-NGS), calibrated iREC fills in spatial and temporal gaps in creel sampling, so it follows that the summed yearly estimates are typically higher when you include calibrated iREC, especially in regions with many gaps to fill. As the comparison plots for the other fisheries shows in Appendix \@ref(app:revised-catch-timeseries), in most cases kept catch was increased with a few exceptions where there was a small reduction in specific years. For example, in Johnston Strait in 2019, the iREC-included catch estimate was lower than the unfiltered creel and logbook estimate (see Figure \@ref(fig:fig-catch-revised-1-appendix). This could be due to overestimation by the unfiltered creel: 7 out of 9 estimates in this region/year failed the criteria and are therefore likely estimating catch based only on a small number of days in a month (e.g. a long weekend) that is then expanded to the entire month.

```{r results example summer m plots, results='asis', echo=FALSE}
#| label: fig-ngs-catch-revised-1
#| fig-cap: "North Strait of Georgia yearly kept catch estimates based on both previous unfiltered creel/logbook estimates and the revised estimates that includes iREC"


### make this by finescale fishery old... sum up. 
jdf_graph_data <-   
  sport_catch_models$model_catch |> 
  filter(region == "NGS",
         between(year, 2013, 2023),
         kept_status == "kept") |>
  select(year, creel_unfiltered_plus, catch_estimate_predicted) |>
  pivot_longer(c(creel_unfiltered_plus, catch_estimate_predicted),
               names_to = "estimate_type",
               values_to = "catch_estimate") |>
  group_by(year, estimate_type) |>
  summarize(catch_estimate = sum(catch_estimate, na.rm = TRUE), 
            .groups = "drop")

ggplot(jdf_graph_data,
       aes(y = catch_estimate,
           x = year,
           shape = estimate_type,
           linetype = estimate_type))+
  geom_point(size = 2.5,
             colour = "#440154")+
  geom_line(colour = "#440154") +
  scale_shape_manual(values = c(16, 1),
                     labels = c("calibrated iREC included", "unfiltered creel and logbook")
  ) +
  scale_linetype_manual(values = c(1, 2),
                        labels = c("calibrated iREC included", "unfiltered creel and logbook")
  ) +
  scale_x_continuous(breaks = c(2013, 2015, 2017, 2019, 2021, 2023)) +
  ggtitle("Northern Strait of Georgia")  + 
  theme_bw() + 
  ylab("Kept catch estimate") + 
  xlab("Year") +
  labs(linetype = "Catch Estimate", shape = "Catch Estimate")

```

\clearpage

## Revised Chinook catch estimates before iREC (2005-2012)

### Model fit

When revising Chinook catch estimates prior to the introduction of iREC, the following regional and temporal stratification and time series were used to produce separate models to better represent fishery and catch monitoring dynamics in different regions and seasons.

- Summer models for each of Northern West Coast Vancouver Island (WCVI), Southern WCVI, Northern Strait of Georgia, Southern Strait of Georgia, Juan de Fuca Strait, and Johnstone Strait as six separate models using 2013-2023 time period.
- Spring/Fall models for each of WCVI, Northern Strait of Georgia, Southern Strait of Georgia, Juan de Fuca Strait, and Johnstone Strait as five separate models using 2013-2023 time period.
- Annual models for Central BC, Haida Gwaii (Northern BC AABM), and Northern BC ISBM, as three separate models using 2015 to 2023 time period.

The model selection process determined that separating South Coast into smaller regional models made for a better fit overall. This makes intuitive sense as creel programs in each of the regions are operated in different ways. For example, WCVI may include more extensive log book programs than other regions. Furthermore, as described above, the quality of creel surveys may vary region to region based on geography, logistics, and weather. In Central BC, Haida Gwaii, and Northern BC, there is very limited off-season fisheries and therefore a single annual model was fit for each of these regions. Furthermore, while this process allowed us to find the best fitting model among a number of candidates, residual plots show the summer models overall tend to fit better than the seasonal or annual models, and generally the summer models had excellent residual fit and show minimal patterns in the residuals (see Appendix \@ref(app:second-appendix) for residual plots).

These models develop a relationship between filtered creel + iREC catch in the a given season (spring (months 1:4), summer (months 5:9), or fall (months 10:12)) and the unfiltered creel and logbook only (i.e. no iREC) catch in the creel summer period. This relationship is developed with each data point as a single year in the 2013-2023 time series (or 2015-2023 for NBC and CBC). The relationship is then used to predict the kept and released catch by mark status from 2005 to year before the model fit. For example the Haida Gwaii Annual model is used to predict annual estimates of kept and released catch by mark status from 2005 to 2014 based on the existing creel and logbook catch estimates.

In the following subsections we describe the chosen model and its parameters and show an example of the modeled relationships from the best fitting models for each spatio-temporal combination. The full suite of model outputs can be found in Appendix \@ref(app:third-appendix).

```{r plotting_function, include = FALSE}
#plotting function that will be used for the next few sections
# Helper: runs predict_with_ci + plot_region_model (not by-year) for one region,
# returns a named list of ggplot objects (one per fishery). For main text use.

ModelYearEnd <- 2023
SouthModelYearStart <- 2013
NorthModelYearStart <- 2015
most_recent_data_year <- 2023
PredYearStart <- 2005


main_regions <- list(

  # --- Summer fisheries ---
  "SGS Summer"      = list(model_name = "SGS_summer_model",   x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = "summer",
                            fishery_codes = c("SGS S SUMMER")),
  "NGS Summer"      = list(model_name = "NGS_summer_model",   x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = "summer",
                            fishery_codes = c("NGS S SUMMER")),
  "JNST Summer"     = list(model_name = "jnst_summer_model",  x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = "summer",
                            fishery_codes = c("JOHN ST S SUMMER")),
  "SWCVI Summer"    = list(model_name = "SWCVI_summer_model", x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = "summer",
                            fishery_codes = c("SWCVI S SUMMER AABM", "SWCVI S SUMMER ISBM")),
  "NWCVI Summer"    = list(model_name = "NWCVI_summer_model", x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = "summer",
                            fishery_codes = c("NWCVI S SUMMER AABM", "NWCVI S SUMMER ISBM")),
  "JDF Summer"      = list(model_name = "jdf_summer_model",   x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = "summer",
                            fishery_codes = c("CA JDF S SUMMER")),

  # --- Seasonal (spring/fall) fisheries ---
  "SGS Seasonal"    = list(model_name = "seasonal_SGS_only", x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = c("spring", "fall"),
                            fishery_codes = c("SGS S SPRING", "SGS S FALL")),
  "NGS Seasonal"    = list(model_name = "seasonal_NGS_only", x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = c("spring", "fall"),
                            fishery_codes = c("NGS S SPRING", "NGS S FALL")),
  "JDF Seasonal"    = list(model_name = "seasonal_JDF",      x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = c("spring", "fall"),
                            fishery_codes = c("CA JDF S SPRING", "CA JDF S FALL")),
  "JNST Seasonal"   = list(model_name = "seasonal_JNST",     x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = c("spring", "fall"),
                            fishery_codes = c("JOHN ST S SPRING", "JOHN ST S FALL")),
  "WCVI Seasonal"   = list(model_name = "seasonal_WCVI",     x_var = "creel_plus_summer",
                            x_var_mr = "mr_creel_plus_summer", x_var_rr = "rr_creel_plus_summer",
                            year_start = SouthModelYearStart, season_codes = c("spring", "fall"),
                            fishery_codes = c("NWCVI S SPRING AABM", "SWCVI S SPRING AABM",
                                               "NWCVI S FALL AABM",   "SWCVI S FALL AABM")),

  # --- Northern BC (ISBM/AABM) fisheries ---
  "North AABM"      = list(model_name = "north_aabm",   x_var = "historic_plus",
                            x_var_mr = "mr_historic_plus", x_var_rr = "rr_historic_plus",
                            year_start = NorthModelYearStart, season_codes = "annual",
                            fishery_codes = c("NORTH AABM S")),
  "North ISBM"      = list(model_name = "north_isbm",   x_var = "historic_plus",
                            x_var_mr = "mr_historic_plus", x_var_rr = "rr_historic_plus",
                            year_start = NorthModelYearStart, season_codes = "annual",
                            fishery_codes = c("NORTH ISBM S")),
  "Central BC ISBM kept" = list(model_name = "central_isbm_kept", x_var = "historic_plus",
                            x_var_mr = "mr_historic_plus", x_var_rr = "rr_historic_plus",
                            year_start = NorthModelYearStart, season_codes = "annual",
                            fishery_codes = c("CENTRAL S")), 
   "Central BC ISBM released" = list(model_name = "central_isbm_released", x_var = "historic_plus",
                            x_var_mr = "mr_historic_plus", x_var_rr = "rr_historic_plus",
                            year_start = NorthModelYearStart, season_codes = "annual",
                            fishery_codes = c("CENTRAL S"))
)


run_main_region <- function(model_name, x_var, x_var_mr, x_var_rr, year_start,
                             season_codes, fishery_codes) {

  model <- sport_catch_models$model_list[[model_name]]

  if (is.null(model)) {
    stop("No model found for model_name = '", model_name,
         "'. Available names: ", paste(names(sport_catch_models$model_list), collapse = ", "))
  }

  df <- prepare_region_data(
    data            = sport_catch_finescale,
    combine_to      = NULL,
    fishery_codes   = fishery_codes,
    season_codes    = season_codes,
    status_codes    = c("marked_kept_total", "unmarked_kept_total",
                         "marked_released_total", "unmarked_released_total"),
    pred_year_start = PredYearStart,
    year_start      = year_start,
    year_end        = ModelYearEnd,
    x_var           = x_var,
    x_var_mr        = x_var_mr,
    x_var_rr        = x_var_rr,
    exclude_years   = NULL
  )

  pred_out <- predict_with_ci(
    model, df,
    pred_year_start = PredYearStart,
    year_start      = year_start,
    year_end        = ModelYearEnd,
    x_var           = x_var,
    x_var_mr        = x_var_mr,
    x_var_rr        = x_var_rr
  )

  pred_df   <- pred_out$predictions
  fisheries <- unique(df$finescale_fishery)

  plots <- purrr::map(
    fisheries,
    ~ plot_region_model(
      fishery    = .x,
      x_var      = x_var,
      x_var_mr   = x_var_mr,
      x_var_rr   = x_var_rr,
      plot_data  = df,
      model_data = pred_df,
      year_start = year_start,
      year_end   = ModelYearEnd
    )
  )
  names(plots) <- fisheries
  plots
}

```

### Summer models

#### Northern WCVI

The best model for Northern WCVI Summer fishery catch ($C_{summer}$), was a model with a gaussian distribution that included summer creel ($C_{creel,summer}$), marked status ($m$) and management ($M$), an interaction between mark status and management, and an intercept. Kept status was dropped from the model during model selection, as were more complex variations of interactions. 

$$C_{summer} = C_{creel,summer} + M + m + M:m + 1
\label{eq:nwcvi_summer}$$

This model parameterization allows for filtered creel + calibrated iREC catch in the whole summer period (months 5:9) to be predicted independently by the unfiltered creel and logbook catch, management type (e.g. AABM vs. ISBM), and mark status (e.g. marked vs. unmarked). In addition, the interaction between management and mark status allows the effect of management type on catch to be modified by mark status. 

Figure \@ref(fig:fig-nwcvi-isbm-summer-model) shows ISBM modeled relationship for NWCVI Summer. The AABM counterpart plot can be found in Appendix \@ref(app:third-appendix). Overall, we found a strong linear relationship between summer creel and summer iREC-included catch (p<0.001, pseudo r^2 = 0.98), the intercept (but not slope) of which changed with mark status (overall there were more unmarked fish than marked fish), and the interaction between mark status and management type, such that more unmarked fish were caught during ISBM than AABM.   


```{r results example seasonal summer, results='asis', echo=FALSE}
#| label: fig-nwcvi-isbm-summer-model
#| fig-cap: "Northern WCVI ISBM Summer fishery modelled relationship for between summertime unfiltered creel estimate (months 5 and 9) and total catch estimate that incorporates iREC. Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."

cfg <- main_regions[["NWCVI Summer"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)


#(plots[["NWCVI S SUMMER AABM"]] + ylab("iREC-included catch estimate")) + 
  (plots[["NWCVI S SUMMER ISBM"]] + ylab("iREC-included catch estimate")) 

#+ facet_wrap(~management, scales="fixed")

```

#### Southern WCVI

The best model for Northern WCVI Summer fishery catch ($C_{summer}$), was a model with a gaussian distribution that included summer creel ($C_{creel,summer}$), kept status ($k$) , marked status ($m$) and management ($M$), an interaction between kept status and summer creel, an interaction between mark status and summer creel, and an intercept. More complex variations and interactions were dropped during model selection.

$$C_{summer} = C_{creel,summer} + M + m + k + C_{creel,summer}:k + C_{creel,summer}:m + 1
\label{eq:swcvi_summer}$$

This model parameterization allows for filtered creel + calibrated iREC catch in the whole summer period (months 5:9) to be predicted independently by the unfiltered creel and logbook catch, management type (e.g. AABM vs. ISBM), mark status (e.g. marked vs. unmarked), and kept status (kept vs. released). In addition, the interactions allow the effect of creel summer on catch (i.e. the slope) to be modified by kept status and mark status separately. 

Figure \@ref(fig:fig-swcvi-aabm-summer-model) shows the AABM modeled relationship for SWCVI Summer. The ISBM counterpart plot can be found in Appendix \@ref(app:third-appendix). the Overall, we found a strong positive linear relationship between summer creel and summer iREC-included catch (p<0.001, pseudo r^2 = 0.98), the slope of which was modestly modified by kept status (released steeper than kept) and marked status (unmarked steeper than marked). No significant independent effects were found for kept or mark status. There was significant effect of management (AABM caught more fish overall than ISBM), and AABM appeared to have a tighter relationship than ISBM. 


```{r results example seasonal summer, results='asis', echo=FALSE}
#| label: fig-swcvi-aabm-summer-model
#| fig-cap: "Southern WCVI AABM Summer fishery modelled relationship for between summertime unfiltered creel estimate (months 5 and 9) and total catch estimate that incorporates iREC. Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."

cfg <- main_regions[["SWCVI Summer"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)


(plots[["SWCVI S SUMMER AABM"]] + ylab("iREC-included catch estimate"))
  #+ facet_wrap(~management, scales="fixed")) + (plots[["SWCVI S SUMMER ISBM"]] + ylab("iREC-included catch estimate")) + facet_wrap(~management, scales="fixed")

```

#### Northern Georgia Strait 
The best model for Northern Georgia Strait summer fishery catch ($C_{summer}$), was a model with a gaussian distribution that included summer creel ($C_{creel,summer}$) and mark status ($m$) and an intercept. Kept status was dropped from the model during model selection, as were more complex variations and interactions.

$$C_{summer} = C_{creel,summer} + m + 1
\label{eq:ngs_summer}$$

This model parameterization allows for filtered creel + calibrated iREC catch in the whole summer period (months 5:9) to be predicted independently by the unfiltered creel and logbook catch and mark status (e.g. marked vs. unmarked).  

Figure \@ref(fig:fig-ngs-summer-model) shows the modeled relationship for NGS Summer. Overall, we found a strong positive linear relationship between summer creel and summer iREC-included catch (p<0.001, pseudo r^2 = 0.94). There were more unmarked fish than marked fish caught. 

```{r results example seasonal summer, results='asis', echo=FALSE}
#| label: fig-ngs-summer-model
#| fig-cap: "Northern Strait of Georgia fishery modelled relationship between summertime unfiltered creel estimate and total catch estimate including iREC in summer months (5-9). Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."
cfg <- main_regions[["NGS Summer"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plot <- plots[[1]] + ylab("iREC-included catch estimate")
plot
```


#### Southern Georgia Strait

The best model for Southern Georgia Strait summer fishery catch ($C_{summer}$), was also a model with a gaussian distribution that included summer creel ($C_{creel,summer}$) and mark status ($m$) and an intercept. Kept status was dropped from the model during model selection, as were more complex variations and interactions.

$$C_{summer} = C_{creel,summer} + m + 1
\label{eq:sgs_summer}$$

This model parameterization allows for filtered creel + calibrated iREC catch in the whole summer period (months 5:9) to be predicted independently by the unfiltered creel and logbook catch and mark status (e.g. marked vs. unmarked).  

Figure \@ref(fig:fig-sgs-summer-model) shows the modeled relationship for SGS Summer. Overall, we found a moderate positive linear relationship between summer creel and summer iREC-included catch (p<0.001, pseudo r^2 = 0.71). Overall, there were more unmarked fish than marked fish caught. 

```{r results example seasonal summer, results='asis', echo=FALSE}
#| label: fig-sgs-summer-model
#| fig-cap: "Southern Strait of Georgia fishery modelled relationship between summertime unfiltered creel estimate and total catch estimate including iREC in summer months (5-9). Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."
cfg <- main_regions[["SGS Summer"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plot <- plots[[1]] + ylab("iREC-included catch estimate")
plot
```

#### Johnstone Strait

The best model for Johnstone Strait summer fishery catch ($C_{summer}$), was a model with a Gamma distribution that included summer creel ($C_{creel,summer}$), kept_status ($k$), mark status ($m$), an interaction between kept status and summer creel, an interaction between mark status and summer creel, and an intercept. More complex variations and interactions were dropped. 

$$C_{summer} = C_{creel,summer} + m + k + C_{creel,summer}:k + C_{creel,summer}:m + 1
\label{eq:jnst_summer}$$

This model parameterization allows for filtered creel + calibrated iREC catch in the whole summer period (months 5:9) to be predicted independently by the unfiltered creel and logbook catch, mark status (e.g. marked vs. unmarked), and kept status (kept vs. released). In addition, the interactions allow the effect of creel summer on catch (i.e. the slope) to be modified by kept status and mark status separately.  

Figure \@ref(fig:fig-jnst-summer-model) shows the modeled relationship for JNST Summer. Overall, we found a strong non-linear positive relationship between summer creel and summer iREC-included catch (p<0.001, pseudo r^2 = 0.96), the slope of which was modestly modified by kept status (released steeper than kept) and marked status (marked steeper than unmarked). There was an overall effect of mark status, as above, more unmarked fish were encountered than marked fish, and an overall effect of kept_status, more fish kept than released. 

```{r results example seasonal summer, results='asis', echo=FALSE}
#| label: fig-jnst-summer-model
#| fig-cap: "Johnstone Strait fishery modelled relationship between summertime unfiltered creel estimate and total catch estimate including iREC in summer months (5-9). Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."
cfg <- main_regions[["JNST Summer"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

 plots[[1]] + ylab("iREC-included catch estimate") 

```



#### Juan de Fuca

The best model for Juan de Fuca Strait summer fishery catch ($C_{summer}$), was a model with a Gamma distribution that included summer creel ($C_{creel,summer}$), kept_status ($k$), mark status ($m$), an interaction between kept status and summer creel, an interaction between mark status and summer creel, and an intercept. More complex variations and interactions were dropped. 

$$C_{summer} = C_{creel,summer} + m + k + C_{creel,summer}:k + C_{creel,summer}:m + 1
\label{eq:jdf_summer}$$

This model parameterization allows for filtered creel + calibrated iREC catch in the whole summer period (months 5:9) to be predicted independently by the unfiltered creel and logbook catch, mark status (e.g. marked vs. unmarked), and kept status (kept vs. released). In addition, the interactions allow the effect of creel summer on catch (i.e. the slope) to be modified by kept status and mark status separately.  

Figure \@ref(fig:fig-jdf-summer-model) shows the modeled relationship for JDF Summer. Overall, we found a strong non-linear positive relationship between summer creel and summer iREC-included catch (p<0.001, pseudo r^2 = 0.91), the slope of which was modestly modified by kept status (released steeper than kept) and marked status (marked steeper than unmarked). There was an overall effect of mark status, as above, more unmarked fish were encountered than marked fish, and an overall effect of kept_status, more fish kept than released. 


```{r results example seasonal summer, results='asis', echo=FALSE}
#| label: fig-jdf-summer-model
#| fig-cap: "Juan de Fuca Summer fishery modelled relationship between unfiltered creel and total catch estimates including iREC in summer months (5-9). Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."

cfg <- main_regions[["JDF Summer"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plots[[1]] + ylab("iREC-included catch estimate") 


```

### Spring and Fall Models

#### WCVI
The best model for WCVI Spring and Fall AABM fishery catch ($C_{seasonal}$), was a model with a gaussian distribution that included region ($r$), season ($s$), their interaction, and an intercept. Notably, summer creel, marked status, and kept status were dropped from the model during model selection, as were more complex variations of interactions. 

$$C_{seasonal} = r + s + r:s + 1
\label{eq:wcvi_seasonal}$$

This model parameterization allows for filtered creel + calibrated iREC catch in the off-season to be predicted independently by region (NWCVI vs. SWCVI) and season (spring vs. fall), and for the effect of season to be modified by region.

Figure \@ref(fig:fig-swcvi-aabm-spring-model) shows the SWCVI AABM modeled relationship for Spring. The Fall and NWCVI counterpart plots can be found in Appendix \@ref(app:third-appendix). Overall, we found no relationship between summer creel and off-season iREC-included catch. There was a significant interaction such that SWCVI Spring had more catch than the other regions and season combinations, but no overall main effect of region or season. This flat relationship explains minimal variation in the data (pseudo r^2 = 0.18), but gives an offset (intercept) to use as a predictor for off-season catch. Unlike the summer models for NWCVI and SWCVI, which both demonstrated overall more unmarked catch than marked catch, mark status was not included in the model, indicating similar catch for marked and unmarked fish in the off-season. 


```{r results example seasonal summer, results='asis', echo=FALSE}
#| label: fig-swcvi-aabm-spring-model
#| fig-cap: "Southern WCVI AABM Summer fishery modelled relationship for between summertime unfiltered creel estimate (months 5 and 9) and total catch estimate that incorporates iREC. Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."

cfg <- main_regions[["WCVI Seasonal"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)


  (plots[["SWCVI S SPRING AABM"]] + ylab("iREC-included catch estimate")) 

#+ facet_wrap(~management, scales="fixed")

```
#### Northern Georgia Strait 
The best model for Northern Georgia Strait off-season fishery catch ($C_{seasonal}$), was a model with a gaussian distribution that included summer creel ($C_{creel,summer}$), kept status ($k$), season ($s$), an interaction between summer creel and season, an interaction between summer creel and kept status, and an intercept. Mark status and more complex variations and interactions were dropped during model selection.

$$C_{seasonal} = C_{creel,summer} + k + s + C_{creel,summer}:s + C_{creel,summer}:k + 1
\label{eq:ngs_seasonal}$$

This model parameterization allows for off-season filtered creel + calibrated iREC catch to be predicted independently by the unfiltered creel and logbook summer catch, kept status (kept vs. released), mark status (e.g. marked vs. unmarked), and season (spring vs. fall). Furthermore, this model formula allows for the effect of summer creel on off-season catch to be modified by season and kept status.   

Figure \@ref(fig:fig-ngs-spring-model) shows the modeled relationship for NGS Spring. The Fall counterpart plot can be found in Appendix \@ref(app:third-appendix). Overall, we found a moderate positive linear relationship between summer creel and off-season iREC-included catch (p<0.001, pseudo r^2 = 0.57). There was an overall effect of season (more catch in spring than fall), and season altered the slope of the summer creel to seasonal iREC such that the slope is steeper for spring than for fall catch. There was no overall effect of kept status, but kept status modified the slope of the creel summer catch to iREC seasonal catch: the slope is steeper for released than kept fish. Unlike the summer model for NGS, which demonstrated overall more unmarked catch than marked catch, mark status was dropped from this model, indicating similar catch for marked and unmarked fish in the off-season. 

```{r results example seasonal summer, results='asis', echo=FALSE}
#| label: fig-ngs-spring-model
#| fig-cap: "Northern Strait of Georgia fishery modelled relationship between summertime unfiltered creel estimate and total catch estimate including iREC in spring (months 1-4). Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."
cfg <- main_regions[["NGS Seasonal"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plots[["NGS S SPRING"]] + ylab("iREC-included catch estimate")
```
#### Southern Georgia Strait 
The best model for Southern Georgia Strait off-season fishery catch ($C_{seasonal}$), was a model with a gaussian distribution that included summer creel ($C_{creel,summer}$), mark status ($m$), season ($s$), an interaction between summer creel and mark status, and an intercept. Kept status and more complex variations and interactions were dropped during model selection.

$$C_{seasonal} = C_{creel,summer} + m + s + C_{creel,summer}:m + 1
\label{eq:sgs_seasonal}$$

This model parameterization allows for off-season filtered creel + calibrated iREC catch to be predicted independently by the unfiltered creel and logbook summer catch, mark status (e.g. marked vs. unmarked), and season (spring vs. fall). Furthermore, this model formula allows for the effect of summer creel on off-season catch to be modified by mark status.    

Figure \@ref(fig:fig-sgs-spring-model) shows the modeled relationship for SGS Spring. The Fall counterpart plot can be found in Appendix \@ref(app:third-appendix). Overall, we found a weak positive linear relationship between summer creel and off-season iREC-included catch (p<0.011, pseudo r^2 = 0.39). There was an overall effect of season (more catch in spring than fall). Similar to the summer SGS model, there was an overall effect of mark status (more unmarked than marked catch). However, in this model mark status modified the slope of the creel summer catch to iREC seasonal catch: the slope is steeper for marked than unmarked fish. 

```{r results example seasonal summer, results='asis', echo=FALSE}
#| label: fig-sgs-spring-model
#| fig-cap: "Southern Strait of Georgia fishery modelled relationship between summertime unfiltered creel estimate and total catch estimate including iREC in spring (months 1-4). Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."
cfg <- main_regions[["SGS Seasonal"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plots[["SGS S SPRING"]] + ylab("iREC-included catch estimate")
```


#### Johnstone Strait

The best model for Johnstone Strait seasonal fishery catch ($C_{seasonal}$), was a model with a gaussian distribution that included mark status ($m$) and an intercept. Notably, summer creel, kept status, season, and any more complex variations and interactions were dropped. 

$$C_{seasonal} = m + 1
\label{eq:jnst_seasonal}$$

This model parameterization allows for filtered creel + calibrated iREC catch in the off-season to be predicted independently by mark status (e.g. marked vs. unmarked). 

Figure \@ref(fig:fig-jnst-spring-model) shows the modeled relationship for JNST Spring. The Fall counterpart plot can be found in Appendix \@ref(app:third-appendix). Overall, we found no relationship between summer creel and off-season iREC-included catch. As in the summer model for JNST, there was a significant effect of mark status, more unmarked fish caught than marked fish. The flat relationship explains minimal variation in the data (pseudo r^2 = 0.08), but gives an offset (intercept) to use as a predictor for off-season catch. 

```{r results example seasonal summer, results='asis', echo=FALSE}
#| label: fig-jnst-spring-model
#| fig-cap: "Johnstone Strait fishery modelled relationship between summertime unfiltered creel estimate and total catch estimate including iREC in spring months (1:4). Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."
cfg <- main_regions[["JNST Seasonal"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

 plots[["JOHN ST S SPRING"]] + ylab("iREC-included catch estimate") 

```


#### Juan de Fuca 

The best model for Juan de Fuca Strait off-season fishery catch ($C_{seasonal}$), was a model with a Gamma distribution that included summer creel ($C_{creel,summer}$), kept_status ($k$), mark status ($m$), season($s$), an interaction between mark status and summer creel, and an intercept. More complex variations and interactions were dropped. 

$$C_{seasonal} = C_{creel,summer} + m + k + s + C_{creel,summer}:m + 1
\label{eq:jdf_seasonal}$$

This model parameterization allows for filtered creel + calibrated iREC catch in the off-season to be predicted independently by the summertime unfiltered creel and logbook catch, mark status (e.g. marked vs. unmarked), and kept status (kept vs. released). In addition, the interaction allows the effect of creel summer on off-season catch (i.e. the slope) to be modified by mark status.  

Figure \@ref(fig:fig-jdf-spring-model) shows the modeled relationship for JDF Spring. The Fall counterpart plot can be found in Appendix \@ref(app:third-appendix). Overall, we found a modest non-linear positive relationship between summer creel and seasonal iREC-included catch (p<0.003, pseudo r^2 = 0.43), the slope of which was modified by marked status (marked steeper than unmarked), similar to the summer JDF model. Although there was no overall effect of mark status, the trend was in the opposite direction to the summer JDF model, slightly more marked fish than unmarked fish caught in the off-season (p=0.08). There was an overall effect of kept_status, slightly more fish released than kept (p<0.003), the opposite direction to the summer JDF model. Finally, there was an overall effect of season, where more fish were caught in the spring than the fall.  


```{r results example seasonal summer, results='asis', echo=FALSE}
#| label: fig-jdf-spring-model
#| fig-cap: "Juan de Fuca Spring fishery modelled relationship between unfiltered creel and total catch estimates including iREC in spring months (1-4). Data points are yearly estimates from 2013 to 2023. Note varying scales on the y-axes to better visualize the model fit."

cfg <- main_regions[["JDF Seasonal"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plots[["CA JDF S SPRING"]] + ylab("iREC-included catch estimate") 


```


###  Annual Fisheries

#### Haida Gwaii
The best model for the Haida Gwaii annual fishery incorporates the historic data (Haida creel + logbook, $C_{historic}$), kept status ($k$), mark status ($m$), and the interaction between kept and mark status, and an intercept, as identified in Equation \eqref{eq:haida_gwaii_annual}. Unlike other model fits, we used a gaussian link instead of a Gamma log because earlier values in the time series were outside the range of the model fit. As identified in Appendix \@ref(apdx-haida-model), the Gamma log link would cause the model to extrapolate unmarked kept catch estimates more the 3 times the historic catch. This is was inconsistent with the much lower increase across the rest of the time series and for other catch statuses.

$$C_{annual} = C_{historic} + k + m + k:m + 1 
\label{eq:haida_gwaii_annual}$$

This model parameterization allows for calibrated iREC catch to be predicted by the unfiltered creel and logbook (historic) catch in the limited creel summer period. In addition, the effect of the creel and logbook (historic) summer period on catch is modified by kept status and mark status. The interaction between kept and mark status allows the effect of kept status to be moderated by mark status. Figure \@ref(fig:fig-nbcaabm-residuals) shows the qq plot residuals and a plot of the residuals against predicted values to detect patterns in residuals.

Figure \@ref(fig:fig-nbcaabm-model) shows the model results with Haida Gwaii. Overall, we found a strong positive linear relationship between annual Haida creel + logbook and annual iREC catch (p<0.001, pseudo r^2 = 0.91). There were no overall significant effects of either mark or kept status, but a trending effect of mark status (p=0.056), which suggested that there were more unmarked fish encountered than marked fish. Furthermore, the interaction suggests that for unmarked fish, more fish were released than kept, while there were similar numbers of kept and released fish in the marked category. 


```{r results NBC AABM Annual, results='asis', echo=FALSE}
#| label: fig-nbcaabm-model
#| fig-cap: "Haida Gwaii annual fishery modelled relationship between historic unfiltered creel estimate and total iREC catch estimates in the whole year. Data points are yearly estimates from 2015 to 2023. Note varying scales on the y-axes to better visualize the model fit."

cfg <- main_regions[["North AABM"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plots[[1]] + ylab("iREC-only catch estimate")

```

#### Northern BC ISBM Fisheries


The best model for annual Northern BC (NBC) ISBM fisheries was one with a Gamma distribution using a log link that included annual historic data (Area 3 and 4 Creel + logbook, $C_{historic,annual}$), mark status ($m$), and an interaction between historic data and mark status. Kept status and more complex variations and interactions were dropped.  

$$C_{annual} = C_{historic,annual} + m + C_{historic,annual}:m \label{eq:nbc_isbm_annual}$$

This model parameterization allows for annual calibrated iREC catch to be predicted both independently by the annual unfiltered Area 3 and 4 creel and logbook catch, by mark status and their interaction. Figure \@ref(fig:fig-nbcisbm-residuals) shows the qq plot residuals and a plot of the residuals against predicted values to detect patterns in residuals. The residuals vs. predicted values show no significant patterns.

Figure \@ref(fig:fig-nbcisbm-model) shows the annual model results in NBC ISBM. There is a positive non-linear relationship between creel and logbook data and iREC annual catch estimates (p<0.0001, pseudo r^2 = 0.91). The slope of this relationship was slightly steeper for marked fish than unmarked fish. There was a significant effect of mark status, such that overall the marked catch was substantially less (\< 1,500 pieces) than the catch of unmarked fish (\< 15,000 pieces). 

```{r results NBC ISBM annual, results='asis', echo=FALSE}
#| label: fig-nbcisbm-model
#| fig-cap: "NBC ISBM fishery modelled relationship between unfiltered creel estimate and iREC catch estimates over the whole year. Data points are yearly regional estimates from 2015 to 2023. Note varying scales on the y-axes to better visualize the model fit."


cfg <- main_regions[["North ISBM"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plots[[1]] + ylab("iREC-included catch estimate")
#+ facet_wrap(~mark_status, scales="fixed")



```

=======

#### Central BC annual fishery

Central BC was split into two models, one for kept fish and one for released fish where there were almost exclusively zeros provided in the historical data. 

The best model for the kept Central BC annual catch was one with a Gamma distribution with a log link which included historic estimates (logbook only) and mark status. 

$$ C_{irec,annual,kept} = C_{historic,annual} + m \label{eq:cbc_annual_kept}$$

This model parameterization allows for calibrated iREC catch to be predicted independently by the annual logbook catch and by mark status. Figure \@ref(fig:fig-cbc-residuals) shows the qq plot residuals and a plot of the residuals against predicted values to detect patterns in residuals. There were minimal patterns in the residuals.


The best model for the released Central BC annual catch was one with a Gamma distribution with a log link which included mark status and an intercept. As historic logbook estimates were mostly zero, this factor was excluded during model selection, as was kept status and various interactions. 

$$ C_{irec,annual,released} = m + 1 \label{eq:cbc_annual_released}$$

This model parameterization allows for calibrated iREC catch to be predicted independently by mark status. Figure \@ref(fig:fig-cbc-residuals) shows the qq plot residuals and a plot of the residuals against predicted values to detect patterns in residuals. There were minimal patterns in the residuals.

Figure \@ref(fig:fig-cbc-model) shows the combined CBC model results. Generally for low catch estimates a flat line is predicted, but for unmarked Kept fish, the iREC estimate is an increasing function of summertime logbook catch. As released catch is not captured well by the log book program, the modelling of release data before iREC is likely to not be that informative. In both models there was an effect of mark status such that there were more unmarked fish encountered than marked fish. There was an overall positive non-linear relationship between kept logbook catch and iREC (p<0.01, pseudo r^2=0.90). 



```{r results CBC model, results='asis', echo=FALSE}
#| label: fig-cbc-model
#| fig-cap: "CBC annual fishery modelled relationship between annual logbook estimates and annual iREC catch estimates. Data points are yearly estimates from 2015 to 2023. Note varying scales on the y-axes to better visualize the model fit."

cfg <- main_regions[["Central BC ISBM kept"]]

plots <- run_main_region(
  model_name    = cfg$model_name,
  x_var         = cfg$x_var,
  x_var_mr      = cfg$x_var_mr,
  x_var_rr      = cfg$x_var_rr,
  year_start    = cfg$year_start,
  season_codes  = cfg$season_codes,
  fishery_codes = cfg$fishery_codes
)

plot <- plots[[1]] + ylab("annual iREC catch estimate") + xlab("annual logbook catch estimate")
plot

```

\clearpage

## Validation using sport head recoveries

```{r shrp-head-catch-model-fits, echo=FALSE}

lm_eqn <- function(df){
  m <- lm(head_total ~ catch_estimate_predicted + 0, df)
  
  eq <- substitute(italic(y) == a %.% italic(x)*", Adj."~~italic(R)^2~"="~r2*"," ~~italic(n)*"="~df_n,
                   list(a = format(unname(coef(m)[1]), digits = 3),
                        b = format(unname(coef(m)[2]), digits = 3),
                        r2 = format(summary(m)$adj.r.squared, digits = 3),
                        df_n = nrow(df)))
  as.character(as.expression(eq))
}

lm_eqn_vec <- function(head_total, catch_estimate){
  m <- lm(head_total ~ catch_estimate + 0)
  
  eq <- substitute(italic(y) == a %.% italic(x)*", Adj."~~italic(R)^2~"="~r2*"," ~~italic(n)*"="~df_n,
                   list(a = format(unname(coef(m)[1]), digits = 3),
                        b = format(unname(coef(m)[2]), digits = 3),
                        r2 = format(summary(m)$adj.r.squared, digits = 3),
                        df_n = length(head_total)))
  as.character(as.expression(eq))
}

lm_log_eqn_vec <- function(head_total, catch_estimate){
  m <- lm(log(head_total) ~ log(catch_estimate) + 0)
  
  eq <- substitute(italic(ln(y)) == a %.% italic(ln(x))*", Adj."~~italic(R)^2~"="~r2*"," ~~italic(n)*"="~df_n,
                   list(a = format(unname(coef(m)[1]), digits = 3),
                        b = format(unname(coef(m)[2]), digits = 3),
                        r2 = format(summary(m)$adj.r.squared, digits = 3),
                        df_n = length(head_total)))
  as.character(as.expression(eq))
}

creel_to_heads <-
  sport_catch_models$model_catch |>
  filter(status == "marked_kept_total") |> 
  full_join(pacRecCatch::get_head_totals(), by = c("year", finescale_fishery = "region")) |>
  mutate(finescale_fishery = if_else(grepl("NWCVI|SWCVI", finescale_fishery), 
                                     trimws(gsub(" ISBM| AABM", "", finescale_fishery)),
                                     finescale_fishery))  |>
  group_by(year, finescale_fishery) |>
  select(year, finescale_fishery, creel_unfiltered_plus, head_total, catch_estimate_predicted, catch_estimate) |>
  summarize(across(c(head_total, catch_estimate_predicted, creel_unfiltered_plus, catch_estimate),
                   \(x) sum(x, na.rm=TRUE)),
            .groups = "drop") |>
  filter(head_total > 0, head_total < 6000)


south_creel_to_heads <- 
  creel_to_heads |>
  filter(!grepl("^NORTH ", finescale_fishery), !grepl("^CENTRAL " ,finescale_fishery))

all_heads_lm <- lm(head_total ~ catch_estimate_predicted + 0, creel_to_heads)


pre_irec_catch_model_df <- 
  creel_to_heads |>
  filter(year < 2013)

pre_irec_lm <- lm(head_total ~ catch_estimate_predicted + 0, pre_irec_catch_model_df)


post_irec_catch_model_df <- 
  creel_to_heads |>
  filter(year >= 2013)

post_irec_lm <- lm(head_total ~ catch_estimate_predicted + 0, post_irec_catch_model_df)


south_creel_to_heads_df <-
  south_creel_to_heads |>
  filter(creel_unfiltered_plus > 0)

south_creel_to_heads_lm <- lm(head_total ~ creel_unfiltered_plus + 0, south_creel_to_heads_df)

south_combine_to_heads_df <-
  south_creel_to_heads |>
  filter(catch_estimate_predicted > 0)

combine_creel_to_heads_lm <- lm(head_total ~ catch_estimate_predicted + 0, south_combine_to_heads_df)
```

As the total number of heads collected by the SHRP is expected to be primarily driven by the the total kept marked catch in recreational fisheries, the total kept marked catch and total heads collected by the SHRP should be correlated. In contrast to the creel program, the SHRP relies on several sampling approaches (e.g., lodge sampling, public collection points, avid anglers) and should not necessarily be driven by creel survey coverage. To better align with the CWT estimation methods, the catch and heads where spatially group by the region defined in Figure \@ref(fig:fig-CwtMap) and the seasonal period. To provide a clearer comparison, a single observation that consisted of over 6000 head recoveries from the SWVI summer period in 2010. The majority of the heads originate from a single depot that seemed to have contributed substantially more samples compared to other years.

In Figure \@ref(fig:fig-heads-to-creel-and-updated-catch), the kept mark Chinook catch in Southern BC recreational fisheries (i.e. excluding Northern and Central regions) are compared to just Creel estimated mark kept catch (left panel) and catch that combines creel and calibrated iREC (right panel). As the adjusted $R^2$ show, the combined catch explains slightly more of the variability in the number of heads submitted to the SHRP. Also, the lower slope value of `r format(unname(coef(combine_creel_to_heads_lm)[1]), digits = 3)` against the combined catch, compared to the slope of `r format(unname(coef(south_creel_to_heads_lm)[1]), digits = 3)` with the just the creel catch, further supports the concern that just creel catch provides an under estimate of catch that could impact assessment of CWT indicators. Additionally, the combined creel and iREC catch was able to directly estimate more regions and seasons with `r nrow(south_combine_to_heads_df)` estimates, compared to `r nrow(south_creel_to_heads_df)` estimates with creel only.

The SHRP has generally run consistently over the transition time between the modelled catch (pre-2013) and subsequent catch period during which iREC has been incorporated into estimates. So, the linear relationship between marked kept catch and head recoveries from both the 2005 - 2012 modelled catch time series and the 2013 - 2023 combined creel and iREC catch time series should be similar.

```{r heads-to-creel, echo=FALSE}
#| label: fig-heads-to-creel-and-updated-catch
#| fig-cap: "Relationship between the marked kept catch estimates from the the creel program (left panel) and the combined creel and iREC catch estimates (right panel) compared to the Chinook head recoveries provided to the SHRP"

creel_to_heads_graph <-
  south_creel_to_heads_df |>
  select(year, finescale_fishery, head_total, creel_unfiltered_plus) |>
  ggplot(aes(x=creel_unfiltered_plus, y=head_total)) +
  geom_point(size=1.5) +
  geom_smooth(method='lm', formula= y~x + 0) +
  geom_text(x = 1, 
            y = max(south_creel_to_heads_df$head_total) * 0.9, 
            label = lm_eqn_vec(south_creel_to_heads_df$head_total,
                               south_creel_to_heads_df$creel_unfiltered_plus), 
            parse = TRUE, hjust=0, vjust=0) +
  labs(title = "Southern BC - Creel Only",
       x = "Chinook Marked Kept Creel Catch",
       y = "Chinook Head Recoveries") +
  theme_bw()

combine_catch_to_heads <-
  south_combine_to_heads_df |>
  select(year, finescale_fishery, head_total, catch_estimate_predicted) |>
  ggplot(aes(x=catch_estimate_predicted, y=head_total)) +
  geom_point(size=1.5) +
  geom_smooth(method='lm', formula= y~x + 0) +
  geom_text(x = 1, 
            y = max(south_combine_to_heads_df$head_total) * 0.9,
            label = lm_eqn_vec(south_combine_to_heads_df$head_total,
                               south_combine_to_heads_df$catch_estimate_predicted), 
            parse = TRUE, hjust=0, vjust=0) +
  labs(title = "Southern BC - Creel and iREC Combined",
       x = "Chinook Marked Kept Creel Catch",
       y = "Chinook Head Recoveries") +
  theme_bw()  

creel_to_heads_graph + combine_catch_to_heads
```

Figure \@ref(fig:fig-heads-kept-current) shows the linear relationship between Chinook marked kept catch and heads submitted to the SHRP from 2005 to 2023. As the adjusted $R^2$ of `r round(summary(all_heads_lm)$adj.r.squared, digits = 3)` suggests, there is a strong relationship between the marked kept catch across the years of this analysis with an average sample rate of `r format(unname(coef(all_heads_lm)[1]), digits = 3)` for all Pacific Recreational fisheries.

```{r results Head to Kept Mark Catch, echo=FALSE}
#| label: fig-heads-kept-current
#| fig-cap: "Chinook heads to marked kept catch based on combined creel and iREC.  Each point is a yearly CWT Region based on the seasonal or annual time periods in the southern or nothern portion of BC (i.e., NBC and CBC), respectively, from 2005 to 2023."

creel_to_heads |>
  ggplot(aes(x=catch_estimate_predicted, y=head_total)) +
  geom_point(size=2) +
  geom_smooth(method='lm', formula= y~x + 0) +
  geom_text(x = 1, 
            y = max(creel_to_heads$head_total) * 0.9,
            label = lm_eqn_vec(creel_to_heads$head_total, 
                               creel_to_heads$catch_estimate_predicted), 
            parse = TRUE, 
            hjust=0, 
            vjust=0) +
  labs(title = "Chinook Head to Marked Kept Catch",
       x = "Chinook Marked Kept Catch",
       y = "Chinook Head Recoveries") +
  theme_bw()

```

When comparing the model fit before iREC (Figure \@ref(fig:fig-heads-kept-irec) left panel) and with iREC (Figure \@ref(fig:fig-heads-kept-irec right) right panel), CWT sample rates are fairly consistent with a before iREC average sample rate (i.e. slope) of `r format(unname(coef(pre_irec_lm)[1]), digits = 3)` and an iREC period sample rate of `r format(unname(coef(post_irec_lm)[1]), digits = 3)` with similar adjusted $R^2$ values. This suggests that there is no consistent bias in the modelled marked kept catch estimates, when compared to the estimates that include iREC. The time series of sample rates (i.e. submission rates) for each region is shown in Figure \@ref(fig:fig-fishery-rates).

```{r results Pre iRec Head to Kept Mark Catch, echo=FALSE}
#| label: fig-heads-kept-irec
#| fig-cap: "Chinook heads to marked kept catch based on combined creel and iREC. Each point is a yearly CWT Region based on the seasonal or annual time period in the southern or nothern portions of BC (i.e., NBC and CBC), respectively.  The left panel shows the relationship from 2005 to 2012 and the right panel shows the relationship from 2013 to 2023"

pre_irec <-
  ggplot(data = pre_irec_catch_model_df,
         aes(x=catch_estimate_predicted, y=head_total)) +
  geom_point(size=2) +
  geom_smooth(method='lm', formula= y~x + 0) +
  geom_text(x = 500, 
            y = max(pre_irec_catch_model_df$head_total, na.rm = TRUE) * 0.9, 
            label = lm_eqn_vec(pre_irec_catch_model_df$head_total, 
                               pre_irec_catch_model_df$catch_estimate_predicted), 
            parse = TRUE, 
            hjust=0, 
            vjust=0) +
  labs( title = "All Regions - Before iREC",
        x = "Chinook Marked Kept Catch",
        y = "Chinook Head Recoveries")+
  theme_bw()

post_irec <- ggplot(data = post_irec_catch_model_df,
                    aes(x=catch_estimate_predicted, y=head_total)) +
  geom_point(size=2) +
  geom_smooth(method='lm', formula= y~x + 0) +
  geom_text(x = 500, 
            y =  max(post_irec_catch_model_df$head_total, na.rm= TRUE) * 0.9, 
            label = lm_eqn_vec(post_irec_catch_model_df$head_total, 
                               post_irec_catch_model_df$catch_estimate_predicted), 
            parse = TRUE, 
            hjust=0, 
            vjust=0) +
  labs(title = "All Regions - Combined iREC and Creel",
       x = "Chinook Marked Kept Catch",
       y = "Chinook Head Recoveries") +
  theme_bw()

pre_irec + post_irec
```

````{=tex}
\newpage
\begin{landscape}

```{r results Fishery Rates, echo=FALSE, out.width="100%"}
#| label: fig-fishery-rates
#| fig-cap: "Chinook Head submission rate time series for each CWT region and associated time periods (i.e. season or annual)"

region_labels <- c(
  "CA JDF" = "Juan de Fuca Sport",
  "CENTRAL S" = "Central BC Sport",
  "JOHN ST" = "Johnstone Strait Sport",
  "NORTH AABM" = "Haida Gwaii Sport",
  "NORTH ISBM" = "Nothern BC ISBM Sport",
  "NGS S" = "Georgia Strait Sport (North)",
  "NWCVI S" = "West Coast Vancouver Island Sport (North)",
  "SGS S" = "Georgia Strait Sport (South)",
  "SWCVI S" = "West Coast Vancouver Island Sport (South)"
)

region_df <- 
  creel_to_heads |>
  mutate(submit_rate = head_total / catch_estimate_predicted) |>
  filter(catch_estimate_predicted > 0, catch_estimate_predicted > head_total) |>
  #extract the first + second word in the fishery column so I can facet by "general" region
  mutate(region_ish = word(finescale_fishery, 1, 2)) %>%
  #season is the last word
  mutate(season = factor(word(finescale_fishery, -1), levels=c("SPRING", "SUMMER", "FALL", "S")),
         region_ish = factor(word(finescale_fishery, 1, 2), 
                             levels=c("NORTH AABM", "NORTH ISBM", 
                                      "CENTRAL S", "JOHN ST",
                                      "NWCVI S", "SWCVI S",
                                      "NGS S", "SGS S",
                                      "CA JDF")))

ggplot(data = region_df,
       aes(x=year, y=submit_rate, color=season)) +
  geom_line(linewidth = 1,
            alpha = 0.6) +
  geom_point(size = 1)+
  facet_wrap(~region_ish,
             ncol = 2,
             labeller = labeller(region_ish = region_labels)) +
  scale_color_manual(labels=c("SPRING" = "Spring",
                              "SUMMER"="Summer", 
                              "FALL"="Fall",
                              "S" = "Annual"),
                     values =  c("#440154", "#d4be02", "#35b779", "#31688e"
                     )) +
  labs(x = "Year", 
       y = "Chinook Head Submission Rate", 
       color = "Season")+
  theme_bw() +
  theme(legend.position="bottom") 


```

\end{landscape}
\clearpage
````

## Revised catch data in CTC processes

Revised recreational catch estimates are inputs to the CTC's analyses and reporting processes in three primary ways.

- Reported total kept and released catch
- Marked kept catch estimates are used to estimate CWT recoveries
- Marked and unmarked release rates are used by the ERA to estimate fishery-related incidental mortality

Recreational catch estimate revisions from this analysis were first incorporated into the February 2025 CTC ERA [@CTC2026]. Upon further review of the methods and implementation of the `pacRecCatch` R package for this technical report, some improvements and corrections were identified. While report focuses on the results of latest version, the changes to methods and data are summarized in Appendix \@ref(apdx-analytic-improvements).

### Reporting Kept and Released Catch

As noted previously, revised estimates derived using the methods described in this report were incorporated into CTC reports starting with the February 2025 ERA. Both kept and released catch estimates are typically reported in the Catch and Escapement report by ERA fishery. Here we plot the difference between the revised kept (Figure \@ref(fig:fig-cande-kept)) and released (Figure \@ref(fig:fig-cande-released)) catch estimates, and those recently reported in the 2024 Catch and Escapement report. Again, we see that the revised catch estimates were generally higher than previously reported estimates that did not incorporate iREC information. In some fisheries, iREC information has been included in recent years (e.g., since 2018 in Strait of Georgia Sport), but our revised estimates incorporate iREC in a standardized way across all fisheries. The largest changes in kept catch occurred in CBC where iREC had not been previously incorporated, but the inclusion of iREC more than doubles recent estimates. Other large changes occur in Strait of Georgia sport before 2012 (iREC has already been incorporated from 2013 onward).

```{r results CandEkept, results='asis', echo=FALSE}
#| label: fig-cande-kept
#| fig-cap: "Revised (blue) and previously reported (green) kept catch estimates from 2005-2023. The vertical line on each graph identifies the first year that revised estimates are based on only on observed data (e.g. iREC and creel).  Years before the vertical line are modelled estimates that utilize past catch estimates."

era_model_catch <- 
  sport_release_rates %>% 
  filter(grepl("^T", finescale_fishery) == FALSE) |> #Exclude Terminal finescale fisheries, these just duplicate non-terminal catch in marine
  mutate(erafishery = case_when(
    grepl("CA JDF S", finescale_fishery) ~ "BC JF S",
    grepl("NGS S|SGS S", finescale_fishery) ~ "GEO ST S",
    grepl("WCVI", finescale_fishery) & grepl("AABM", finescale_fishery) ~  "WCVI AABM S",
    grepl("WCVI", finescale_fishery) & grepl("ISBM", finescale_fishery) ~  "WCVI ISBM S",
    grepl("NORTH AABM", finescale_fishery) ~  "NBC AABM S",
    grepl("NBC ISBM|NORTH ISBM", finescale_fishery) ~  "NBC ISBM S",
    grepl("CENTRAL", finescale_fishery) ~  "CBC S",
    grepl("JOHN ST", finescale_fishery) ~  "JNST S",
    .default = NA_character_
  )) %>%
  mutate(kept_total = marked_kept_total + unmarked_kept_total,
         released_total = marked_released_total + unmarked_released_total) %>%
  group_by(year, erafishery) %>%
  summarize(updated_kept_total = sum(kept_total, na.rm=TRUE), 
            updated_released_total = sum(released_total, na.rm=TRUE),
            .groups = "drop")

fishery_model_line <- data.frame(
  erafishery = c("BC JF S", "CBC S", "GEO ST S", 
                 "JNST S", "NBC AABM S", "NBC ISBM S",
                 "WCVI AABM S", "WCVI ISBM S"),
  xintercept = c(2013, 2015, 2013, 2013, 2015, 2015, 2013, 2013))


rec_summary <- 
  read.csv("data/rec_summary_2005-2023.csv") %>% 
  as_tibble() |>
  rename_with(tolower)

total_catch_compare <- 
  era_model_catch  %>% 
  left_join(rec_summary, by = c("erafishery", "year")) |>
  filter(!is.na(erafishery))


## Kept only
k<-ggplot() +
  geom_point(data=total_catch_compare, aes(y=updated_kept_total, x=year, col="lightblue4")) + 
  geom_line(data=total_catch_compare, aes(y=updated_kept_total, x=year, col="lightblue4"))+
  geom_point(data=total_catch_compare, aes(y=total_kept, x=year, col="lightgreen")) + 
  geom_line(data=total_catch_compare, aes(y=total_kept, x=year, col="lightgreen"))+
  scale_color_manual(values=c("lightblue4", "lightgreen"),  
                     labels = c("Revised estimates", "2024 C&E Report"))+
  facet_wrap(~erafishery, 
             scales = "free", 
             ncol=4)+
  geom_vline(data = fishery_model_line, aes(xintercept = xintercept)) +
  labs(col = "Data source",
       y = "Kept Catch",
       x = "Year") +
  theme_bw() +
  theme(legend.position="bottom")

print(k)

```

```{r results CandErelease, results='asis', echo=FALSE}
#| label: fig-cande-released
#| fig-cap: "Revised (blue) and previously reported (green) legal-sized release catch estimates from 2005-2023. The vertical line on each graph identifies the first year that revised estimates are based only on observed data (e.g., iREC and creel). Years before the vertical line are modelled based on past catch estimates."

rec_summary <- 
  read.csv("data/rec_summary_2005-2023.csv") |>
  as_tibble() |>
  rename_with(tolower)

total_catch_compare <- 
  era_model_catch  %>% 
  left_join(rec_summary, by = c("erafishery", "year")) |>
  filter(!is.na(erafishery))


## Released only
ggplot() +
  geom_point(data=total_catch_compare, aes(y=updated_released_total, x=year, col="lightblue4")) + 
  geom_line(data=total_catch_compare, aes(y=updated_released_total, x=year, col="lightblue4"))+
  geom_point(data=total_catch_compare, aes(y=legal_releases, x=year, col="lightgreen")) + 
  geom_line(data=total_catch_compare, aes(y=legal_releases, x=year, col="lightgreen"))+
  scale_color_manual(values=c("lightblue4", "lightgreen"),  
                     labels = c("Revised estimates", "2024 C&E Report"))+
  facet_wrap(~erafishery, 
             scales = "free", 
             ncol=4)+
  geom_vline(data = fishery_model_line, aes(xintercept = xintercept)) +
  labs(col = "Data source",
       y = "Legal-Size Released Catch",
       x = "Year") +
  theme_bw() +
  theme(legend.position="bottom")

```

The largest changes in released catch also occurred in CBC where previously very few releases were counted. This is because lodges in CBC do not report released catch, whereas iREC allows for released catch to be reported. We also now predict releases prior to 2012 in the Strait of Georgia, NBC ISBM, and BC Juan de Fuca, where previously very few releases were reported. The other fisheries had recorded releases but the revised estimates that include iREC increase the number of these released fish.

### Updating CWT estimates

As described in the methods, kept and marked catch estimates are used to expand CWT observations from the SHRP to CWT estimates uploaded to RMIS. Canada updated CWT estimates based on revised catch and published them to RMIS in support of ERA process starting in February 2025. Figure \@ref(fig:fig-indcwtest) shows the timeseries of total CWT estimates for all ERA indicator stocks, including both the original 2024 estimates and the revised estimates based on updated catch data. As expected, most of the differences between the 2024 and revised CWT estimates are again in earlier years, before the incorporation of iREC (identified by the vertical line). Furthermore, many of the fisheries show more consistent ranges in values over the whole revised timeseries (e.g., ERA fishery 63 - GEO ST S) compared to those in the 2024 timeseries. This suggests that the general fishery sample rate has been relatively consistent and that changes in indicator stock CWT estimates were potentially driven more by biases in historic marked catch estimates. Therefore, revised CWT estimates should provide a more consistent timeseries to evaluate indicator stock impacts.

```{r results indcwtest, results='asis', echo=FALSE}
#| label: fig-indcwtest
#| fig-cap: "Timeseries of CWT estimate totals for all PSC indicator stocks within the CTC 2025 ERA for each ERA fishery. The vertical line on each panel identifies the first year that revised CWT estimates used catch estimates based primarily on observed data (e.g., iREC and creel). Years before the vertical line use modelled catch based on past catch estimates."

cwt_est_filename <- "data/cwt_est_by_fishery.sql"
cwt_est_sql <- readChar(cwt_est_filename, file.info(cwt_est_filename)$size)

new_camp <- ctcUtil::open_camp_connection("camp_2025_local.yaml")
prev_camp <- ctcUtil::open_camp_connection("camp_2024.yaml")

new_cwt_est <- 
  dbGetQuery(new_camp, cwt_est_sql) |>
  mutate(fishery = paste0(fishery_id, " - ", fishery_name))

prev_cwt_est <- 
  dbGetQuery(prev_camp, cwt_est_sql) |>
  mutate(fishery = paste0(fishery_id, " - ", fishery_name))

dbDisconnect(prev_camp)
dbDisconnect(new_camp)

cwt_fishery_model_line <-
  data.frame(
    fishery = c("65 - BC JF S", "66 - TBC JF TERM S", 
                "63 - GEO ST S", "64 - TGEO ST TERM S", "61 - JNST S", 
                "62 - TJNST TERM S", 
                "50 - WCVI AABM S", "51 - WCVI ISBM S", "52 - TWCVI TERM S",
                "45 - CBC S", "46 - TCBC TERM S",
                "47 - NBC AABM S", "48 - NBC ISBM S", "49 - TNBC TERM S"),
    xintercept = c(2013, 2013, 2013, 2013, 2013, 2013, 2013, 2013, 2013,
                   2015, 2015, 2015, 2015, 2015)
  )

## CWT Estimate Total Timeseries
ggplot() +
  geom_point(data=new_cwt_est, aes(y=cwt_estimate, x=run_year, col="lightblue4")) + 
  geom_line(data=new_cwt_est, aes(y=cwt_estimate, x=run_year, col="lightblue4"))+
  geom_point(data=prev_cwt_est, aes(y=cwt_estimate, x=run_year, col="lightgreen")) + 
  geom_line(data=prev_cwt_est, aes(y=cwt_estimate, x=run_year, col="lightgreen"))+
  scale_color_manual(values=c( "lightblue4", "lightgreen"),
                     labels = c("Revised CWT Estimates", "2024 CWT Estimates"))+
  ylab("CWT Estimate (All Indicators)")+
  xlab("Run Year")+
  facet_wrap(~fishery, scales = "free", ncol=4)+
  theme_bw() + 
  theme(legend.position="bottom")+
  geom_vline(data = cwt_fishery_model_line, aes(xintercept = xintercept)) +
  labs(col = "Data source")

```

### Updating Chinook non-retention (release) estimates

Kept and released catch estimates are incorporated for most sport fisheries into the Chinook ERA non-retention method. In 2024, the Chinook ERA model was modified to only accounted for Chinook mark selective retention fisheries with the addition of the marked release rate ($MRR$) and unmarked kept rate ($UKR$). We calculated the $MRR$ and $UKR$ at the PFMA- and month-level, and then used a weighted average based on total catch estimate per area and month to generate an $MRR$ and $UKR$ for each finescale fishery. For more direct comparison to $MRR$ in the figures we show the unmarked release rate $URR$ (i.e. $1 - UKR$), although $UKR$ is actually used in the ERA. We can compare these marked and unmarked release rates based on revised catch estimates to release rates generated by the ratio between kept and released fish in the Chinook Non-Retention (CNR) file used in the 2024 ERA (Figure \@ref(fig:fig-mrrukr)). In the revised estimates, we allow release rates to vary by season, whereas in the past, finescale fisheries were not split into seasons and therefore release rates in each ERA fishery applied to all seasons.

We can see that some fisheries have been accounting for releases since 2005, while other fisheries only more recently incorporated release information. With the exception of Juan de Fuca Strait, which did have mark-selective fisheries before 2012, all other sport fisheries were non-selective before 2013. Since the data is modelled, we did not want to bias the previous non-selective fisheries by allowing for selectivity, therefore we have fixed the $MRR$ and $URR$ to be equal in the modelled data, by calculating a single release rate for marked and unmarked fish combined. After 2013, the data is observed and therefore allowed to vary selectively even if no specific mark selective fishery was specified in regulations. When $MRR$ and $URR$ are identical, this indicates non-selectivity. If $MRR$ is higher than $URR$, this indicates marked fish are released at a higher rate than unmarked fish, which would be indicative of preferential selection of unmarked fish for retention. If $URR$ is higher than $MRR$, then marked fish are released at a lower rate than unmarked fish, indicating preferential mark retention. Generally, the $MRR$ and $URR$ results suggest that releases are occurring but there is no strong, consistent selectivity to the releases. This aligns with MSF regulations that are very limited in space and time and are only beginning to be implemented in a small number of sport fisheries.

```{r results mrrukr, results='asis', echo=FALSE, fig.height = 8}
#| label: fig-mrrukr
#| fig-cap: "Release rate timeseries at the finescale fishery level, from 2005 to 2023. Light blue lines are release rates of unmarked fish and dark blue lines are release rates of marked fish from revised data. Green lines are release rates (regardless of mark status) used in the 2024 CNR files applied to the 2024 ERA."

fishery_map <-
  sport_catch_models$model_catch %>% 
  select(region, management) %>%
  mutate(erafishery = case_when(
    region %in% c("SWCVI", "NWCVI") & management == "AABM" ~ "WCVI AABM S",
    region %in% c("SWCVI", "NWCVI") & management == "ISBM" ~ "WCVI ISBM S",
    region %in% c("NGS", "SGS") ~  "GEO ST S",
    region == "JDF" ~ "BC JF S",
    region == "CBC" ~ "CBC S",
    region == "NC" & management == "AABM" ~ "NBC AABM S",
    region == "NC" & management == "ISBM" ~ "NBC ISBM S",
    region == "JNST" ~ "JNST S",
    .default = "Unknown"
  )) %>% 
  distinct()

cnr_canada_sport <- 
  read.csv("data/cnr_canada_sport.csv") %>% 
  as_tibble() %>% 
  rename_with(tolower) %>%
  filter(year > 2004) %>%
  full_join(fishery_map, by= c("erafishery"), relationship = "many-to-many")


finescale_fishery_map <- 
  sport_catch_models$model_catch %>% 
  select(finescale_fishery, region, management) %>% 
  distinct()


updated_relese_rates <-
  sport_release_rates |>
  inner_join(finescale_fishery_map, by = c("finescale_fishery")) |>
  group_by(year, region, management) |>
  summarize(marked_kept_total = sum(marked_kept_total, na.rm = TRUE),
            marked_released_total = sum(marked_released_total, na.rm = TRUE),
            unmarked_kept_total = sum(unmarked_kept_total, na.rm = TRUE),
            unmarked_released_total = sum(unmarked_released_total, na.rm = TRUE),
            .groups = "drop") |>
  mutate(mark_total = if_else(marked_kept_total > 0 | marked_released_total > 0,
                              marked_kept_total + marked_released_total,
                              0.0000001), #small value to prevent divide by zero
         unmark_total = if_else(unmarked_kept_total > 0 | unmarked_released_total > 0,
                                unmarked_kept_total + unmarked_released_total,
                                0.0000001), #small value to prevent divide by zero
         mrr = marked_released_total / mark_total,
         urr = unmarked_released_total / unmark_total) |>
  select(year, region, management, mrr, urr)


sport_mark_rate_mrr <-
  cnr_canada_sport |>
  full_join(updated_relese_rates, by=c("year", "region", "management")) |>
  mutate(across(c(mrr, urr), \(x) coalesce(x, 0)),
         region = if_else(region == "NC", "NBC", region),
         region_mgmt = paste(region, management))

mrr_fishery_model_line <-
  data.frame(
    region_mgmt = c("CBC ISBM", "NBC AABM", 
                    "NBC ISBM", "SWCVI AABM", "NWCVI AABM", 
                    "NWCVI ISBM", "SWCVI ISBM", "JNST ISBM",
                    "NGS ISBM", "SGS ISBM", "JDF ISBM"),
    xintercept = c(rep(2015, 3), rep(2013, 8))
  )

mu <- 
  ggplot() +
  geom_point(data=sport_mark_rate_mrr, aes(y=urr, x=year, col="lightblue")) +
  geom_line(data=sport_mark_rate_mrr, aes(y=urr, x=year, col="lightblue"))+
  geom_point(data=sport_mark_rate_mrr, aes(y=mrr, x=year, col="lightblue4")) +
  geom_line(data=sport_mark_rate_mrr, aes(y=mrr, x=year, col="lightblue4"))+
  geom_point(data=sport_mark_rate_mrr, aes(y=release_rate, x=year, col="lightgreen")) +
  geom_line(data=sport_mark_rate_mrr, aes(y=release_rate, x=year, col="lightgreen"))+
  scale_color_manual(values=c("lightblue", "lightblue4", "lightgreen"),  
                     labels = c("Revised Unmarked Release Rate", 
                                "Revised Marked Release Rate", 
                                "2024 CNR Release Rate"))+
  facet_wrap(~region_mgmt) +
  theme_bw() + 
  geom_vline(data = mrr_fishery_model_line, aes(xintercept = xintercept)) +
  labs(col="Data source",
       x = "Year",
       y = "Release Rate") +
  theme(legend.position = "bottom")

print(mu)
```
