`r if(knitr:::is_latex_output()) '% begin csasdown appendix'` `r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'`

\clearpage

# Stratification Mapping {#apdx-strattab}

```{r methods strattab, echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-stratatab

strat_table <- 
  sport_catch_area |> 
  distinct(finescale_fishery, area, month) |>
  arrange(finescale_fishery,  area, month)

fishery_map <-
  sport_catch_area |>
  distinct(region, management, finescale_fishery) |>
  mutate(erafishery = case_when(
    region %in% c("SWCVI", "NWCVI") & management == "AABM" ~ "WCVI AABM S",
    region %in% c("NGS", "SGS") ~  "GEO ST S",
    region == "JDF" ~ "BC JF S",
    region == "CBC" ~ "CENTRAL S",
    region == "JNST" ~ "JNST S",
    region %in% c("SWCVI", "NWCVI") & management == "ISBM" ~ "WCVI ISBM S",
    region == "NC" & management == "ISBM" ~ "NBC ISBM S",
    region == "NC" & management == "AABM" ~ "NBC AABM S",
    region == "TSWCVI" ~ "SWCVI TERMINAL",
    .default = NA_character_)) |>
  select(finescale_fishery, erafishery) |>
  distinct()

strat_table <- 
  strat_table |>
  left_join(fishery_map, by = "finescale_fishery") |>
  arrange(erafishery) |> 
  relocate(erafishery) |>
  group_by(erafishery, finescale_fishery, area) |>
  summarize(month_category = paste0(min(month), " to ", max(month)),
            .groups = "drop") |>
  distinct() |> 
  arrange(erafishery, finescale_fishery, area, month_category) |>
  rename("CTC ERA Fishery" = erafishery,  
         "CTC Finescale Fishery" = finescale_fishery, 
         "PFMA" = area,
         "Months" = month_category)


csas_table(strat_table, format = "latex")
#strat_table %>%
#  kableExtra::kable("latex", 
#                    escape = FALSE,
#                    caption = "Stratification mapping",
#                    align = "l") #%>%
#  kableExtra::kable_classic(full_width = F, html_font = "Cambria") %>%
#  kableExtra::kable_styling(bootstrap_options = c("bordered")) %>%
# kableExtra::collapse_rows(columns = 1:3, valign = "middle")


```

\clearpage

# Revised Estimates 2005-2023 {#app:first-appendix}

```{r m plots, results='asis', echo=FALSE}
#| label: fig-catch-revised-1-appendix
#| fig-cap: "Revised yearly kept catch estimates"

### make this by finescale fishery old... sum up. 
models_combined_old<-
  sport_catch_models$model_catch |>
  filter(region %notin% c("CBC", "NC")) |>
  mutate(region_mgmt = if_else(region %in% c("SWCVI", "NWCVI"),
                               paste(region, management),
                               region)) |>
  group_by(year, kept_status, pred_cat, region_mgmt) |> 
  summarise(across(where(is.double), sum, na.rm = TRUE), .groups = "drop")


fishery_name_south <- sort(unique(models_combined_old$region_mgmt))


for (i in 1:length(fishery_name_south)){
  graph_data <- 
    models_combined_old  %>% 
    filter(region_mgmt == fishery_name_south[i],
           between(year, 2005, 2023),
           kept_status == "kept") %>%
    select(year, creel_unfiltered_plus, catch_estimate_predicted) %>%
    pivot_longer(c(creel_unfiltered_plus, catch_estimate_predicted),
                 names_to = "estimate_type",
                 values_to = "catch_estimate")
  
  m <- ggplot(graph_data, aes(y = catch_estimate,
                              x = year,
                              shape = estimate_type,
                              linetype = estimate_type))+
    geom_point(size = 2.5,
               colour = "#440154")+
    geom_line(colour = "#440154") +
    scale_shape_manual(values = c(1, 16),
                       labels = c("iREC included", "Unfiltered creel and logbook")
    ) +
    scale_linetype_manual(values = c(2, 1),
                          labels = c("iREC included", "Unfiltered creel and logbook")
    ) +
    scale_x_continuous(breaks = seq.int(2005, 2023, by = 2)) +
    ggtitle(paste(fishery_name_south[i])) + 
    theme_bw() + 
    ylab("Kept catch estimate") + 
    xlab("Year") +
    labs(linetype = "Catch Estimate", shape = "Catch Estimate")
  
  print(m)
}
```

```{r m plots, results='asis', echo=FALSE}
#| label: fig-catch-revised-2-appendix
#| fig-cap: "revised yearly catch estimates"



### make this by finescale fishery old... sum up. 
models_combined_old2 <-
  sport_catch_models$model_catch |>
  filter(region %in% c("CBC", "NC")) |>
  mutate(region_mgmt = if_else(region == "NC",
                               paste(region, management),
                               region)) |>
  group_by(year, kept_status, pred_cat, region_mgmt) |> 
  summarise(across(where(is.double), sum, na.rm = TRUE), .groups = "drop")

fishery_name_north<- sort(unique(models_combined_old2$region_mgmt))

for (i in 1:length(fishery_name_north)) {
  graph_data <- 
    models_combined_old2  %>% 
    filter(region_mgmt == fishery_name_north[i],
           between(year, 2005, 2023),
           kept_status == "kept") %>%
    select(year, historic_plus, catch_estimate_predicted) %>%
    pivot_longer(c(historic_plus, catch_estimate_predicted),
                 names_to = "estimate_type",
                 values_to = "catch_estimate")
  
  m <- ggplot(graph_data, aes(y = catch_estimate,
                              x = year,
                              shape = estimate_type,
                              linetype = estimate_type)) +
    geom_point(size = 2.5,
               colour = "#440154")+
    geom_line(colour = "#440154") +
    scale_shape_manual(values = c(1, 16),
                       labels = c("iREC included", "Historic")
    ) +
    scale_linetype_manual(values = c(2, 1),
                          labels = c("iREC included", "Historic")
    ) +
    scale_x_continuous(breaks = seq.int(2005, 2023, by = 2)) +
    ggtitle(paste(fishery_name_north[i])) + 
    theme_bw() + 
    ylab("Kept catch estimate") + 
    xlab("Year") +
    labs(linetype = "Catch Estimate", shape = "Catch Estimate")
  
  print(m)
}
```

\clearpage

# Model Residuals {#app:second-appendix}

```{r Modelling summer subsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-south-sum-residuals
#| fig-cap: "Validation plots for BC South Summer model"

res_gam <- simulateResiduals(sport_catch_models$model_list$south_summer, plot = T, quantreg=T)
```

```{r Modelling spring and fall subsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-south-sf-residuals
#| fig-cap: "Validation plots for BC South Spring and Fall model"

res_gam <- simulateResiduals(sport_catch_models$model_list$south_season, plot = T, quantreg=T)

```

```{r Modelling nbc aabm subsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-nbcaabm-residuals
#| fig-cap: "Validation plots for NBC AABM annual model"

res_gam <- simulateResiduals(sport_catch_models$model_list$north_aabm, plot = T, quantreg=T)

```

```{r Modelling summer nbc isbm subsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-nbcisbm-sum-residuals
#| fig-cap: "Validation plots for NBC ISBM Summer model"

res_gam <- simulateResiduals(sport_catch_models$model_list$north_isbm_summer, plot = T, quantreg=T)

```

```{r Modelling NBC ISBM spring and fall subsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-nbcisbm-sf-residuals
#| fig-cap: "Validation plots for NBC ISBM Spring and Fall model"

res_gam <- simulateResiduals(sport_catch_models$model_list$north_isbm_season, plot = T, quantreg=T)

```

```{r Modelling CBCsubsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-cbc-residuals
#| fig-cap: "Validation plots for CBC annual model"

res_gam <- simulateResiduals(sport_catch_models$model_list$central_isbm, plot = T, quantreg=T)

```

\clearpage

# Model Output {#app:third-appendix}

```{r appendix_plotting_function, include = FALSE}
#plotting function that will be used for the next few sections
create_facet_plot <- function(finescale_fishery_name, x_var, plot_data, model_data) {
  # Labels for facets
  status_labels <- c(
    "marked_kept_total" = "Marked Kept Total",
    "marked_released_total" = "Marked Released Total",
    "unmarked_kept_total" = "Unmarked Kept Total",
    "unmarked_released_total" = "Unmarked Released Total"
  )
  
  # Colors for status
  status_colors <- c(
    "marked_kept_total" = "#440154", 
    "marked_released_total" = "#31688e",
    "unmarked_kept_total" = "#35b779",
    "unmarked_released_total" = "#d4be02"
  )
  
  # Filter data for the specific finescale fishery
  filtered_data <- plot_data %>% filter(finescale_fishery == finescale_fishery_name)
  filtered_model_data <- model_data %>% filter(finescale_fishery == finescale_fishery_name)
  
  #check if all catch_estimate values are zero for each status
  #wont want ribbon and line if all data are 0 like in figure 10
  has_nonzero <- filtered_data %>%
    group_by(status) %>%
    summarize(has_nonzero = any(catch_estimate != 0), .groups = "drop")
  
  # Create the plot
  facet_plot <- ggplot(filtered_data, aes_string(x = x_var, y = "catch_estimate", col = "status", fill = "status")) +
    geom_point() +
    theme_bw() +
    scale_size_continuous(range = c(1, 3)) +
    scale_colour_manual(values = status_colors, guide = "none") +
    scale_fill_manual(values = status_colors, guide = "none") +
    xlab("Creel summer catch estimate") +
    ylab("Calibrated iREC-only catch estimate") +
    facet_wrap(~status,
               labeller = labeller(status = status_labels),
               scales = "free")
  
  # Conditionally add geom_line and geom_ribbon for statuses with non-zero data
  for (status in unique(filtered_data$status)) {
    if (has_nonzero %>% filter(status == !!status) %>% pull(has_nonzero)) {
      facet_plot <- facet_plot +
        geom_line(data = filtered_model_data %>% filter(status == !!status),
                  aes_string(y = "fit", x = x_var, col = "status")) +
        geom_ribbon(data = filtered_model_data %>% filter(status == !!status),
                    aes_string(y = "fit", x = x_var, ymin = "right_lwr", ymax = "right_upr", fill = "status"), 
                    alpha = 0.10)
    }
  }
  
  return(facet_plot)
}
```

## Southern BC Summer fisheries

```{r forloop seasonal summer appendix, results='asis', echo=FALSE}
south_summer_model <- sport_catch_models$model_list$south_summer
south_summer_no_nas <- south_summer_model$data

family_set <- family(south_summer_model)
ilink_family_set<- family_set$linkinv


ndata <-
  south_summer_no_nas %>% 
  group_by(status, finescale_fishery, season) %>%  
  tidyr::expand(creel_plus_summer = seq(0, max(creel_plus_summer), length=100)) %>%
  
  
  ## add the fitted values by predicting from the model for the new data
  add_column(fit = predict(south_summer_model, 
                           newdata = ., 
                           type = 'response')) %>%
  bind_cols(setNames(as_tibble(predict(south_summer_model, 
                                       ., 
                                       se.fit = TRUE)[1:2]),
                     c('fit_link','se_link'))) %>%
  mutate(fit_resp  = ilink_family_set(fit_link),
         right_upr = ilink_family_set(fit_link + (2 * se_link)),
         right_lwr = ilink_family_set(fit_link - (2 * se_link)))

fishery_name_south <- sort(unique(south_summer_no_nas$finescale_fishery))

for (i in 1:length(fishery_name_south)){
  
  plot <- create_facet_plot(finescale_fishery_name = fishery_name_south[i], 
                            x_var = "creel_plus_summer", 
                            plot_data = south_summer_no_nas, 
                            model_data = ndata) +
    ggtitle(paste(fishery_name_south[i])) + 
    #changing the y lab
    ylab("IREC-included catch estimate")
  
  print(plot)
}
```

## Southern BC Spring and Fall fisheries

```{r forloop seasonal spring and fall appendix, results='asis', echo=FALSE}

south_season_model <- sport_catch_models$model_list$south_season
south_season_no_nas <- south_season_model$data

family_set <- family(south_season_model)
ilink_family_set<- family_set$linkinv

ndata <-
  south_season_no_nas %>% 
  group_by(status, region_mgmt, finescale_fishery, season) %>%  
  tidyr::expand(creel_plus_summer = seq(0, max(creel_plus_summer), length=100)) %>%
  ## add the fitted values by predicting from the model for the new data
  add_column(fit = predict(south_season_model, newdata = ., type = 'response')) %>%
  bind_cols(setNames(as_tibble(predict(south_season_model, ., se.fit = TRUE)[1:2]),
                     c('fit_link','se_link'))) %>%
  mutate(fit_resp  = ilink_family_set(fit_link),
         right_upr = ilink_family_set(fit_link + (2 * se_link)),
         right_lwr = ilink_family_set(fit_link - (2 * se_link)))


fishery_name_south<- sort(unique(south_season_no_nas$finescale_fishery))

for (i in 1:length(fishery_name_south)){
  
  plot <- create_facet_plot(finescale_fishery_name = fishery_name_south[i], 
                            x_var = "creel_plus_summer", 
                            plot_data = south_season_no_nas, 
                            model_data = ndata) +
    ggtitle(paste(fishery_name_south[i]))
  
  print(plot)
}

```

## Northern BC AABM fisheries

```{r forloop NBC AABM annual appendix, results='asis', echo=FALSE}

north_aabm_model <- sport_catch_models$model_list$north_aabm

family_set <- family(north_aabm_model)
ilink_family_set<- family_set$linkinv

ndata <-
  north_aabm_model$data %>% 
  group_by(status, finescale_fishery, season) %>%  
  tidyr::expand(historic_plus = seq(0, max(historic_plus), length=100)) %>%
  ## add the fitted values by predicting from the model for the new data
  add_column(fit = predict(north_aabm_model, 
                           newdata = ., 
                           type = 'response')) %>%
  bind_cols(setNames(as_tibble(predict(north_aabm_model, 
                                       ., 
                                       se.fit = TRUE)[1:2]),
                     c('fit_link','se_link'))) %>%
  
  mutate(fit_resp  = ilink_family_set(fit_link),
         right_upr = ilink_family_set(fit_link + (2 * se_link)),
         right_lwr = ilink_family_set(fit_link - (2 * se_link)))


fishery_name_north_aabm<- sort(unique(north_aabm_model$data$finescale_fishery))


for (i in 1:length(fishery_name_north_aabm)){
  plot <- create_facet_plot(finescale_fishery_name = fishery_name_north_aabm[i], 
                            x_var = "historic_plus", 
                            plot_data = north_aabm_model$data, 
                            model_data = ndata) +
    ggtitle(paste(fishery_name_north_aabm[i]))
  
  print(plot)
}
```

## Northern BC ISBM Spring and Fall fisheries

```{r forloop NBC ISBM seasonal spring and fall appendix, results='asis', echo=FALSE}


north_isbm_season_model <- sport_catch_models$model_list$north_isbm_season
north_isbm_season_no_nas <- north_isbm_season_model$data

family_set <- family(north_isbm_season_model)
ilink_family_set<- family_set$linkinv

ndata <-
  north_isbm_season_no_nas %>% 
  group_by(status, finescale_fishery, season) %>%  
  tidyr::expand(historic_summer = seq(0, max(historic_summer), length=100)) %>%
  ## add the fitted values by predicting from the model for the new data
  add_column(fit = predict(north_isbm_season_model, 
                           newdata = ., 
                           type = 'response')) %>%
  bind_cols(setNames(as_tibble(predict(north_isbm_season_model, 
                                       ., 
                                       se.fit = TRUE)[1:2]),
                     c('fit_link','se_link'))) %>%
  
  mutate(fit_resp  = ilink_family_set(fit_link),
         right_upr = ilink_family_set(fit_link + (2 * se_link)),
         right_lwr = ilink_family_set(fit_link - (2 * se_link)))


fishery_name_north_isbm<- sort(unique(north_isbm_season_no_nas$finescale_fishery))


for (i in 1:length(fishery_name_north_isbm)){
  
  plot <- create_facet_plot(finescale_fishery_name = fishery_name_north_isbm[i], 
                            x_var = "historic_summer", 
                            plot_data = north_isbm_season_no_nas, 
                            model_data = ndata) +
    ggtitle(paste(fishery_name_north_isbm[i]))
  
  print(plot)
}
```

`r if(knitr:::is_latex_output()) '% end csasdown appendix'`


\clearpage

# Modified Haida Gwaii Model {#app:haida-appendix}


## Catch Source Time Series
```{r NBC AABM Catch Time Series appendix, results='asis', echo=FALSE}

#Limit catch to Historic Plus and iRec Calibrated
nbc_aabm <-
  sport_catch_finescale |>
  filter(management == "AABM", region == "NC") |>
  group_by(year, status) |>
  summarize(across(c(historic_plus, irec_calibrated), sum, na.rm=TRUE),
            .groups = "drop") |>
  select(year, status, historic_plus, irec_calibrated) |>
  tidyr::pivot_longer(c(historic_plus, irec_calibrated))


#Incorporate Summer Historic Catch
nbc_aabm <-
  sport_catch_finescale |>
  filter(management == "AABM", region == "NC", season == "summer") |>
  group_by(year, status) |>
  summarize(across(c(historic_summer), sum, na.rm=TRUE),
            .groups = "drop") |>
  select(year, status, historic_summer) |>
  tidyr::pivot_longer(c(historic_summer)) |>
  bind_rows(nbc_aabm) 


nbc_aabm |>
  mutate(name = case_when(name == "historic_summer" ~ "Historic Summer",
                          name == "historic_plus" ~ "All Historic",
                          name == "irec_calibrated" ~ "Calibrated iREC",
                          .default = "")) |>
  ggplot(aes(x = year, y = value, color = status)) +
  geom_line(linewidth = 1) +
  geom_point() +
  facet_wrap(~ name, nrow = 3) +
  labs(title = "NBC AABM Catch Source Time Series",
       x = "year",
       y = "Catch") +
  theme_minimal()

```


## Previous Model Fits

### Spring/Fall

$$ C_{spring/fall} = Z + S + C_{historic, summer}:Z $$

```{r fit NBC AABM Gamma log models, results='asis', echo=FALSE}
season_north_aabm <-
  sport_catch_area |>
  filter(region == "NC",
         management == "AABM") |>
  mutate(season = case_when(between(month, 1,4) ~ "spring",
                            between(month, 5, 9) ~ "summer",
                            between(month, 10, 12) ~ "fall",
                            .default = "")) |>
  mutate(historic_summer = if_else(between(month, 7, 8), coalesce(historic_plus, 0), 0))

historic_summer_df <- 
  season_north_aabm |>
  group_by(year, status) |>
  summarize(historic_summer = sum(historic_summer, na.rm = TRUE), .groups = "drop")

summer_north_aabm <-
  season_north_aabm |>
  filter(season == "summer") |>
  group_by(year, status) |>
  summarize(historic_plus = sum(historic_plus, na.rm = TRUE), 
            catch_estimate  = sum(catch_estimate , na.rm = TRUE), 
            .groups = "drop") |>
  inner_join(historic_summer_df, by = c("year", "status"))

season_north_aabm <-
  season_north_aabm |>
  filter(season != "summer") |>
  group_by(year, status, season) |>
  summarize(historic_plus = sum(historic_plus, na.rm = TRUE), 
            catch_estimate  = sum(catch_estimate , na.rm = TRUE),
            .groups = "drop") |>
  right_join(historic_summer_df, by = c("year", "status"))


season_north_aabm_gamma_spec <-
  season_north_aabm %>%
  filter(year >= 2015) %>%
  glm(formula = catch_estimate + 1 ~ season + status + historic_summer:season +
        1 + historic_summer,
      family=Gamma(link = "log"),
      data = .)

season_north_aabm_full <-
  season_north_aabm %>%
  filter(year < 2015) %>%
  mutate(catch_estimate_adj = predict.glm(season_north_aabm_gamma_spec,
                                          newdata = .,
                                          type = "response")) |>
  bind_rows(filter(season_north_aabm, year >= 2015)) |>
  mutate(catch_estimate_adj = coalesce(catch_estimate_adj, catch_estimate))

season_north_aabm_full |>
  group_by(year, status) |>
  summarize(catch_estimate_adj = sum(catch_estimate_adj, na.rm = TRUE), .groups = "drop") |> 
  ggplot(aes(x = year, y = catch_estimate_adj, color = status)) +
  geom_line(linewidth = 1) +
  geom_point() +
  labs(title = "NBC AABM Catch Source Time Series",
       x = "year",
       y = "Catch") +
  theme_minimal()
```



### Summer

$$ C_{summer} = C_{historic, summer} + S + C_{historic, summer}:S $$

```{r}

summer_north_aabm_gamma_spec <-
  summer_north_aabm %>%
  filter(between(year, 2015, 2023)) %>%
  glm(formula = catch_estimate + 1 ~ status + historic_summer:status +
        1 + historic_summer,
      family=Gamma(link = "log"),
      data = .)

summer_north_aabm_full <-
  summer_north_aabm %>%
  filter(year < 2015) %>%
  mutate(catch_estimate_adj = predict.glm(summer_north_aabm_gamma_spec,
                                          newdata =  .,
                                          type = "response")) |>
  bind_rows(filter(summer_north_aabm, year >= 2015)) |>
  mutate(catch_estimate_adj = coalesce(catch_estimate_adj, catch_estimate))


summer_north_aabm_full |>
  group_by(year, status) |>
  summarize(catch_estimate_adj = sum(catch_estimate_adj, na.rm = TRUE), .groups = "drop") |> 
  ggplot(aes(x = year, y = catch_estimate_adj, color = status)) +
  geom_line(linewidth = 1) +
  geom_point() +
  labs(title = "NBC AABM Summer Catch Source Time Series",
       x = "year",
       y = "Catch") +
  theme_minimal()
```

## New Model Fits

$$ C_{all} = C_{historic} + S $$

```{r}
all_north_aabm <-
  sport_catch_finescale |>
  filter(grepl("^NORTH AABM S", finescale_fishery)) |>
  group_by(year, status) |>
  summarize(across(c(irec_calibrated, historic_plus), sum, na.rm = TRUE),
            .groups = "drop")

all_north_aabm_predict <-
  all_north_aabm |>
  filter(year < 2015)

fit_north_aabm_no_nas <-
  all_north_aabm |>
  filter(between(year, 2015, 2023)) |>
  mutate(catch_estimate = irec_calibrated)

all_north_aabm_gaussian_spec <-
  glm(formula = catch_estimate + 1 ~ status + 1 + historic_plus,
      data = fit_north_aabm_no_nas)

all_north_aabm_full <-
  all_north_aabm_predict |>
  mutate(catch_estimate_adj = predict.glm(all_north_aabm_gaussian_spec,
                                          newdata = all_north_aabm_predict,
                                          type = "response")) |>
  bind_rows(fit_north_aabm_no_nas) |>
  mutate(catch_estimate_adj = coalesce(catch_estimate_adj, irec_calibrated))



all_north_aabm_full |>
  group_by(year, status) |>
  summarize(catch_estimate_adj = sum(catch_estimate_adj, na.rm = TRUE), .groups = "drop") |> 
  ggplot(aes(x = year, y = catch_estimate_adj, color = status)) +
  geom_line(linewidth = 1) +
  geom_point() +
  labs(title = "NBC AABM Catch Source Time Series",
       x = "year",
       y = "Catch") +
  theme_minimal()
```

## AIC Comparision


```{r}


fit_aic_df <-
  data.frame(Model = c("Original Summer (Gamma Log)", "Original Spring/Fall (Gamma Log)", "New Annual Gaussian Fit"),
             AIC = c(summer_north_aabm_gamma_spec$aic,
                     season_north_aabm_gamma_spec$aic,
                     all_north_aabm_gaussian_spec$aic))

knitr::kable(fit_aic_df)

```

### Change in Annual catch

```{r}

previous_model_catch <-
  season_north_aabm_full |>
  select(year, status, spring_catch = catch_estimate_adj) |>
  group_by(year, status) |>
  summarize(spring_catch = sum(spring_catch, na.rm = TRUE), .groups = "drop") |>
  full_join(select(summer_north_aabm_full, year, status, summer_catch = catch_estimate_adj), 
            by = c("year", "status"))

catch_diff <-
  previous_model_catch |>
  full_join(select(all_north_aabm_full, year, status, new_annual_catch = catch_estimate_adj), 
            by = c("year", "status")) |>
  mutate(old_annual_catch = spring_catch + summer_catch,
         catch_diff = new_annual_catch - old_annual_catch)


# Find max absolute value
max_y <- max(abs(catch_diff$catch_diff))


catch_diff |>
  ggplot(aes(x = year, y = catch_diff, color = status)) +
  geom_line(linewidth = 1) +
  geom_point() +
  coord_cartesian(ylim = c(-max_y, max_y)) +
  labs(title = "NBC AABM Catch Change Time Series",
       x = "year",
       y = "Catch") +
  theme_minimal()
```



