`r if(knitr:::is_latex_output()) '% begin csasdown appendix'` `r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'`

\clearpage

# REVISED ESTIMATES 2013-2022 {#app:first-appendix}

```{r m plots, results='asis', echo=FALSE}
#| label: fig-catch-revised-1-appendix
#| fig-cap: "Revised yearly kept catch estimates"


library(ROracle)
library(pacRecCatch)
library(tidyverse)
library(odbc)
library(lubridate)
library(ggnewscale)
library(patchwork)
library(DHARMa)
library(lme4)
library(bbmle)
library(SuppDists)
library(MuMIn)

models_combined <- pacRecCatch::model_rec_catch(Sport_mark_rate_finescale_combined)


### make this by finescale fishery old... sum up. 
models_combined_old<-models_combined %>% group_by(YEAR,kept_status,pred_cat, finescale_fishery_old) %>% summarise_if(is.numeric, sum, na.rm = TRUE) %>% filter(!str_detect(finescale_fishery_old, "CBC|NBC"))

fishery_name_south<- sort(unique(models_combined_old$finescale_fishery_old))

for (i in 1:length(fishery_name_south)){

 m<-ggplot(models_combined_old  %>% filter(finescale_fishery_old==fishery_name_south[i], YEAR %in% c(2013:2023), kept_status=="Kept")) +
    geom_point(size=2.5,  aes(y=creel_unfiltered_plus, x=YEAR,col=kept_status, fill=kept_status, shape=pred_cat))+
    geom_line(aes(y=creel_unfiltered_plus, x=YEAR,col=kept_status, linetype = " Unfiltered creel and logbook"))+
    scale_shape_manual(values=c(1,1), guide="none")+
    scale_x_continuous(breaks = c(2013, 2015, 2017, 2019, 2021, 2023))+
    new_scale("shape")+
    scale_colour_viridis_d()+
    geom_point(size=2.5, aes(y=catch_estimate_predicted, x=YEAR,col=kept_status, fill=kept_status, shape=pred_cat))+
    geom_line(aes(y=catch_estimate_predicted, x=YEAR,col=kept_status, linetype = "iREC included"))+
    scale_shape_manual(values=c(16,17), guide="none")+
    scale_linetype_manual(values=c(2,1))+
   ggtitle(paste(fishery_name_south[i])) + theme_bw() + scale_x_continuous(breaks = c(2013, 2015, 2017, 2019, 2021, 2023))+ ylab("Kept catch estimate") + xlab("Year")
 
 
 print(m)
}
```

```{r m plots, results='asis', echo=FALSE}
#| label: fig-catch-revised-2-appendix
#| fig-cap: "revised yearly catch estimates"



### make this by finescale fishery old... sum up. 
models_combined_old2<-models_combined %>% group_by(YEAR,kept_status,pred_cat, finescale_fishery_old) %>% summarise_if(is.numeric, sum, na.rm = TRUE) %>% filter(str_detect(finescale_fishery_old, "CBC|NBC"))

fishery_name_north<- sort(unique(models_combined_old2$finescale_fishery_old))

for (i in 1:length(fishery_name_north)){

 m2<-ggplot(models_combined_old2  %>% filter(finescale_fishery_old==fishery_name_north[i], YEAR %in% c(2013:2023), kept_status=="Kept")) +
    geom_point(size=2.5,  aes(y=historic_plus, x=YEAR,col=kept_status, fill=kept_status, shape=pred_cat))+
    geom_line(aes(y=historic_plus, x=YEAR,col=kept_status, linetype = "Historic"))+
    scale_shape_manual(values=c(1,1), guide="none")+
    scale_x_continuous(breaks = c(2013, 2015, 2017, 2019, 2021, 2023))+
    new_scale("shape")+
    scale_colour_viridis_d()+
    geom_point(size=2.5, aes(y=catch_estimate_predicted, x=YEAR,col=kept_status, fill=kept_status, shape=pred_cat))+
    geom_line(aes(y=catch_estimate_predicted, x=YEAR,col=kept_status, linetype = "iREC included"))+
    scale_shape_manual(values=c(16,17), guide="none")+
    scale_linetype_manual(values=c(2,1))+
   ggtitle(paste(fishery_name_north[i])) + theme_bw() + scale_x_continuous(breaks = c(2013, 2015, 2017, 2019, 2021, 2023))+ ylab("Kept catch estimate") + xlab("Year")
 
 
 print(m2)
}
```

\clearpage

# MODEL RESIDUALS {#app:second-appendix}

### Southern BC Summer fisheries

```{r Modelling summer subsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-south-sum-residuals
#| fig-cap: "Validation plots for BC South Summer model"


Summer_south<-Sport_mark_rate_finescale_combined %>%
  filter(YEAR %in% c(2013:2023))%>%
  filter(!str_detect(finescale_fishery, "CBC|NBC")) %>%
  filter(season %in% c("summer"))

Summer_south_no_nas<-Summer_south %>% drop_na(any_of(c("creel_plus_summer", "status", "finescale_fishery_old")))
write_rds(Summer_south_no_nas, "Summer_south_no_nas.RDS")

Summer_model<- glm(formula = catch_estimate ~ finescale_fishery_old + status + creel_plus_summer:finescale_fishery_old + creel_plus_summer:status +
finescale_fishery_old:status + 1 + creel_plus_summer,  family=Gamma(link = "log"), data = Summer_south_no_nas)

res_gam <- simulateResiduals(Summer_model, plot = T, quantreg=T)
```

### Northern BC AABM Summer fisheries

```{r Modelling summer nbc aabm subsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-nbcaabm-sum-residuals
#| fig-cap: "Validation plots for NBC AABM Summer model"

Summer_north_aabm<-Sport_mark_rate_finescale_combined%>%
  filter(YEAR %in% c(2015:2023)) %>%
  filter(finescale_fishery_old == "NBC AABM S")%>%
  filter( season=="summer")

#Modelling comparisons need to be done on models with same # of NAs - so drop nas
Summer_north_aabm_no_nas<-Summer_north_aabm %>% drop_na(any_of(c("historic_summer", "status", "finescale_fishery_old", "season", "historic_effort")))

###Chosen model
Summer_north_aabm_model_gamma_spec<- glm(formula = catch_estimate + 1 ~ status + historic_summer:status + 1 + historic_summer,  family=Gamma(link = "log"), data = Summer_north_aabm_no_nas)

res_gam <- simulateResiduals(Summer_north_aabm_model_gamma_spec, plot = T, quantreg=T)

```

### Northern BC ISBM Summer fisheries

```{r Modelling summer nbc isbm subsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-nbcisbm-sum-residuals
#| fig-cap: "Validation plots for NBC ISBM Summer model"

Summer_north_isbm<-Sport_mark_rate_finescale_combined%>% filter(YEAR %in% c(2015:2023)) %>% filter(finescale_fishery_old == "NBC ISBM S")%>%
  filter(season %in% c("summer"))
Summer_north_isbm_no_nas<-Summer_north_isbm %>% drop_na(any_of(c("historic_summer", "status", "finescale_fishery_old", "season", "historic_effort")))
write_rds(Summer_north_isbm_no_nas, "Summer_north_isbm_no_nas.RDS")

#Chosen model
Summer_north_isbm_model_full_gamma_spec<- glm(formula = catch_estimate + 1 ~ status + 1 + historic_summer,  family=Gamma(link = "log"), data = Summer_north_isbm_no_nas)
res_spec <- simulateResiduals(Summer_north_isbm_model_full_gamma_spec, plot = T, quantreg=T)


```

### Southern BC Spring and Fall fisheries

```{r Modelling spring and fall subsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-south-sf-residuals
#| fig-cap: "Validation plots for BC South Spring and Fall model"

### Data needed
Season_south_sf<-Sport_mark_rate_finescale_combined %>%
  filter(YEAR %in% c(2013:2023))%>%
  filter(!str_detect(finescale_fishery, "CBC|NBC")) %>%
  filter(season %in% c("spring", "fall"))

#Modelling comparisons need to be done on models with same # of NAs - so drop nas
Season_south_no_nas<-Season_south_sf %>% drop_na(any_of(c("creel_plus_summer", "status", "finescale_fishery_old", "season")))

Spring_fall_model<-  glm(formula =catch_estimate + 3 ~ finescale_fishery_old + season +   status + creel_plus_summer:finescale_fishery_old + creel_plus_summer:status +  finescale_fishery_old:season + 1 + creel_plus_summer,  family=Gamma(link = "log"), data = Season_south_no_nas)

res_gam <- simulateResiduals(Spring_fall_model, plot = T, quantreg=T)
```

### Northern BC AABM Spring and Fall fisheries

```{r Modelling NBC AABM spring and fall subsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-nbcaabm-sf-residuals
#| fig-cap: "Validation plots for NBC AABM Spring and Fall model"

Season_north_aabm<-Sport_mark_rate_finescale_combined%>% filter(YEAR %in% c(2015:2023)) %>% filter(finescale_fishery_old == "NBC AABM S") %>%
  filter(season %in% c("spring", "fall"))
Season_north_aabm_no_nas<-Season_north_aabm %>% drop_na(any_of(c("historic_summer", "status", "finescale_fishery_old", "season", "historic_effort")))


###Chosen model
North_aabm_model<- glm(formula = catch_estimate + 1 ~ season + status + historic_summer:season + 1 + historic_summer,  family=Gamma(link = "log"), data = Season_north_aabm_no_nas)

res_gam <- simulateResiduals(North_aabm_model, plot = T, quantreg=T)
```

### Northern BC ISBM Spring and Fall fisheries

```{r Modelling NBC ISBM spring and fall subsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-nbcisbm-sf-residuals
#| fig-cap: "Validation plots for NBC ISBM Spring and Fall model"


Season_nbc_isbm<-Sport_mark_rate_finescale_combined%>% filter(YEAR %in% c(2015:2023)) %>% filter(finescale_fishery_old == "NBC ISBM S")%>%
  filter(season %in% c("spring", "fall"))
Season_nbc_isbm_no_nas<-Season_nbc_isbm %>% drop_na(any_of(c("historic_summer", "status", "finescale_fishery_old", "season", "historic_effort")))


###Chosen model
nbc_isbm_model_full_gamma_spec<- glm(formula = catch_estimate + 1 ~ season + status + season:status + 1 + historic_summer,  family=Gamma(link = "log"), data = Season_nbc_isbm_no_nas)

res_gam_drop_kept_spec <- simulateResiduals(nbc_isbm_model_full_gamma_spec, plot = T, quantreg=T)


```

### Central BC All seasons fisheries

```{r Modelling CBCsubsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-cbc-residuals
#| fig-cap: "Validation plots for CBC model"

Season_cbc_isbm<-Sport_mark_rate_finescale_combined%>% filter(YEAR %in% c(2015:2023)) %>% filter(finescale_fishery_old == "CBC S")

#Modelling comparisons need to be done on models with same # of NAs - so drop nas
Season_cbc_isbm_no_nas<-Season_cbc_isbm %>% drop_na(any_of(c("historic_summer", "status")))

#Chosen model:
cbc_isbm_model_full_gamma_spec<- glm(formula = catch_estimate + 1 ~  status + 1 + historic_summer,  family=Gamma(link = "log"), data = Season_cbc_isbm_no_nas)

res_gam_drop_kept_spec <- simulateResiduals(cbc_isbm_model_full_gamma_spec, plot = T, quantreg=T)
```

\clearpage

# MODEL OUTPUT {#app:third-appendix}

### Southern BC Summer fisheries

```{r forloop seasonal summer appendix, results='asis', echo=FALSE}

family.set <- family(Summer_model)
ilink.family.set<- family.set$linkinv

mod<-Summer_model

ndata<-Summer_south_no_nas %>% group_by(status, finescale_fishery_old, finescale_fishery, season) %>%  tidyr::expand(creel_plus_summer = seq(0, max(creel_plus_summer), length=100))
  

## add the fitted values by predicting from the model for the new data
ndata<- add_column(ndata, fit = predict(mod, newdata = ndata, type = 'response'))
ndata<- bind_cols(ndata, setNames(as_tibble(predict(mod, ndata, se.fit = TRUE)[1:2]),
                                                   c('fit_link','se_link')))

ndata <- mutate(ndata,
                fit_resp  = ilink.family.set(fit_link),
                right_upr = ilink.family.set(fit_link + (2 * se_link)),
                right_lwr = ilink.family.set(fit_link - (2 * se_link)))

fishery_name_south<- sort(unique(Summer_south_no_nas$finescale_fishery))

for (i in 1:length(fishery_name_south)){


dataminmax_marked_kept<- Summer_south_no_nas %>% filter(finescale_fishery==fishery_name_south[i], status=="marked_Kept_total")
dataminmax_marked_released<- Summer_south_no_nas %>% filter(finescale_fishery==fishery_name_south[i], status=="marked_Released_total")
dataminmax_unmarked_kept<- Summer_south_no_nas %>% filter(finescale_fishery==fishery_name_south[i], status=="unmarked_Kept_total")
dataminmax_unmarked_released<- Summer_south_no_nas %>% filter(finescale_fishery==fishery_name_south[i], status=="unmarked_Released_total")


g1<-ggplot(Summer_south_no_nas %>% filter(finescale_fishery==fishery_name_south[i], status=="marked_Kept_total"), aes(x=creel_plus_summer,  y= catch_estimate, col=status))+
  geom_point()+
  geom_line(ndata %>% filter(finescale_fishery==fishery_name_south[i], status=="marked_Kept_total"), mapping=aes(y= fit, x=creel_plus_summer, col=status))+
  geom_ribbon(ndata %>% filter(finescale_fishery==fishery_name_south[i], status=="marked_Kept_total"), mapping=aes(y= fit,x=creel_plus_summer, ymin = right_lwr, ymax = right_upr, fill=status), alpha = 0.10)+
   theme_bw() +
  scale_size_continuous(range=c(1,3))+ scale_colour_manual(values=c("#440154"))+  scale_fill_manual(values=c("#440154"))+
  coord_cartesian(ylim = c(0,max(dataminmax_marked_kept$catch_estimate)), xlim = c(0,max(dataminmax_marked_kept$creel_plus_summer))) + ggtitle( paste0("", fishery_name_south[fishery_name_south== fishery_name_south[i]], " "))

g2<-ggplot(Summer_south_no_nas %>% filter(finescale_fishery==fishery_name_south[i], status=="marked_Released_total"), aes(x=creel_plus_summer,  y= catch_estimate, col=status))+
  geom_point()+
  geom_line(ndata %>% filter(finescale_fishery==fishery_name_south[i], status=="marked_Released_total"), mapping=aes(y= fit, x=creel_plus_summer, col=status))+
  geom_ribbon(ndata %>% filter(finescale_fishery==fishery_name_south[i], status=="marked_Released_total"), mapping=aes(y= fit,x=creel_plus_summer, ymin = right_lwr, ymax = right_upr, fill=status), alpha = 0.10)+
  theme_bw() +
 scale_size_continuous(range=c(1,3))+ scale_colour_manual(values=c("#31688e"))+  scale_fill_manual(values=c("#31688e"))+
  coord_cartesian(ylim = c(0,max(dataminmax_marked_released$catch_estimate)), xlim = c(0,max(dataminmax_marked_released$creel_plus_summer)))

g3<-ggplot(Summer_south_no_nas %>% filter(finescale_fishery==fishery_name_south[i], status=="unmarked_Kept_total"), aes(x=creel_plus_summer,  y= catch_estimate, col=status))+
  geom_point()+
  geom_line(ndata %>% filter(finescale_fishery==fishery_name_south[i], status=="unmarked_Kept_total"), mapping=aes(y= fit, x=creel_plus_summer, col=status))+
  geom_ribbon(ndata %>% filter(finescale_fishery==fishery_name_south[i], status=="unmarked_Kept_total"), mapping=aes(y= fit,x=creel_plus_summer, ymin = right_lwr, ymax = right_upr, fill=status), alpha = 0.10)+
  theme_bw() +
  scale_size_continuous(range=c(1,3))+ scale_colour_manual(values=c("#35b779"))+  scale_fill_manual(values=c("#35b779"))+
  coord_cartesian(ylim = c(0,max(dataminmax_unmarked_kept$catch_estimate)), xlim = c(0,max(dataminmax_unmarked_kept$creel_plus_summer)))

g4<-ggplot(Summer_south_no_nas%>% filter(finescale_fishery==fishery_name_south[i], status=="unmarked_Released_total"), aes(x=creel_plus_summer,  y= catch_estimate, col=status))+
  geom_point()+
  geom_line(ndata %>% filter(finescale_fishery==fishery_name_south[i], status=="unmarked_Released_total"), mapping=aes(y= fit, x=creel_plus_summer, col=status))+
  geom_ribbon(ndata %>% filter(finescale_fishery==fishery_name_south[i], status=="unmarked_Released_total"), mapping=aes(y= fit,x=creel_plus_summer, ymin = right_lwr, ymax = right_upr, fill=status, col=status), alpha = 0.10)+
   theme_bw() +
  scale_size_continuous(range=c(1,3))+ scale_colour_manual(values=c("#d4be02"))+  scale_fill_manual(values=c("#d4be02"))+
    coord_cartesian(ylim = c(0,max(dataminmax_unmarked_released$catch_estimate)), xlim = c(0,max(dataminmax_unmarked_released$creel_plus_summer)))

g<-g1+g2 + g3+g4 + plot_layout(guides = "collect") 
print(g)
}
```

### Southern BC Spring and Fall fisheries

```{r forloop seasonal spring and fall appendix, results='asis', echo=FALSE}

family.set <- family(Spring_fall_model)
ilink.family.set<- family.set$linkinv

mod<-Spring_fall_model

ndata<-Season_south_no_nas %>% group_by(status, finescale_fishery_old, finescale_fishery, season) %>%  tidyr::expand(creel_plus_summer = seq(0, max(creel_plus_summer), length=100))
  

## add the fitted values by predicting from the model for the new data
ndata<- add_column(ndata, fit = predict(mod, newdata = ndata, type = 'response'))
ndata<- bind_cols(ndata, setNames(as_tibble(predict(mod, ndata, se.fit = TRUE)[1:2]),
                                                   c('fit_link','se_link')))

ndata <- mutate(ndata,
                fit_resp  = ilink.family.set(fit_link),
                right_upr = ilink.family.set(fit_link + (2 * se_link)),
                right_lwr = ilink.family.set(fit_link - (2 * se_link)))


fishery_name_south<- sort(unique(Season_south_no_nas$finescale_fishery))

  
for (i in 1:length(fishery_name_south)){
 
dataminmax_marked_kept<- Season_south_no_nas %>% filter(finescale_fishery==fishery_name_south[i], status=="marked_Kept_total")
dataminmax_marked_released<- Season_south_no_nas %>% filter(finescale_fishery==fishery_name_south[i], status=="marked_Released_total")
dataminmax_unmarked_kept<- Season_south_no_nas %>% filter(finescale_fishery==fishery_name_south[i], status=="unmarked_Kept_total")
dataminmax_unmarked_released<- Season_south_no_nas %>% filter(finescale_fishery==fishery_name_south[i], status=="unmarked_Released_total")


g1<-ggplot(Season_south_no_nas %>% filter(finescale_fishery==fishery_name_south[i], status=="marked_Kept_total"), aes(x=creel_plus_summer,  y= catch_estimate, col=status))+
  geom_point()+
  geom_line(ndata %>% filter(finescale_fishery==fishery_name_south[i], status=="marked_Kept_total"), mapping=aes(y= fit, x=creel_plus_summer, col=status))+
  geom_ribbon(ndata %>% filter(finescale_fishery==fishery_name_south[i], status=="marked_Kept_total"), mapping=aes(y= fit,x=creel_plus_summer, ymin = right_lwr, ymax = right_upr, fill=status), alpha = 0.10)+
   theme_bw() +
  scale_size_continuous(range=c(1,3))+ scale_colour_manual(values=c("#440154"))+  scale_fill_manual(values=c("#440154"))+
  coord_cartesian(ylim = c(0,max(dataminmax_marked_kept$catch_estimate)), xlim = c(0,max(dataminmax_marked_kept$creel_plus_summer))) + ggtitle( paste0("", fishery_name_south[fishery_name_south== fishery_name_south[i]], " "))

g2<-ggplot(Season_south_no_nas %>% filter(finescale_fishery==fishery_name_south[i], status=="marked_Released_total"), aes(x=creel_plus_summer,  y= catch_estimate, col=status))+
  geom_point()+
  geom_line(ndata %>% filter(finescale_fishery==fishery_name_south[i], status=="marked_Released_total"), mapping=aes(y= fit, x=creel_plus_summer, col=status))+
  geom_ribbon(ndata %>% filter(finescale_fishery==fishery_name_south[i], status=="marked_Released_total"), mapping=aes(y= fit,x=creel_plus_summer, ymin = right_lwr, ymax = right_upr, fill=status), alpha = 0.10)+
  theme_bw() +
 scale_size_continuous(range=c(1,3))+ scale_colour_manual(values=c("#31688e"))+  scale_fill_manual(values=c("#31688e"))+
  coord_cartesian(ylim = c(0,max(dataminmax_marked_released$catch_estimate)), xlim = c(0,max(dataminmax_marked_released$creel_plus_summer)))

g3<-ggplot(Season_south_no_nas %>% filter(finescale_fishery==fishery_name_south[i], status=="unmarked_Kept_total"), aes(x=creel_plus_summer,  y= catch_estimate, col=status))+
  geom_point()+
  geom_line(ndata %>% filter(finescale_fishery==fishery_name_south[i], status=="unmarked_Kept_total"), mapping=aes(y= fit, x=creel_plus_summer, col=status))+
  geom_ribbon(ndata %>% filter(finescale_fishery==fishery_name_south[i], status=="unmarked_Kept_total"), mapping=aes(y= fit,x=creel_plus_summer, ymin = right_lwr, ymax = right_upr, fill=status), alpha = 0.10)+
  theme_bw() +
  scale_size_continuous(range=c(1,3))+ scale_colour_manual(values=c("#35b779"))+  scale_fill_manual(values=c("#35b779"))+
  coord_cartesian(ylim = c(0,max(dataminmax_unmarked_kept$catch_estimate)), xlim = c(0,max(dataminmax_unmarked_kept$creel_plus_summer)))

g4<-ggplot(Season_south_no_nas %>% filter(finescale_fishery==fishery_name_south[i], status=="unmarked_Released_total"), aes(x=creel_plus_summer,  y= catch_estimate, col=status))+
  geom_point()+
  geom_line(ndata %>% filter(finescale_fishery==fishery_name_south[i], status=="unmarked_Released_total"), mapping=aes(y= fit, x=creel_plus_summer, col=status))+
  geom_ribbon(ndata %>% filter(finescale_fishery==fishery_name_south[i], status=="unmarked_Released_total"), mapping=aes(y= fit,x=creel_plus_summer, ymin = right_lwr, ymax = right_upr, fill=status, col=status), alpha = 0.10)+
   theme_bw() +
  scale_size_continuous(range=c(1,3))+ scale_colour_manual(values=c("#d4be02"))+  scale_fill_manual(values=c("#d4be02"))+
    coord_cartesian(ylim = c(0,max(dataminmax_unmarked_released$catch_estimate)), xlim = c(0,max(dataminmax_unmarked_released$creel_plus_summer)))

g<-g1+g2 + g3+g4 + plot_layout(guides = "collect") 
print(g)
}
```

### Northern BC AABM Spring and Fall fisheries

```{r forloop NBC AABM seasonal spring and fall appendix, results='asis', echo=FALSE}

family.set <- family(North_aabm_model)
ilink.family.set<- family.set$linkinv

mod<-North_aabm_model

ndata<-Season_north_aabm_no_nas %>% group_by(status, finescale_fishery_old, finescale_fishery, season) %>%  tidyr::expand(historic_summer = seq(0, max(historic_summer), length=100))


## add the fitted values by predicting from the model for the new data
ndata<- add_column(ndata, fit = predict(mod, newdata = ndata, type = 'response'))
ndata<- bind_cols(ndata, setNames(as_tibble(predict(mod, ndata, se.fit = TRUE)[1:2]),
                                                   c('fit_link','se_link')))

ndata <- mutate(ndata,
                fit_resp  = ilink.family.set(fit_link),
                right_upr = ilink.family.set(fit_link + (2 * se_link)),
                right_lwr = ilink.family.set(fit_link - (2 * se_link)))


fishery_name_north_aabm<- sort(unique(Season_north_aabm_no_nas$finescale_fishery))

for (i in 1:length(fishery_name_north_aabm)){



dataminmax_marked_kept<- Season_north_aabm_no_nas %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="marked_Kept_total")
dataminmax_marked_released<- Season_north_aabm_no_nas %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="marked_Released_total")
dataminmax_unmarked_kept<- Season_north_aabm_no_nas %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="unmarked_Kept_total")
dataminmax_unmarked_released<- Season_north_aabm_no_nas %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="unmarked_Released_total")

dataminmax_marked_kept<- Season_north_aabm_no_nas %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="marked_Kept_total")
dataminmax_marked_released<- Season_north_aabm_no_nas %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="marked_Released_total")
dataminmax_unmarked_kept<- Season_north_aabm_no_nas %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="unmarked_Kept_total")
dataminmax_unmarked_released<- Season_north_aabm_no_nas %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="unmarked_Released_total")

g1<-ggplot(Season_north_aabm_no_nas %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="marked_Kept_total"), aes(x=historic_summer,  y= catch_estimate, col=status))+
  geom_point()+
  geom_line(ndata %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="marked_Kept_total"), mapping=aes(y= fit, x=historic_summer, col=status))+
  geom_ribbon(ndata %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="marked_Kept_total"), mapping=aes(y= fit,x=historic_summer, ymin = right_lwr, ymax = right_upr, fill=status), alpha = 0.10)+
   theme_bw() +
  scale_size_continuous(range=c(1,3))+ scale_colour_manual(values=c("#440154"))+  scale_fill_manual(values=c("#440154"))+
  coord_cartesian(ylim = c(0,max(dataminmax_marked_kept$catch_estimate)), xlim = c(0,max(dataminmax_marked_kept$historic_summer))) + ggtitle(paste0(fishery_name_north_aabm[fishery_name_north_aabm== fishery_name_north_aabm[i]], " "))

g2<-ggplot(Season_north_aabm_no_nas %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="marked_Released_total"), aes(x=historic_summer,  y= catch_estimate, col=status))+
  geom_point()+
  geom_line(ndata %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="marked_Released_total"), mapping=aes(y= fit, x=historic_summer, col=status))+
  geom_ribbon(ndata %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="marked_Released_total"), mapping=aes(y= fit,x=historic_summer, ymin = right_lwr, ymax = right_upr, fill=status), alpha = 0.10)+
  theme_bw() +
 scale_size_continuous(range=c(1,3))+ scale_colour_manual(values=c("#31688e"))+  scale_fill_manual(values=c("#31688e"))+
  coord_cartesian(ylim = c(0,max(dataminmax_marked_released$catch_estimate)), xlim = c(0,max(dataminmax_marked_released$historic_summer)))

g3<-ggplot(Season_north_aabm_no_nas %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="unmarked_Kept_total"), aes(x=historic_summer,  y= catch_estimate, col=status))+
  geom_point()+
  geom_line(ndata %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="unmarked_Kept_total"), mapping=aes(y= fit, x=historic_summer, col=status))+
  geom_ribbon(ndata %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="unmarked_Kept_total"), mapping=aes(y= fit,x=historic_summer, ymin = right_lwr, ymax = right_upr, fill=status), alpha = 0.10)+
  theme_bw() +
  scale_size_continuous(range=c(1,3))+ scale_colour_manual(values=c("#35b779"))+  scale_fill_manual(values=c("#35b779"))+
  coord_cartesian(ylim = c(0,max(dataminmax_unmarked_kept$catch_estimate)), xlim = c(0,max(dataminmax_unmarked_kept$historic_summer)))

g4<-ggplot(Season_north_aabm_no_nas %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="unmarked_Released_total"), aes(x=historic_summer,  y= catch_estimate, col=status))+
  geom_point()+
  geom_line(ndata %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="unmarked_Released_total"), mapping=aes(y= fit, x=historic_summer, col=status))+
  geom_ribbon(ndata %>% filter(finescale_fishery==fishery_name_north_aabm[i], status=="unmarked_Released_total"), mapping=aes(y= fit,x=historic_summer, ymin = right_lwr, ymax = right_upr, fill=status, col=status), alpha = 0.10)+
   theme_bw() +
  scale_size_continuous(range=c(1,3))+ scale_colour_manual(values=c("#d4be02"))+  scale_fill_manual(values=c("#d4be02"))+
    coord_cartesian(ylim = c(0,max(dataminmax_unmarked_released$catch_estimate)), xlim = c(0,max(dataminmax_unmarked_released$historic_summer)))

g<-g1+g2 + g3+g4 + plot_layout(guides = "collect") 
print(g)
}
```

### Northern BC ISBM Spring and Fall fisheries

```{r forloop NBC ISBM seasonal spring and fall appendix, results='asis', echo=FALSE}


family.set <- family(nbc_isbm_model_full_gamma_spec)
ilink.family.set<- family.set$linkinv

mod<-nbc_isbm_model_full_gamma_spec

ndata<-Season_north_isbm_no_nas %>% group_by(status, finescale_fishery_old, finescale_fishery, season) %>%  tidyr::expand(historic_summer = seq(0, max(historic_summer), length=100))

## add the fitted values by predicting from the model for the new data
ndata<- add_column(ndata, fit = predict(mod, newdata = ndata, type = 'response'))
ndata<- bind_cols(ndata, setNames(as_tibble(predict(mod, ndata, se.fit = TRUE)[1:2]),
                                                   c('fit_link','se_link')))

ndata <- mutate(ndata,
                fit_resp  = ilink.family.set(fit_link),
                right_upr = ilink.family.set(fit_link + (2 * se_link)),
                right_lwr = ilink.family.set(fit_link - (2 * se_link)))


fishery_name_north_isbm<- sort(unique(Season_north_isbm_no_nas$finescale_fishery))

for (i in 1:length(fishery_name_north_isbm)){

 
dataminmax_marked_kept<- Season_north_isbm_no_nas %>% filter(finescale_fishery==fishery_name_north_isbm[i], status=="marked_Kept_total")
dataminmax_marked_released<- Season_north_isbm_no_nas %>% filter(finescale_fishery==fishery_name_north_isbm[i], status=="marked_Released_total")
dataminmax_unmarked_kept<- Season_north_isbm_no_nas %>% filter(finescale_fishery==fishery_name_north_isbm[i], status=="unmarked_Kept_total")
dataminmax_unmarked_released<- Season_north_isbm_no_nas %>% filter(finescale_fishery==fishery_name_north_isbm[i], status=="unmarked_Released_total")


g1<-ggplot(Season_north_isbm_no_nas %>% filter(finescale_fishery==fishery_name_north_isbm[i], status=="marked_Kept_total"), aes(x=historic_summer,  y= catch_estimate, col=status))+
  geom_point()+
  geom_line(ndata %>% filter(finescale_fishery==fishery_name_north_isbm[i], status=="marked_Kept_total"), mapping=aes(y= fit, x=historic_summer, col=status))+
  geom_ribbon(ndata %>% filter(finescale_fishery==fishery_name_north_isbm[i], status=="marked_Kept_total"), mapping=aes(y= fit,x=historic_summer, ymin = right_lwr, ymax = right_upr, fill=status), alpha = 0.10)+
   theme_bw() +
  scale_size_continuous(range=c(1,3))+ scale_colour_manual(values=c("#440154"))+  scale_fill_manual(values=c("#440154"))+
  coord_cartesian(ylim = c(0,max(dataminmax_marked_kept$catch_estimate)), xlim = c(0,max(dataminmax_marked_kept$historic_summer))) + ggtitle(paste0(fishery_name_north_isbm[fishery_name_north_isbm== fishery_name_north_isbm[i]], " "))

g2<-ggplot(Season_north_isbm_no_nas %>% filter(finescale_fishery==fishery_name_north_isbm[i], status=="marked_Released_total"), aes(x=historic_summer,  y= catch_estimate, col=status))+
  geom_point()+
  geom_line(ndata %>% filter(finescale_fishery==fishery_name_north_isbm[i], status=="marked_Released_total"), mapping=aes(y= fit, x=historic_summer, col=status))+
  geom_ribbon(ndata %>% filter(finescale_fishery==fishery_name_north_isbm[i], status=="marked_Released_total"), mapping=aes(y= fit,x=historic_summer, ymin = right_lwr, ymax = right_upr, fill=status), alpha = 0.10)+
  theme_bw() +
 scale_size_continuous(range=c(1,3))+ scale_colour_manual(values=c("#31688e"))+  scale_fill_manual(values=c("#31688e"))+
  coord_cartesian(ylim = c(0,max(dataminmax_marked_released$catch_estimate)), xlim = c(0,max(dataminmax_marked_released$historic_summer)))

g3<-ggplot(Season_north_isbm_no_nas %>% filter(finescale_fishery==fishery_name_north_isbm[i], status=="unmarked_Kept_total"), aes(x=historic_summer,  y= catch_estimate, col=status))+
  geom_point()+
  geom_line(ndata %>% filter(finescale_fishery==fishery_name_north_isbm[i], status=="unmarked_Kept_total"), mapping=aes(y= fit, x=historic_summer, col=status))+
  geom_ribbon(ndata %>% filter(finescale_fishery==fishery_name_north_isbm[i], status=="unmarked_Kept_total"), mapping=aes(y= fit,x=historic_summer, ymin = right_lwr, ymax = right_upr, fill=status), alpha = 0.10)+
  theme_bw() +
  scale_size_continuous(range=c(1,3))+ scale_colour_manual(values=c("#35b779"))+  scale_fill_manual(values=c("#35b779"))+
  coord_cartesian(ylim = c(0,max(dataminmax_unmarked_kept$catch_estimate)), xlim = c(0,max(dataminmax_unmarked_kept$historic_summer)))

g4<-ggplot(Season_north_isbm_no_nas %>% filter(finescale_fishery==fishery_name_north_isbm[i], status=="unmarked_Released_total"), aes(x=historic_summer,  y= catch_estimate, col=status))+
  geom_point()+
  geom_line(ndata %>% filter(finescale_fishery==fishery_name_north_isbm[i], status=="unmarked_Released_total"), mapping=aes(y= fit, x=historic_summer, col=status))+
  geom_ribbon(ndata %>% filter(finescale_fishery==fishery_name_north_isbm[i], status=="unmarked_Released_total"), mapping=aes(y= fit,x=historic_summer, ymin = right_lwr, ymax = right_upr, fill=status, col=status), alpha = 0.10)+
   theme_bw() +
  scale_size_continuous(range=c(1,3))+ scale_colour_manual(values=c("#d4be02"))+  scale_fill_manual(values=c("#d4be02"))+
    coord_cartesian(ylim = c(0,max(dataminmax_unmarked_released$catch_estimate)), xlim = c(0,max(dataminmax_unmarked_released$historic_summer)))

g<-g1+g2 + g3+g4 + plot_layout(guides = "collect") 
print(g)
}
```

`r if(knitr:::is_latex_output()) '% end csasdown appendix'`
