`r if(knitr:::is_latex_output()) '% begin csasdown appendix'` `r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'`

\clearpage

# Stratification Mapping

```{r methods strattab, echo=FALSE, message=FALSE, warning=FALSE}
#| label: strattab

Sport_mark_rate_finescale<-wrangle_rec_catch("pacRecCatch_db.yaml", by="area")

Strat_table <- 
  Sport_mark_rate_finescale %>% 
  dplyr::select(finescale_fishery_old, finescale_fishery, AREA, MONTH) %>% 
  unique() %>% 
  arrange(finescale_fishery_old, finescale_fishery,  AREA, MONTH)

fishery_map <-
  Sport_mark_rate_finescale %>% 
  dplyr::select(finescale_fishery_old, finescale_fishery) %>%
  mutate(erafishery = case_when(
    finescale_fishery_old %in% c("NWCVI S AABM", "SWCVI S AABM") ~ "WCVI AABM S",
    finescale_fishery_old %in% c("NGS S", "SGS S") ~  "GEO ST S",
    finescale_fishery_old == "CA JDF S" ~ "BC JF S",
    finescale_fishery_old %in%  c("NWCVI S ISBM", "SWCVI S ISBM") ~ "WCVI ISBM S",
    TRUE ~ finescale_fishery_old)) %>% 
  unique()

Strat_table <- 
  Strat_table %>% 
  left_join(fishery_map) %>% 
  arrange(erafishery) %>% 
  relocate(erafishery) %>%
  group_by(finescale_fishery) %>% 
  mutate(month_category = case_when(
    finescale_fishery == "CBC S" ~ "1:12",
    finescale_fishery == "NWCVI S SUMMER AABM" & AREA %in% c("Area 25", "Area 26", "Area 27") ~ "5:6",
    finescale_fishery == "NWCVI S SUMMER ISBM"  ~ "7:9",
    finescale_fishery == "SWCVI S SUMMER AABM" & AREA %in% c("Area 21", "Area 23", "Area 23 (Alberni Canal)", "Area 23 (Barkley)", "Area 24") ~ "5:7",
    finescale_fishery == "SWCVI S SUMMER ISBM"  ~ "8:9",
    MONTH %in% c(10,11,12) ~ "10:12", 
    MONTH %in% c(1:4) ~ "1:4", 
    MONTH %in% c(5:9) ~ "5:9",
    MONTH %in% c(7:8) ~ "7:8",
    TRUE ~ "NA"
  )) %>% 
  dplyr::select(-MONTH) %>% unique() %>% drop_na() %>% 
  rename("ERA Fishery" = "erafishery", "Finescale Fishery Old" = "finescale_fishery_old", 
         "Finescale Fishery New" ="finescale_fishery", "PFMA" ="AREA","Months" ="month_category")


Strat_table %>%
  kableExtra::kable("html", 
                    escape = FALSE,
                    caption = "Stratification mapping",
                    align = "l") %>%
  kableExtra::kable_classic(full_width = F, html_font = "Cambria") %>%
  kableExtra::kable_styling(bootstrap_options = c("bordered")) %>%
  kableExtra::collapse_rows(columns = 1:3, valign = "middle")


```

# REVISED ESTIMATES 2013-2023 {#app:first-appendix}

```{r m plots, results='asis', echo=FALSE}
#| label: fig-catch-revised-1-appendix
#| fig-cap: "Revised yearly kept catch estimates"


library(ROracle)
library(pacRecCatch)
library(tidyverse)
library(odbc)
library(lubridate)
library(ggnewscale)
library(patchwork)
library(DHARMa)
library(lme4)
library(bbmle)
library(SuppDists)
library(MuMIn)

models_combined <- pacRecCatch::model_rec_catch(Sport_mark_rate_finescale_combined)


### make this by finescale fishery old... sum up. 
models_combined_old<-models_combined %>% group_by(YEAR,kept_status,pred_cat, finescale_fishery_old) %>% summarise_if(is.numeric, sum, na.rm = TRUE) %>% filter(!str_detect(finescale_fishery_old, "CBC|NBC"))

fishery_name_south<- sort(unique(models_combined_old$finescale_fishery_old))


for (i in 1:length(fishery_name_south)){

 m<-ggplot(graph_data <- models_combined_old  %>% 
             filter(finescale_fishery_old == fishery_name_south[i],
                    YEAR %in% c(2013:2023),
                    kept_status == "Kept") %>%
  ungroup() %>%
  select(YEAR, creel_unfiltered_plus, catch_estimate_predicted) %>%
  pivot_longer(c(creel_unfiltered_plus, catch_estimate_predicted),
               names_to = "estimate_type",
               values_to = "catch_estimate"))
 
m <- ggplot(graph_data, aes(y = catch_estimate,
           x = YEAR,
           shape = estimate_type,
           linetype = estimate_type))+
  geom_point(size = 2.5,
             colour = "#440154")+
  geom_line(colour = "#440154") +
  scale_shape_manual(values = c(1, 16),
                     labels = c("iREC included", "Unfiltered creel and logbook")
                     ) +
  scale_linetype_manual(values = c(2, 1),
                      labels = c("iREC included", "Unfiltered creel and logbook")
                      ) +
  scale_x_continuous(breaks = c(2013, 2015, 2017, 2019, 2021, 2023)) +
  ggtitle(paste(fishery_name_south[i])) + 
  theme_bw() + 
  ylab("Kept catch estimate") + 
  xlab("Year") +
  labs(linetype = "Catch Estimate", shape = "Catch Estimate")
 
 
 
 print(m)
}
```

```{r m plots, results='asis', echo=FALSE}
#| label: fig-catch-revised-2-appendix
#| fig-cap: "revised yearly catch estimates"



### make this by finescale fishery old... sum up. 
models_combined_old2<-models_combined %>% group_by(YEAR,kept_status,pred_cat, finescale_fishery_old) %>% summarise_if(is.numeric, sum, na.rm = TRUE) %>% filter(str_detect(finescale_fishery_old, "CBC|NBC"))

fishery_name_north<- sort(unique(models_combined_old2$finescale_fishery_old))


for (i in 1:length(fishery_name_north)){

 m<-ggplot(graph_data <- models_combined_old2  %>% 
             filter(finescale_fishery_old == fishery_name_north[i],
                    YEAR %in% c(2013:2023),
                    kept_status == "Kept") %>%
  ungroup() %>%
  select(YEAR, historic_plus, catch_estimate_predicted) %>%
  pivot_longer(c(historic_plus, catch_estimate_predicted),
               names_to = "estimate_type",
               values_to = "catch_estimate"))
 
m <- ggplot(graph_data, aes(y = catch_estimate,
           x = YEAR,
           shape = estimate_type,
           linetype = estimate_type))+
  geom_point(size = 2.5,
             colour = "#440154")+
  geom_line(colour = "#440154") +
  scale_shape_manual(values = c(1, 16),
                     labels = c("iREC included", "Historic")
                     ) +
  scale_linetype_manual(values = c(2, 1),
                      labels = c("iREC included", "Historic")
                      ) +
  scale_x_continuous(breaks = c(2013, 2015, 2017, 2019, 2021, 2023)) +
  ggtitle(paste(fishery_name_north[i])) + 
  theme_bw() + 
  ylab("Kept catch estimate") + 
  xlab("Year") +
  labs(linetype = "Catch Estimate", shape = "Catch Estimate")
 
 
 
 print(m)
}
```

\clearpage

# MODEL RESIDUALS {#app:second-appendix}

### Southern BC Summer fisheries <br/>

```{r Modelling summer subsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-south-sum-residuals
#| fig-cap: "Validation plots for BC South Summer model"


Summer_south<-Sport_mark_rate_finescale_combined %>%
  filter(YEAR %in% c(2013:2023))%>%
  filter(!str_detect(finescale_fishery, "CBC|NBC")) %>%
  filter(season %in% c("summer"))

Summer_south_no_nas<-Summer_south %>% drop_na(any_of(c("creel_plus_summer", "status", "finescale_fishery_old")))
write_rds(Summer_south_no_nas, "Summer_south_no_nas.RDS")

Summer_model<- glm(formula = catch_estimate ~ finescale_fishery_old + status + creel_plus_summer:finescale_fishery_old + creel_plus_summer:status +
finescale_fishery_old:status + 1 + creel_plus_summer,  family=Gamma(link = "log"), data = Summer_south_no_nas)

res_gam <- simulateResiduals(Summer_model, plot = T, quantreg=T)
```

### Northern BC AABM Summer fisheries <br/>

```{r Modelling summer nbc aabm subsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-nbcaabm-sum-residuals
#| fig-cap: "Validation plots for NBC AABM Summer model"

Summer_north_aabm<-Sport_mark_rate_finescale_combined%>%
  filter(YEAR %in% c(2015:2023)) %>%
  filter(finescale_fishery_old == "NBC AABM S")%>%
  filter( season=="summer")

#Modelling comparisons need to be done on models with same # of NAs - so drop nas
Summer_north_aabm_no_nas<-Summer_north_aabm %>% drop_na(any_of(c("historic_summer", "status", "finescale_fishery_old", "season", "historic_effort")))

###Chosen model
Summer_north_aabm_model_gamma_spec<- glm(formula = catch_estimate + 1 ~ status + historic_summer:status + 1 + historic_summer,  family=Gamma(link = "log"), data = Summer_north_aabm_no_nas)

res_gam <- simulateResiduals(Summer_north_aabm_model_gamma_spec, plot = T, quantreg=T)

```

### Northern BC ISBM Summer fisheries <br/>

```{r Modelling summer nbc isbm subsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-nbcisbm-sum-residuals
#| fig-cap: "Validation plots for NBC ISBM Summer model"

Summer_north_isbm<-Sport_mark_rate_finescale_combined%>% filter(YEAR %in% c(2015:2023)) %>% filter(finescale_fishery_old == "NBC ISBM S")%>%
  filter(season %in% c("summer"))
Summer_north_isbm_no_nas<-Summer_north_isbm %>% drop_na(any_of(c("historic_summer", "status", "finescale_fishery_old", "season", "historic_effort")))
write_rds(Summer_north_isbm_no_nas, "Summer_north_isbm_no_nas.RDS")

#Chosen model
Summer_north_isbm_model_full_gamma_spec<- glm(formula = catch_estimate + 1 ~ status + 1 + historic_summer,  family=Gamma(link = "log"), data = Summer_north_isbm_no_nas)
res_spec <- simulateResiduals(Summer_north_isbm_model_full_gamma_spec, plot = T, quantreg=T)


```

### Southern BC Spring and Fall fisheries <br/>

```{r Modelling spring and fall subsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-south-sf-residuals
#| fig-cap: "Validation plots for BC South Spring and Fall model"

### Data needed
Season_south_sf<-Sport_mark_rate_finescale_combined %>%
  filter(YEAR %in% c(2013:2023))%>%
  filter(!str_detect(finescale_fishery, "CBC|NBC")) %>%
  filter(season %in% c("spring", "fall"))

#Modelling comparisons need to be done on models with same # of NAs - so drop nas
Season_south_no_nas<-Season_south_sf %>% drop_na(any_of(c("creel_plus_summer", "status", "finescale_fishery_old", "season")))

Spring_fall_model<-  glm(formula =catch_estimate + 3 ~ finescale_fishery_old + season +   status + creel_plus_summer:finescale_fishery_old + creel_plus_summer:status +  finescale_fishery_old:season + 1 + creel_plus_summer,  family=Gamma(link = "log"), data = Season_south_no_nas)

res_gam <- simulateResiduals(Spring_fall_model, plot = T, quantreg=T)
```

### Northern BC AABM Spring and Fall fisheries <br/>

```{r Modelling NBC AABM spring and fall subsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-nbcaabm-sf-residuals
#| fig-cap: "Validation plots for NBC AABM Spring and Fall model"

Season_north_aabm<-Sport_mark_rate_finescale_combined%>% filter(YEAR %in% c(2015:2023)) %>% filter(finescale_fishery_old == "NBC AABM S") %>%
  filter(season %in% c("spring", "fall"))
Season_north_aabm_no_nas<-Season_north_aabm %>% drop_na(any_of(c("historic_summer", "status", "finescale_fishery_old", "season", "historic_effort")))


###Chosen model
North_aabm_model<- glm(formula = catch_estimate + 1 ~ season + status + historic_summer:season + 1 + historic_summer,  family=Gamma(link = "log"), data = Season_north_aabm_no_nas)

res_gam <- simulateResiduals(North_aabm_model, plot = T, quantreg=T)
```

### Northern BC ISBM Spring and Fall fisheries <br/>

```{r Modelling NBC ISBM spring and fall subsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-nbcisbm-sf-residuals
#| fig-cap: "Validation plots for NBC ISBM Spring and Fall model"


Season_nbc_isbm<-Sport_mark_rate_finescale_combined%>% filter(YEAR %in% c(2015:2023)) %>% filter(finescale_fishery_old == "NBC ISBM S")%>%
  filter(season %in% c("spring", "fall"))
Season_nbc_isbm_no_nas<-Season_nbc_isbm %>% drop_na(any_of(c("historic_summer", "status", "finescale_fishery_old", "season", "historic_effort")))


###Chosen model
nbc_isbm_model_full_gamma_spec<- glm(formula = catch_estimate + 1 ~ season + status + season:status + 1 + historic_summer,  family=Gamma(link = "log"), data = Season_nbc_isbm_no_nas)

res_gam_drop_kept_spec <- simulateResiduals(nbc_isbm_model_full_gamma_spec, plot = T, quantreg=T)


```

### Central BC All seasons fisheries <br/>

```{r Modelling CBCsubsetting appendix, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
#| label: fig-cbc-residuals
#| fig-cap: "Validation plots for CBC model"

Season_cbc_isbm<-Sport_mark_rate_finescale_combined%>% filter(YEAR %in% c(2015:2023)) %>% filter(finescale_fishery_old == "CBC S")

#Modelling comparisons need to be done on models with same # of NAs - so drop nas
Season_cbc_isbm_no_nas<-Season_cbc_isbm %>% drop_na(any_of(c("historic_summer", "status")))

#Chosen model:
cbc_isbm_model_full_gamma_spec<- glm(formula = catch_estimate + 1 ~  status + 1 + historic_summer,  family=Gamma(link = "log"), data = Season_cbc_isbm_no_nas)

res_gam_drop_kept_spec <- simulateResiduals(cbc_isbm_model_full_gamma_spec, plot = T, quantreg=T)
```

\clearpage

# MODEL OUTPUT {#app:third-appendix}

```{r plotting_function, include = FALSE}
#plotting function that will be used for the next few sections
create_facet_plot <- function(finescale_fishery_name, x_var, plot_data, model_data) {
  # Labels for facets
  status_labels <- c(
    "marked_Kept_total" = "Marked Kept Total",
    "marked_Released_total" = "Marked Released Total",
    "unmarked_Kept_total" = "Unmarked Kept Total",
    "unmarked_Released_total" = "Unmarked Released Total"
  )
  
  # Colors for status
  status_colors <- c(
    "marked_Kept_total" = "#440154", 
    "marked_Released_total" = "#31688e",
    "unmarked_Kept_total" = "#35b779",
    "unmarked_Released_total" = "#d4be02"
  )
  
  # Filter data for the specific finescale fishery
  filtered_data <- plot_data %>% filter(finescale_fishery == finescale_fishery_name)
  filtered_model_data <- model_data %>% filter(finescale_fishery == finescale_fishery_name)
  
  #check if all catch_estimate values are zero for each status
  #wont want ribbon and line if all data are 0 like in figure 10
  has_nonzero <- filtered_data %>%
    group_by(status) %>%
    summarize(has_nonzero = any(catch_estimate != 0), .groups = "drop")
  
  # Create the plot
  facet_plot <- ggplot(filtered_data, aes_string(x = x_var, y = "catch_estimate", col = "status", fill = "status")) +
    geom_point() +
    theme_bw() +
    scale_size_continuous(range = c(1, 3)) +
    scale_colour_manual(values = status_colors, guide = "none") +
    scale_fill_manual(values = status_colors, guide = "none") +
    xlab("Creel summer catch estimate") +
    ylab("iREC-only catch estimate") +
    facet_wrap(~status,
               labeller = labeller(status = status_labels),
               scales = "free")
  
  # Conditionally add geom_line and geom_ribbon for statuses with non-zero data
  for (status in unique(filtered_data$status)) {
    if (has_nonzero %>% filter(status == !!status) %>% pull(has_nonzero)) {
      facet_plot <- facet_plot +
        geom_line(data = filtered_model_data %>% filter(status == !!status),
                  aes_string(y = "fit", x = x_var, col = "status")) +
        geom_ribbon(data = filtered_model_data %>% filter(status == !!status),
                    aes_string(y = "fit", x = x_var, ymin = "right_lwr", ymax = "right_upr", fill = "status"), 
                    alpha = 0.10)
    }
  }
  
  return(facet_plot)
}
```

### Southern BC Summer fisheries

```{r forloop seasonal summer appendix, results='asis', echo=FALSE}

family.set <- family(Summer_model)
ilink.family.set<- family.set$linkinv

mod<-Summer_model

ndata<-Summer_south_no_nas %>% group_by(status, finescale_fishery_old, finescale_fishery, season) %>%  tidyr::expand(creel_plus_summer = seq(0, max(creel_plus_summer), length=100))
  

## add the fitted values by predicting from the model for the new data
ndata<- add_column(ndata, fit = predict(mod, newdata = ndata, type = 'response'))
ndata<- bind_cols(ndata, setNames(as_tibble(predict(mod, ndata, se.fit = TRUE)[1:2]),
                                                   c('fit_link','se_link')))

ndata <- mutate(ndata,
                fit_resp  = ilink.family.set(fit_link),
                right_upr = ilink.family.set(fit_link + (2 * se_link)),
                right_lwr = ilink.family.set(fit_link - (2 * se_link)))

fishery_name_south<- sort(unique(Summer_south_no_nas$finescale_fishery))

for (i in 1:length(fishery_name_south)){

plot <- create_facet_plot(finescale_fishery_name = fishery_name_south[i], 
                          x_var = "creel_plus_summer", 
                          plot_data = Summer_south_no_nas, 
                          model_data = ndata) +
  ggtitle(paste(fishery_name_south[i])) + 
  #changing the y lab
  ylab("IREC-included catch estimate")

print(plot)
}
```

### Southern BC Spring and Fall fisheries

```{r forloop seasonal spring and fall appendix, results='asis', echo=FALSE}

family.set <- family(Spring_fall_model)
ilink.family.set<- family.set$linkinv

mod<-Spring_fall_model

ndata<-Season_south_no_nas %>% group_by(status, finescale_fishery_old, finescale_fishery, season) %>%  tidyr::expand(creel_plus_summer = seq(0, max(creel_plus_summer), length=100))
  

## add the fitted values by predicting from the model for the new data
ndata<- add_column(ndata, fit = predict(mod, newdata = ndata, type = 'response'))
ndata<- bind_cols(ndata, setNames(as_tibble(predict(mod, ndata, se.fit = TRUE)[1:2]),
                                                   c('fit_link','se_link')))

ndata <- mutate(ndata,
                fit_resp  = ilink.family.set(fit_link),
                right_upr = ilink.family.set(fit_link + (2 * se_link)),
                right_lwr = ilink.family.set(fit_link - (2 * se_link)))


fishery_name_south<- sort(unique(Season_south_no_nas$finescale_fishery))

for (i in 1:length(fishery_name_south)){

plot <- create_facet_plot(finescale_fishery_name = fishery_name_south[i], 
                          x_var = "creel_plus_summer", 
                          plot_data = Season_south_no_nas, 
                          model_data = ndata) +
  ggtitle(paste(fishery_name_south[i]))

print(plot)
}

```

### Northern BC AABM Spring and Fall fisheries

```{r forloop NBC AABM seasonal spring and fall appendix, results='asis', echo=FALSE}

family.set <- family(North_aabm_model)
ilink.family.set<- family.set$linkinv

mod<-North_aabm_model

ndata<-Season_north_aabm_no_nas %>% group_by(status, finescale_fishery_old, finescale_fishery, season) %>%  tidyr::expand(historic_summer = seq(0, max(historic_summer), length=100))


## add the fitted values by predicting from the model for the new data
ndata<- add_column(ndata, fit = predict(mod, newdata = ndata, type = 'response'))
ndata<- bind_cols(ndata, setNames(as_tibble(predict(mod, ndata, se.fit = TRUE)[1:2]),
                                                   c('fit_link','se_link')))

ndata <- mutate(ndata,
                fit_resp  = ilink.family.set(fit_link),
                right_upr = ilink.family.set(fit_link + (2 * se_link)),
                right_lwr = ilink.family.set(fit_link - (2 * se_link)))


fishery_name_north_aabm<- sort(unique(Season_north_aabm_no_nas$finescale_fishery))


for (i in 1:length(fishery_name_north_aabm)){

plot <- create_facet_plot(finescale_fishery_name = fishery_name_north_aabm[i], 
                          x_var = "historic_summer", 
                          plot_data = Season_north_aabm_no_nas, 
                          model_data = ndata) +
  ggtitle(paste(fishery_name_north_aabm[i]))

print(plot)
}
```

### Northern BC ISBM Spring and Fall fisheries

```{r forloop NBC ISBM seasonal spring and fall appendix, results='asis', echo=FALSE}


family.set <- family(nbc_isbm_model_full_gamma_spec)
ilink.family.set<- family.set$linkinv

mod<-nbc_isbm_model_full_gamma_spec

ndata<-Season_north_isbm_no_nas %>% group_by(status, finescale_fishery_old, finescale_fishery, season) %>%  tidyr::expand(historic_summer = seq(0, max(historic_summer), length=100))

## add the fitted values by predicting from the model for the new data
ndata<- add_column(ndata, fit = predict(mod, newdata = ndata, type = 'response'))
ndata<- bind_cols(ndata, setNames(as_tibble(predict(mod, ndata, se.fit = TRUE)[1:2]),
                                                   c('fit_link','se_link')))

ndata <- mutate(ndata,
                fit_resp  = ilink.family.set(fit_link),
                right_upr = ilink.family.set(fit_link + (2 * se_link)),
                right_lwr = ilink.family.set(fit_link - (2 * se_link)))


fishery_name_north_isbm<- sort(unique(Season_north_isbm_no_nas$finescale_fishery))

for (i in 1:length(fishery_name_north_isbm)){

plot <- create_facet_plot(finescale_fishery_name = fishery_name_north_isbm[i], 
                          x_var = "historic_summer", 
                          plot_data = Season_north_isbm_no_nas, 
                          model_data = ndata) +
  ggtitle(paste(fishery_name_north_isbm[i]))

print(plot)
}
```

`r if(knitr:::is_latex_output()) '% end csasdown appendix'`
